---
title: "iLand landscape building"
author: "Winslow Hansen"
date: "May 26, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

libraries
```{r}
library(tidyverse)
library(terra)
library(sf)
```

Read in the the temperature product as a template
```{r}
gcm = "NorEsm2-MM"   #"NorEsm2-MM" "TaiESM1"


temp = rast(paste0("Z:/project_data/downscaling/CPCRW/Downscaled ",gcm,"/ssp245/tasmax/",gcm,"-ssp245-tasmax-1950.nc"))
#day.length=rast("Z:/project_data/downscaling/CPCRW/Daymet daily data/clipped_daymet_v4_daily_dayl.nc")
```
generate a climate names grid
```{r}

climate.grid=temp[[4]]
```

Read in an environment grid
```{r}
env.grid= rast("D:/workspace/Winslow/iLand_AK/cpcrw/materials/env/env.grid.tif" )
env.grid
env.grid.coarse= aggregate(env.grid,10)
```
Reproject climate.grid
```{r}
climate.grid.repr = project(climate.grid, env.grid.coarse,method="near")
values(climate.grid.repr) <- 1:ncell(climate.grid.repr)
writeRaster(climate.grid.repr,"D:/workspace/Winslow/iLand_AK/cpcrw/materials/env/climate.grid.tif",overwrite=T)

```
reproject the daylength grid


```{r}
#day.length.repr = project(day.length, env.grid.coarse,method="bilinear")
#plot(day.length.repr[[1]])
```

Okay now we convert the cropped climate.grid to a spatial object, and write it out.
```{r}
env.grid.sp= as.points(env.grid,values=T)
writeVector(env.grid.sp, "D:/workspace/Winslow/iLand_AK/cpcrw/materials/env/env_cells_extract.shp",overwrite=T)
climate.grid.sp=as.points(climate.grid.repr, value=T)
writeVector(climate.grid.sp, "D:/workspace/Winslow/iLand_AK/cpcrw/materials/env/climate_cells_extract.shp",overwrite=T)

```

Now produce a shapefile/dataframe that links the env grid to the climate gridcell number
```{r}
rasValue=as.data.frame(terra::extract(climate.grid.repr, env.grid.sp,df=T, bind=T))
rasValue = rasValue %>% dplyr::rename(env.gridCell= env.grid) %>% rename(climate.gridCell = tasmax_4)
write.table(rasValue, "D:/workspace/Winslow/iLand_AK/cpcrw/materials/climate/env.grid to climate.grid link.txt",row.names = F)
```

Read in historical temperature data starting with tmin
```{r}
ignore_leap_years <- function(year, day_of_year) {
  # Check if the year is a leap year
  if (leap_year(year) && day_of_year > 59) {
    # Add 1 to skip the leap day (February 29)
    day_of_year <- day_of_year + 1
  }
return(day_of_year)
  }
ncol <- function(x) dim(x)[2]

process.climate = function(model,ssp,var){
r=rast(paste0("Z:/project_data/downscaling/CPCRW/Downscaled ",model,"/",ssp,"/",var,"/",model,"-",ssp,"-",var,"-",year,".nc"))
r.repr=project(r,climate.grid.repr, method="bilinear")
if(var=="rsds"){
  r.repr = (r.repr * 86400)/1000000  #Convert radiation from watts/m2 to megajoules m2 day-1 (day length/1000000 joules per megajoule)
}else{
  r.repr=r.repr
}

t=as.data.frame(terra::extract(r.repr,climate.grid.sp,bind=T))
names(t)[1] <- "climate.gridCell"
n=ncol(t)
t_long <- tidyr::gather(t, y_day,value , 2:n)
if(var=="vp" &year<1980){
  t_long <- t_long %>% separate(y_day, c('vp', 'y_day'), sep = "\\.") # Had to add this given the name in vapor pressure changed
}else{
t_long <- t_long %>% separate(y_day, c('vp', 'y_day'), sep = '_')
}
t_long$vp=NULL
t_long = t_long %>% mutate(y_day= as.numeric(y_day)) 
t_long = t_long %>% mutate(day_of_year = mapply(ignore_leap_years,year,y_day))
t_long = t_long %>% mutate(date= make_date(year) + days(day_of_year-1))
t_long$y_day=NULL
t_long$day_of_year=NULL
t_long$value=round(t_long$value, digits=3)
write.table(t_long,paste0("D:/workspace/Winslow/iLand_AK/cpcrw/materials/climate/",model,"/",ssp,"/",model,"-",ssp,"-",var,"-",year,".txt"),row.names = F, sep = "\t")
}  


dir.create(paste0("D:/workspace/Winslow/iLand_AK/cpcrw/materials/climate/",gcm,"/ssp245"),recursive = T)

dir.create(paste0("D:/workspace/Winslow/iLand_AK/cpcrw/materials/climate/",gcm,"/ssp126"),recursive = T)

dir.create(paste0("D:/workspace/Winslow/iLand_AK/cpcrw/materials/climate/",gcm,"/ssp370"),recursive = T)


i="ssp126"
j="vp"
year=1950


for(i in levels(as.factor(c("ssp126","ssp245","ssp370")))){      
  for(j in levels(as.factor(c("hurs","pr","rsds","tasmax","tasmin","vp")))){          #
  for(year in 1950:2100){
  process.climate(gcm,i,j)
  print(paste0("Processed var ",j," and year ",year))
}
  }
}
```








