---
title: "climate_plots_writing_sqlite"
author: "Winslow Hansen"
date: "August 5, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

read in packages

```{r,echo=FALSE}
library(RSQLite)
library(tidyverse)
```

Read in and join historical data  

```{r}
path = "D:/workspace/Winslow/iLand_AK/cpcrw/materials/climate/"
gcm = "NorEsm2-MM"    #"NorEsm2-MM" "TaiESM1"
readin=function(model, ssp, var){
  data=data.frame()
for(year in 1950:2100){
t = read.table(paste0(path,model,"/",ssp,"/",model,"-",ssp,"-",var,"-",year,".txt"),header = T, sep = "\t")
t$date=as.Date(t$date)
names(t)[names(t)=="value"] <- var
data=rbind(data,t)
}
return(data)
}

for(i in levels(as.factor(c("ssp126","ssp245","ssp370")))){      
t_min=readin(gcm, i, "tasmin")
t_max=readin(gcm, i,"tasmax")
vap=readin(gcm, i,"vp")
srad=readin(gcm, i,"rsds")
pre=readin(gcm, i,"pr")
total=list(t_min, t_max, pre,srad, vap) %>% reduce(left_join, by = c("climate.gridCell","date")) #Update this variable name
assign(paste0("total_",i),total)
}
```


convert vap from hpa to Kpa by multiplying by 0.1
```{r}
total_ssp126=total_ssp126%>%mutate(vp=vp*0.1,
                     tasmax=tasmax - 273.15,
                     tasmin = tasmin - 273.15)

total_ssp126=total_ssp126%>%mutate(rsds=ifelse(rsds < 0, 0, rsds),
                     vp=ifelse(vp<0,0,vp))

total_ssp245=total_ssp245%>%mutate(vp=vp*0.1,
                     tasmax=tasmax - 273.15,
                     tasmin = tasmin - 273.15)

total_ssp245=total_ssp245%>%mutate(rsds=ifelse(rsds < 0, 0, rsds),
                     vp=ifelse(vp<0,0,vp))

total_ssp370=total_ssp370%>%mutate(vp=vp*0.1,
                     tasmax=tasmax - 273.15,
                     tasmin = tasmin - 273.15)

total_ssp370=total_ssp370%>%mutate(rsds=ifelse(rsds < 0, 0, rsds),
                     vp=ifelse(vp<0,0,vp))


pre=NULL
srad=NULL
t_max=NULL
t_min=NULL
test=NULL
vap=NULL
```


 Fucntions to rocess vapor pressure (which is already in KPA from previous script) into vapor pressure deficit 
```{r}
# ref for tetens formula based on temp
# Murray, F.W. 1967. On the computation of saturation vapour pressure. J. Applied Meteorology 6: 203-204. https://doi.org/10.1175/1520-0450(1967)006%3C0203:OTCOSV%3E2.0.CO;2
# also incorporated conversion from mb to kpa by dividing a in Murray's equation
# (6.1078) by 10

calc_es = function(temp) {
  es = ifelse(temp < 0,
              0.61078 * exp((21.875 * temp) / (temp + 265.5)),  
              0.61078 * exp((17.269 * temp) / (temp + 237.3)))  
  return(es)  
}  ##This is the function to calculation saturation pressure of water 

# to calc avg daily vpd, following FAO guidelines for meteorological data
# www.fao.org/docrep/X0490E/x0490e07.htm
calc_vpd = function(t_min, t_max, vap) {  # temp in C
  # first calc sat vapor pressure at min and max temp
  es_min = calc_es(t_min)
  es_max = calc_es(t_max)
  
  # average value
  es = (es_min + es_max) / 2
  
  # vpd is difference between sap vp and vp
  vpd = es - vap
  
  return(vpd)
}

```



Using these Functions
```{r}
# Calculate_vapd
# setting 0 as minimum value
total_ssp126$vpd_calc = calc_vpd(total_ssp126$tasmin, total_ssp126$tasmax, total_ssp126$vp)
total_ssp245$vpd_calc = calc_vpd(total_ssp245$tasmin, total_ssp245$tasmax, total_ssp245$vp)
total_ssp370$vpd_calc = calc_vpd(total_ssp370$tasmin, total_ssp370$tasmax, total_ssp370$vp)




total_ssp126$vpd = ifelse(total_ssp126$vpd_calc > 0, total_ssp126$vpd_calc, 0)
total_ssp126$vpd_cal=NULL

total_ssp245$vpd = ifelse(total_ssp245$vpd_calc > 0, total_ssp245$vpd_calc, 0)
total_ssp245$vpd_cal=NULL

total_ssp370$vpd = ifelse(total_ssp370$vpd_calc > 0, total_ssp370$vpd_calc, 0)
total_ssp370$vpd_cal=NULL


```
generate a model climate name that is not a number
```{r}

total_ssp126=total_ssp126%>%mutate(add="cpcrw")%>% unite("model.climate.tableName", c(add,climate.gridCell), remove = T)
total_ssp245=total_ssp245%>%mutate(add="cpcrw")%>% unite("model.climate.tableName", c(add,climate.gridCell), remove = T)
total_ssp370=total_ssp370%>%mutate(add="cpcrw")%>% unite("model.climate.tableName", c(add,climate.gridCell), remove = T)

```

round all variables to two decimal places
```{r}
total_ssp126=total_ssp126%>%mutate(rsds=round(rsds, digits=2),
                     vpd=round(vpd,digits=2))

total_ssp245=total_ssp245%>%mutate(rsds=round(rsds, digits=2),
                     vpd=round(vpd,digits=2))
total_ssp370=total_ssp370%>%mutate(rsds=round(rsds, digits=2),
                     vpd=round(vpd,digits=2))

```
Write out the data to an sqlite for the 1951-2005 period

```{r}

for(i in levels(as.factor(c("ssp126","ssp245","ssp370")))){      
total=get(paste0("total_",i))
data_historical=total%>%mutate(year= lubridate::year(date),
                   month= lubridate::month(date),
                   day= lubridate::day(date)) #add year month d
getwd()
db.conn <<- dbConnect(RSQLite::SQLite(), dbname=paste0(path,gcm,i,"V3.sqlite")) ###ADjust name based on project
 # order is id, year, month, day, min_temp, max_temp, prec, rad, vpd
data_historical$model.climate.tableName=as.factor(data_historical$model.climate.tableName)


for(j in levels(data_historical$model.climate.tableName)){
 
  dat=data_historical[data_historical$model.climate.tableName==j,c(10,11,12,2,4,5,6,9)]
  
  names(dat)[4]="min_temp"
  names(dat)[5]="max_temp"
  names(dat)[6]="prec"
  names(dat)[7]="rad"
  dbWriteTable(db.conn, name=paste(j),dat,row.names=F,overwrite=T)
}
dbDisconnect(db.conn)
}
#write.table(subset_2030,"historicalclimate.txt",sep="\t")
```

Function to plot summer and annual values
```{r}

db.conn <- DBI::dbConnect(RSQLite::SQLite(), dbname="D:/workspace/Winslow/iLand_AK/cpcrw/model/databases/cpcrw_climate_hist_ccsm4.sqlite")
tables <- dbListTables(db.conn)

all_data <- lapply(tables, function(table) {
  # Read the table
  df <- dbReadTable(db.conn, table)
  
  # Optionally, add a column to indicate the table name
  df$table_name <- table
  
  return(df)
}) %>% bind_rows()

dbDisconnect(db.conn)

all_data$model="cmip5" 
all_data = all_data %>% filter(year>1979)


  db.conn <- DBI::dbConnect(RSQLite::SQLite(), dbname="D:/workspace/Winslow/iLand_AK/cpcrw/materials/climate/NorEsm2-MMssp245V2.sqlite")
tables <- dbListTables(db.conn)

new_data <- lapply(tables, function(table) {
  # Read the table
  df <- dbReadTable(db.conn, table)
  
  # Optionally, add a column to indicate the table name
  df$table_name <- table
  
  return(df)
}) %>% bind_rows()

dbDisconnect(db.conn)
new_data$model="cmip6"
new_data=new_data %>% filter(year<2006)
summary(new_data)
summary(all_data)
all_data = rbind(all_data,new_data)


total.summary.an=all_data%>%group_by(year,model)%>%summarise_if(is.numeric, funs(max,sd))
total.summary.sum=all_data%>%filter(month<9&month>4)%>%group_by(year,model)%>%summarise_if(is.numeric, funs(mean,sd))



ggplot(data = total.summary.sum)+
    geom_ribbon(aes(x=year, ymin= max_temp_mean - max_temp_sd,ymax =max_temp_mean + max_temp_sd,fill=model),alpha=0.5)+
  geom_line(aes(x=year, y = max_temp_mean, color=model))+
        theme_bw()+
       labs(title = "mean max summer temp",x="Year")

ggplot(data = total.summary.sum)+
    geom_ribbon(aes(x=year, ymin= min_temp_mean - min_temp_sd,ymax =min_temp_mean + min_temp_sd,fill=model),alpha=0.5)+
  geom_line(aes(x=year, y = min_temp_mean, color=model))+
        theme_bw()+
       labs(title = "mean min summer temp",x="Year")

ggplot(data = total.summary.sum)+
    geom_ribbon(aes(x=year, ymin= vpd_mean - vpd_sd,ymax =vpd_mean + vpd_sd,fill=model),alpha=0.5)+
  geom_line(aes(x=year, y = vpd_mean, color=model))+
        theme_bw()+
       labs(title = "mean summer vpd",x="Year")

ggplot(data = total.summary.sum)+
    geom_ribbon(aes(x=year, ymin= rad_mean - rad_sd,ymax =rad_mean + rad_sd,fill=model),alpha=0.5)+
  geom_line(aes(x=year, y = rad_mean, color=model))+
        theme_bw()+
       labs(title = "mean summer rad",x="Year")

ggplot(data = total.summary.sum)+
    geom_ribbon(aes(x=year, ymin= prec_mean - prec_sd,ymax =prec_mean + prec_sd,fill=model),alpha=0.5)+
  geom_line(aes(x=year, y = prec_mean, color=model))+
        theme_bw()+
       labs(title = "mean summer prec",x="Year")
```
