---
title: "iLand landscape building"
author: "Winslow Hansen"
date: "May 26, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

libraries
```{r}
library(tidyverse)
library(raster)
raster::rasterOptions(maxmemory = 1e+10)
library(sf)
raster::rasterOptions()
```
Read in the clipping region that delineates the landscape in Alaska
```{r}
#update your path here
clip = raster("raw data/master_gridf.tif")
clip
#clip=spatial.tools::modify_raster_margins(clip,extent_delta = c(0, 7, 6, 0),1)
#writeRaster(clip,"E:/post doc 2018-2020/Alaska fire modeling/the model/iLand landscapes/landscape building/raw data/master_gridf.tif",overwrite=T)  # this was done to square up the 10x10m grid to deal with converting to 100m RUs
plot(clip)

```
the first thing I'm going to do is build forest cover layer using the Hansen et al. 2000 forest cover and the hansen et al. 2000-2019 forest loss datasets. Note that for simplicity I reprojected them to my native crs in arc.
Read in the datasets
```{r}
forest.2000 = raster("E:/post doc 2018-2020/Alaska fire modeling/the model/iLand landscapes/landscape building/raw data/forest_data/Hansen_GFC-2019-v1.7_treecover2000_70N_150W_repr.tif")

forest.loss.2019 = raster("E:/post doc 2018-2020/Alaska fire modeling/the model/iLand landscapes/landscape building/raw data/forest_data/Hansen_GFC-2019-v1.7_lossyear_70N_150W_repr.tif")
forest.2000
crs(forest.2000)
crs(forest.loss.2019)
crs(clip)
#plot(forest.2000)
above=stack("raw data/forest_data/ABoVE_LandCover_Bh06v05_repr.tif")
above
above.2000=above[[17]]
above.2000
```
 
  
Now crop the forest cover layers to the clip grid and project it back to the clip grid
```{r}
forest.2000.masked = crop(forest.2000, clip) 
forest.loss.2019.masked = crop(forest.loss.2019, clip) 
above.2000.masked = crop(above.2000, clip)
above.2000.masked
clip
plot(forest.2000.masked)
plot(forest.loss.2019.masked)
plot(above.2000.masked)
```
Looking at forest loss since 2000
```{r}
forest.2019.masked <- overlay(x=forest.2000.masked, y=forest.loss.2019.masked, fun=function(x, y) ifelse(y>0, 0, x)) 

plot(forest.2019.masked)
plot(forest.2000.masked)
```

The goal will be to compare forest cover in the Hansen dataset to the ABOVE dataset at this small scale.
Be sure to compare 2000 to 2000
First reclassify the data For Hansen, anything greater than or equal to 60% canopy cover is forest. This matches the ABOVE cutoffs
For the ABOVE we group the forest vegetation types together into forest.
```{r}

values(forest.2000.masked)[values(forest.2000.masked) <=30] = 0
 values(forest.2000.masked)[values(forest.2000.masked) >30] = 1
 above.2000.masked.forest=above.2000.masked
 above.2000.masked.forest=setExtent(above.2000.masked.forest,clip)

 values(above.2000.masked)[values(above.2000.masked) <=4] = 1
 values(above.2000.masked)[values(above.2000.masked) >4] = 0
 

 
plot(forest.2000.masked)
plot(above.2000.masked)
```
```{r}

above.2000.masked
forest.2000.masked
forest.2000.masked2 <- setExtent(forest.2000.masked, extent(above.2000.masked), keepres=TRUE)
test=stack(above.2000.masked,forest.2000.masked2)



test=stack(above.2000.masked,forest.2000.masked2)
above.2000.masked.df=as.data.frame(above.2000.masked,xy=T)

forest.2000.masked.df=as.data.frame(forest.2000.masked2,xy=T)
total=left_join(above.2000.masked.df,forest.2000.masked.df)
head(total)
total=total%>%rename(Above = ABoVE_LandCover_Bh06v05_repr.17)%>%rename(Hansen = Hansen_GFC.2019.v1.7_treecover2000_70N_150W_repr)

summary(total)
total= total%>% mutate(Above = ifelse(is.na(Above),0,Above))
total=total%>%mutate(common= ifelse(Hansen==1&Above==1,1,0),
                     above.sum =ifelse(Hansen==0&Above==1,1,0),
                     hansen.sum=ifelse(Hansen==1&Above==0,1,0))
summary = total%>%summarize(count.common=sum(common),
                            count.above=sum(above.sum),
                            count.hansen=sum(hansen.sum),
                            all=length(common))

4566604/6062512
291354/6062512
692358/6062512
```

quick plot
```{r}
cpcrw <- st_read(
  "E:/post doc 2018-2020/Alaska fire modeling/Alaska_rawdata/data/LTER_boundaries/436_cpcrw_bound_nad83/cpcrw_bound_nad83.shp")

hansen.p=ggplot()+
  geom_raster(data=total,aes(x=x,y=y,fill=as.factor(Hansen)))+
  scale_fill_viridis_d()+
  geom_sf(data=cpcrw,fill=NA,color="red")+
  theme_bw()+
  labs(fill="Forest")
ggsave("E:/post doc 2018-2020/Alaska fire modeling/the model/iLand landscapes/landscape building/raw data/forest_data/hansen.map.jpg",units="in",height = 5,width = 8.5)
above.p=ggplot()+
  geom_raster(data=total,aes(x=x,y=y,fill=as.factor(Above)))+
  scale_fill_viridis_d()+
  geom_sf(data=cpcrw,fill=NA,color="red")+
  theme_bw()+
  labs(fill="Forest")
ggsave("E:/post doc 2018-2020/Alaska fire modeling/the model/iLand landscapes/landscape building/raw data/forest_data/ABoVE.map.jpg",units="in",height = 5,width = 8.5)



```

plot the actual forest types from the above dataset
```{r}
above.df=as.data.frame(above.2000.masked.forest,xy=T)
summary(above.df)
above.df=above.df%>%rename(landcover=ABoVE_LandCover_Bh06v05_repr.17)

above.df=above.df%>%mutate(vegetation.type = ifelse(landcover==1,"Evergreen forest",
                                                    ifelse(landcover==2,"Deciduous forest",
                                                           ifelse(landcover==3, "Mixed forest",
                                                                  ifelse(landcover==4, "Woodland","Non forest")))))

above.forest.p=ggplot()+
  geom_raster(data=above.df,aes(x=x,y=y,fill=as.factor(vegetation.type)))+
  scale_fill_viridis_d()+
  geom_sf(data=cpcrw,fill=NA,color="red")+
  theme_bw()+
  labs(fill="Forest type")
ggsave("E:/post doc 2018-2020/Alaska fire modeling/the model/iLand landscapes/landscape building/raw data/forest_data/ABoVE.forest.map.jpg",units="in",height = 5,width = 8.5)
```

REad in a dem, aspect, and permafrost to start differentiating locations 
```{r}
dem=raster("E:/post doc 2018-2020/Alaska fire modeling/the model/iLand landscapes/landscape building/raw data/DEM/dem_repr.tif")
aspect=raster("E:/post doc 2018-2020/Alaska fire modeling/the model/iLand landscapes/landscape building/raw data/DEM/aspect_repr.tif")
hillshade=raster("E:/post doc 2018-2020/Alaska fire modeling/the model/iLand landscapes/landscape building/raw data/DEM/hillshade_repr.tif")
permafrost=raster("E:/post doc 2018-2020/Alaska fire modeling/the model/iLand landscapes/landscape building/raw data/permafrost/permafrost_repr_clip_res.tif")
 plot(dem)
plot(aspect)
plot(permafrost)
aspect
permafrost
clip
above.2000.masked.forest
```

Clip and fit extent to forest data
```{r}

dem.masked = crop(dem, clip) 
#values(dem.masked)[values(dem.masked) > -2000] = 1
#writeRaster(dem.masked, "E:/post doc 2018-2020/Alaska fire modeling/the model/iLand landscapes/landscape building/raw data/master_gridf2.tif")
aspect.masked = crop(aspect, clip)
hillshade.masked = crop(hillshade, clip)
dem.masked2=setExtent(dem.masked, extent(clip), keepres=TRUE)
permafrost.masked2=setExtent(permafrost, extent(clip), keepres=TRUE)
aspect.masked2=setExtent(aspect.masked, extent(clip), keepres=TRUE)
hillshade.masked2=setExtent(hillshade.masked, extent(clip), keepres=TRUE)

stack(above.2000.masked.forest,permafrost.masked2,aspect.masked2,dem.masked2)
clip
writeRaster(above.2000.masked.forest,"F:/work/Madison/processed_data/veg.processed.tif")
writeRaster(permafrost.masked2,"F:/work/Madison/processed_data/permafrost.processed.tif")
writeRaster(dem.masked2,"F:/work/Madison/processed_data/dem.processed.tif")
writeRaster(aspect.masked2,"F:/work/Madison/processed_data/aspect.processed.tif")

```
Turn into data.frames
```{r}
dem.df = as.data.frame(dem.masked2,xy=T)
aspect.df = as.data.frame(aspect.masked2, xy=T)
hillshade.df = as.data.frame(hillshade.masked2,xy=T)
permafrost.df = as.data.frame(permafrost.masked2, xy=T)
forest.df= as.data.frame(above.2000.masked.forest,xy=T)

total=list(forest.df, dem.df, aspect.df,hillshade.df, permafrost.df) %>% reduce(left_join, by = c("x","y"))
head(total)
total=total%>%rename(Above = ABoVE_LandCover_Bh06v05_repr.17)%>%rename(elevation = dem_repr)%>%rename(aspect = aspect_repr)%>%rename(hillshade = hillshade_repr)%>%rename(permafrost = permafrost_repr_clip_res)
```
Classify aspect into categories and then do the forest types into categories
```{r}

total=total%>%mutate(aspect.dir=ifelse(aspect == -1, "flat",
                                       ifelse(aspect>-1&aspect<=45,"N",
                                              ifelse(aspect>45&aspect<=135,"E",
                                                     ifelse(aspect>135&aspect<=225,"S",
                                                           ifelse(aspect>225&aspect<=315,"W","N"))))))

total=total%>%mutate(forest.type = ifelse(Above==1,"Evergreen",
                                          ifelse(Above==2,"Deciduous",
                                                 ifelse(Above==3,"Mixed",
                                                        ifelse(Above==4,"Woodland",
                                                               ifelse(Above==5|Above==6|Above==7|Above==8|Above==9,"Potential-forest", "Non-forest"))))))
total=total%>%mutate(permafrost = ifelse(is.na(permafrost),0,permafrost))

```

Now lets generate species level rules
```{r}
summary(total)
plot.diag=total%>%filter(is.na(permafrost))
ggplot()+
   geom_raster(data=plot.diag,aes(x=x,y=y,fill=elevation))+
  scale_fill_viridis_c()+
  #geom_sf(data=cpcrw,fill=NA,color="red")+
  theme_bw()+
  labs(fill="Forest type")

total=total%>%mutate(species = ifelse(Above==1 & aspect.dir=="N" ,"Black.spruce",NA))
total=total%>%mutate(species = ifelse(Above==1 & aspect.dir!="N" ,"White.spruce",species))
total=total%>%mutate(species = ifelse(Above==1 & aspect.dir=="flat" & permafrost==1 ,"Black.spruce",species))
total=total%>%mutate(species = ifelse(Above==1 & aspect.dir=="flat" & permafrost==0 ,"White.spruce",species))
total=total%>%mutate(species = ifelse(Above==4 ,"Black.spruce",species))
total=total%>%mutate(species = ifelse(Above==2 & aspect.dir=="S" ,"Aspen",species))
total=total%>%mutate(species = ifelse(Above==2 & aspect.dir!="S" ,"Birch",species))
total=total%>%mutate(species = ifelse(Above==3, "Mixed",species))
total=total%>%mutate(species = ifelse(Above==5|Above==6|Above==7|Above==8|Above==9,"Potential-forest",species))
total=total%>%filter(forest.type!="Non-forest")
total=total%>%mutate(species = as.factor(species))%>%mutate(aspect.dir=as.factor(aspect.dir))
total%>%filter(is.na(species))
summary(total)
library(viridis)
ggplot()+
  geom_raster(data = total, aes(x = x, 
                                  y = y, 
                                  alpha = hillshade)) +
    scale_alpha(name = "", range = c(1, 0.2), guide = F,na.value="grey")  + 
 geom_raster(data=total, aes(x=x,y=y,fill=species))+
  scale_fill_viridis_d()+
  geom_sf(data=cpcrw,fill=NA,color="red")+
  theme_bw()+
  labs(fill="Forest_species")
total.summary=total%>%group_by(species)%>%summarise(count=length(species))
ggplot(data=total, aes(x=species))+
  geom_bar(aes(y = (..count..)/sum(..count..)))+
  coord_flip()+
  theme_bw()+
  labs(x="Species",y="Forested area (%)")
ggsave("E:/post doc 2018-2020/Alaska fire modeling/the model/iLand landscapes/landscape building/cpcrw/percentages.jpg")
```



Plotting species as an xyz and then writing species and the DEM out to a folder
```{r}
total=total%>%mutate(species2=ifelse(species =="Black.spruce",1,
                     ifelse(species=="White.spruce",2,
                            ifelse(species=="Aspen",3,
                                   ifelse(species=="Birch",4,
                                          ifelse(species=="Mixed",5,
                                          ifelse(species=="Potential-forest",6,7)))))))

total.subset=total%>%dplyr::select(x,y,species2)
forest.species = rasterFromXYZ(total.subset,crs = crs(clip))
forest.species
stack(forest.species,permafrost.masked2)
writeRaster(forest.species,"E:/post doc 2018-2020/Alaska fire modeling/the model/iLand landscapes/landscape building/cpcrw/materials/init/forest_species_init.tif",overwrite=T)
write.table(as.matrix(dem.masked2),"E:/post doc 2018-2020/Alaska fire modeling/the model/iLand landscapes/landscape building/cpcrw/materials/env/DEM_processed.txt" ,col.names = F,row.names = F)
```
