---
title: "generating environmental file"
author: "Winslow Hansen"
date: "9/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read in environment grid to get the id names
```{r}
library(terra)
library(tidyverse)
env.grid=rast("D:/workspace/Winslow/cpcrw/permafrost_param/iLand_landscapes/landscape_building/cpcrw/materials/env/env.grid.tif")
env.grid
```
Read in the spatial points data frame of the environment grid so we can use it for extraction
```{r}
env.grid.sp = vect("D:/workspace/Winslow/cpcrw/permafrost_param/iLand_landscapes/landscape_building/cpcrw/materials/env/env_cells_extract.shp")

env.grid.sp
```

REad in the climate to enviornment grid that I've created in climate_processing_step1
```{r}

climate.link=read.table("D:/workspace/Winslow/cpcrw/permafrost_param/iLand_landscapes/landscape_building/cpcrw/materials/climate/env.grid to climate.grid link.txt",header=T)
climate.link = climate.link %>% rename(env.grid = env.gridCell)
# Perform the left join using merge()
# Join by the 'ID' column, retaining all rows from the SpatVector (left join)
env.grid.sp <- merge(env.grid.sp, climate.link, by = "env.grid", all.x = TRUE)
```

reading in other variables in the environment file
```{r}
soil.depth=rast("D:/workspace/Winslow/cpcrw/permafrost_param/iLand_landscapes/landscape_building/cpcrw/materials/env/depth.processed.tif")
sand=rast("D:/workspace/Winslow/cpcrw/permafrost_param/iLand_landscapes/landscape_building/cpcrw/materials/env/sand.processed.tif")
silt=rast("D:/workspace/Winslow/cpcrw/permafrost_param/iLand_landscapes/landscape_building/cpcrw/materials/env/silt.processed.tif")
clay=rast("D:/workspace/Winslow/cpcrw/permafrost_param/iLand_landscapes/landscape_building/cpcrw/materials/env/clay.processed.tif")

```
Reading in the species raster to set the carbon and nitrogen pools for initial conditions and aggregate using max to 100m grid cell size.
```{r}
species=rast("D:/workspace/Winslow/cpcrw/permafrost_param/iLand_landscapes/landscape_building/cpcrw/materials/init/forest_species_init.tif")

```
Trying extraction
```{r}
depth.table=terra::extract(soil.depth,env.grid.sp,df=T)
sand.table=terra::extract(sand,env.grid.sp,df=T)
silt.table=terra::extract(silt,env.grid.sp,df=T)
clay.table=terra::extract(clay,env.grid.sp,df=T)
species.table=terra::extract(species, env.grid.sp, df=T)
summary(species.table)
```
Now join them together.
```{r}
total=list(depth.table, sand.table, silt.table, clay.table, species.table) %>% reduce(left_join, by = "ID")
summary(total)
```
Turning species in Carbon and nitrogen pools
```{r}
total = total %>% mutate(model.site.youngLabileC = ifelse(forest_species_init ==1, runif(70000,48682,96901) ,
                                                   ifelse(forest_species_init == 2,  runif(70000,48682,96901), 
                                                    ifelse(forest_species_init ==3,  runif(70000,17371,31765), 
                                                    ifelse(forest_species_init ==4,  runif(70000,17371,31765),
                                                    ifelse(forest_species_init == 5, runif(70000,33026.5,64336),runif(70000,48682,96901)))))),
                         model.site.youngLabileN = ifelse(forest_species_init ==1, model.site.youngLabileC/32.62,
                                                   ifelse(forest_species_init == 2,  model.site.youngLabileC/32.62 ,
                                                    ifelse(forest_species_init ==3,  model.site.youngLabileC/23.58 ,
                                                    ifelse(forest_species_init ==4, model.site.youngLabileC/23.58 ,
                                                    ifelse(forest_species_init ==5, model.site.youngLabileC/28.01 , model.site.youngLabileC/32.62))))),
                         model.site.youngRefractoryC = ifelse(forest_species_init ==1,  17000,
                                                   ifelse(forest_species_init == 2,  17000,
                                                    ifelse(forest_species_init ==3,  20020,
                                                    ifelse(forest_species_init ==4, 20020, 
                                                    ifelse(forest_species_init ==5, 18500, 17000))))),
                         model.site.youngRefractoryN = ifelse(forest_species_init ==1, model.site.youngRefractoryC/425 ,
                                                   ifelse(forest_species_init == 2, model.site.youngRefractoryC/425 ,
                                                    ifelse(forest_species_init ==3,  model.site.youngRefractoryC/417.1 ,
                                                    ifelse(forest_species_init ==4,  model.site.youngRefractoryC/417.1 ,
                                                    ifelse(forest_species_init ==5,  model.site.youngRefractoryC/421.05 ,425))))),
                         model.site.somC = 35000,
                         model.site.somN = model.site.somC/20,
                         model.settings.permafrost.moss.biomass = ifelse(forest_species_init ==1, runif(70000,1.55,2.79) ,
                                                   ifelse(forest_species_init == 2,  runif(70000,0.31,0.93), 
                                                    ifelse(forest_species_init ==3,  runif(70000,0,0.0001), 
                                                    ifelse(forest_species_init ==4,  runif(70000,0,0.0001),
                                                    ifelse(forest_species_init == 5, runif(70000,0.62,1.55),runif(70000,1.55,2.79)))))))
                         
summary(total)
```

Rename the climate grid so it doesn't start with a number 

```{r}
env.grid.df=as.data.frame(env.grid.sp)
env.grid.df=env.grid.df%>%mutate(add="cpcrw") %>% unite("model.climate.tableName", c(add,climate.gridCell
), remove = T)

total= total %>% rename(env.grid = ID)
total = left_join(env.grid.df, total, by = "env.grid") 
```


Now write out the environment file
```{r}
total[is.na(total)]=0
env.file=data.frame(id=total$env.grid,
                    model.climate.tableName = total$model.climate.tableName,
                    model.site.availableNitrogen = 45,
                    model.site.soilDepth = round(total$depth.processed,digits = 0),
                    model.site.pctSand = round(total$sand.processed, digits = 0),
                    model.site.pctSilt = round(total$silt.processed, digits = 0),
                    model.site.pctClay = round(total$clay.processed, digits = 0),
                    model.site.youngLabileC = round(total$model.site.youngLabileC, digits = 0),
                    model.site.youngLabileN = round(total$model.site.youngLabileN, digits = 0),
                    model.site.youngRefractoryC = round(total$model.site.youngRefractoryC, digits = 0),
                    model.site.youngRefractoryN = round(total$model.site.youngRefractoryN, digits = 0),
                    model.site.somC = round(total$model.site.somC, digits = 0),
                    model.site.somN = round(total$model.site.somN, digits = 0),
                    model.settings.permafrost.moss.biomass = round(total$model.settings.permafrost.moss.biomass, digits = 4)
                    )
summary(env.file)
write.table(env.file,"D:/workspace/Winslow/cpcrw/permafrost_param/iLand_landscapes/landscape_building/cpcrw/materials/env/env.file_cmip6.txt",row.names = F,sep="\t")

t=read.table("D:/workspace/Winslow/cpcrw/permafrost_param/iLand_landscapes/landscape_building/cpcrw/materials/env/env.file.txt",header=T)
summary(t)
summary(env.file)
summary(total)
```






