---
title: "Area Dominated"
author: "FFL"
date: today
---

# Data {.unnumbered .unlisted .hidden}

```{r}
#| label: data processing
source("./scripts/install.R")

out_dir  <- here("data", "estSOLthickness", "area_dom_processed")
out_file <- here("data", "estSOLthickness", "area_dom_processed", "area_dom_sum.rds")
in_dir   <- "//10.60.2.10/FF_Lab/project_data/na_boreal/sensitivity_analysis/data/output/estSOLthickness/areaDom"
if(!file.exists(in_dir)) {
  stop("Are you connected to Janet?")
}

dir.create(out_dir, recursive = TRUE)

if (!file.exists(out_file)) {
  area_dom_files <- list.files(in_dir, full.names = TRUE, pattern = "*.txt")

  # plan(sequential)
  plan(multisession, workers = 10)
  area_dom_landscape_year <- furrr::future_map_dfr(area_dom_files, \(x) {
    read_delim(x, show_col_types = FALSE) |>
      group_by(treatment, replicate, year, sp.dom) |>
      summarise(area = sum(area_ha), .groups = "drop")
  })
  plan(sequential)

  saveRDS(area_dom_landscape_year, out_file)

} else {
  area_dom_landscape_year <- readRDS(out_file)
}

```

# Figures

Kristin Braziunas:

> My rule of thumb is that epsilon is sensitive in increments of ~0.1. Most iLand sims I'm familiar with these days are using ranges from 2.6-3.0
> I recently ran my own sensitivity experiment where I varied epsilon by 0.1 in this range. I also expanded the range (but used wider increments of 0.4 or 0.5) to 1.8-3.5. 
> I got the range to test from: 
> - Landsberg & Waring, who developed the radiation use efficiency model that was used as the basis (https://doi.org/10.1016/j.foreco.2013.10.035) - they use 1.8. 
- Makela et al. 2007 (https://onlinelibrary.wiley.com/doi/full/10.1111/j.1365-2486.2007.01463.x): 1.9-3.1
- Seidl et al. 2012 (orig iLand paper) gives 1.8 (Europe) and 2.6 (US PNW)

```{r}
#| label: line plots
#| echo: false
#| fig-height: 7

colours <- c(
  "Pima" = "#8c564b", "Pigl" = "#1f77b4",
  "Potr" = "#2ca02c", "Bene" = "green",
  "Mixed.spruce" = "#9467bd", "Mixed.deciduous" = "#bcbd22")

area_dom_landscape_year <- area_dom_landscape_year |>
  filter(sp.dom != "Not forested") |>
  mutate(
    rep_sp  = paste(sp.dom, replicate, sep = "_"),
    est     = as.numeric(str_extract(treatment, "(?<=(?:basefecund_|maxfecund_))[0-9.]+")),
    est     = replace_na(est, 0.5),
    ep_part = replace_na(str_extract(treatment, "ep[0-9.]+"), "ep2.7"),
    fec_type = if_else(str_detect(treatment, "maxfecund"), "max_est", "base_est"),
    treatment = paste0(
      substr(treatment, 1, 6),
      "_", fec_type, est,
      "_", ep_part,
      "_dbh0.1"
    ),
    treatment = factor(treatment, levels = unique(treatment[order(est)]))
  )
# dplyr::filter(x, grepl("max", treatment))
# dplyr::filter(x, !grepl("max", treatment))

```


```{r}
#| echo: false
plot_list_area_dom <- area_dom_landscape_year |>
  split(area_dom_landscape_year$treatment) |>
  lapply(\(x){
    ggplot(x, aes(x = year, y = area, group = rep_sp, colour = sp.dom)) +
      geom_line() +
      scale_colour_manual(values = colours) +
      guides(colour = guide_legend(override.aes = list(linewidth = 3))) +
      ylim(c(0, 55000)) +
      facet_wrap(~ treatment) +
      theme_minimal() +
      theme(
        legend.position = "bottom",
        legend.key.width = unit(2, "cm"))
  }) 

headings <- names(plot_list_area_dom)
```


```{r}
#| echo: false
plot_list_area_dom2 <- area_dom_landscape_year |>
  filter(
    sp.dom %in% c("Bene", "Mixed.deciduous", "Potr"),
    str_detect(treatment, regex("NorEsm"))
  ) |>
  droplevels() |>
  (\(df) split(df, df$treatment))() |>
  lapply(\(x) {
    ggplot(x, aes(x = year, y = area, group = rep_sp, colour = sp.dom)) +
      geom_line() +
      scale_colour_manual(values = colours) +
      guides(colour = guide_legend(override.aes = list(linewidth = 3))) +
      ylim(c(0, 7000)) +
      facet_wrap(~ treatment) +
      theme_minimal() +
      theme(
        legend.position = "bottom"
        )
  })

headings2 <- names(plot_list_area_dom2)
```









## NorEsm new epsilon + est results

New results estSOLthickness = 0.5 and 0.3 and epsilon 1.8 and 2.2

:::{.panel-tabset}
```{r}
#| warning: false
#| results: asis
#| fig-height: 7

epsilon_plots <- headings[grep(x = headings, pattern = "^NorEsm.*est0\\.5|^NorEsm.*est0\\.3")]
epsilon_plot_list <- plot_list_area_dom[names(plot_list_area_dom) %in% epsilon_plots]

for (i in seq_along(epsilon_plot_list)) {
  cat("# ",epsilon_plots[i],"\n")
  print(epsilon_plot_list[[i]])
  cat('\n\n')
}
```
:::


Deciduous only:


:::{.panel-tabset}
```{r}
#| warning: false
#| results: asis
#| fig-height: 7

epsilon_plots2 <- headings[grep(x = headings, pattern = "^NorEsm.*est0\\.5|^NorEsm.*est0\\.3")]
epsilon_plot_list2 <- plot_list_area_dom2[names(plot_list_area_dom2) %in% epsilon_plots2]

for (i in seq_along(epsilon_plot_list2)) {
  cat("# ",epsilon_plots2[i],"\n")
  print(epsilon_plot_list2[[i]])
  cat('\n\n')
}
```
:::









## NorEsm epsilon

Here are figures for NorEsm looking at default estSOLthickness = 0.5 and varying epsilon (default2.7)

:::{.panel-tabset}
```{r}
#| warning: false
#| results: asis
#| fig-height: 7

epsilon_plots <- headings[grep(x = headings, pattern = "^NorEsm.*est0\\.5")]
epsilon_plot_list <- plot_list_area_dom[names(plot_list_area_dom) %in% epsilon_plots]

for (i in seq_along(epsilon_plot_list)) {
  cat("# ",epsilon_plots[i],"\n")
  print(epsilon_plot_list[[i]])
  cat('\n\n')
}
```
:::




## NorEsm epsilon decid

Here are figures for NorEsm looking at default estSOLthickness = 0.5 and varying epsilon (default2.7)

:::{.panel-tabset}
```{r}
#| warning: false
#| results: asis
#| fig-height: 7

epsilon_plots2 <- headings[grep(x = headings, pattern = "^NorEsm.*est0\\.5")]
epsilon_plot_list2 <- plot_list_area_dom2[names(plot_list_area_dom2) %in% epsilon_plots2]

for (i in seq_along(epsilon_plot_list2)) {
  cat("# ",epsilon_plots2[i],"\n")
  print(epsilon_plot_list2[[i]])
  cat('\n\n')
}
```
:::



## NorEsm estSOLthickness decid

These figures vary estSOLthickness and keep epsilon at default 2.7:

:::{.panel-tabset}
```{r}
#| warning: false
#| results: asis
#| fig-height: 7

est_plots2 <- headings[grep(x = headings, pattern = "^NorEsm.*est[0-9.]+.*ep2\\.7")]
est_plot_list2 <- plot_list_area_dom2[names(plot_list_area_dom2) %in% est_plots2]

for (i in seq_along(est_plot_list2)) {
  cat("# ",est_plots2[i],"\n")
  print(est_plot_list2[[i]])
  cat('\n\n')
}
```
:::


## NorEsm estSOLthickness

These figures vary estSOLthickness and keep epsilon at default 2.7:

:::{.panel-tabset}
```{r}
#| warning: false
#| results: asis
#| fig-height: 7

est_plots <- headings[grep(x = headings, pattern = "^NorEsm.*est[0-9.]+.*ep2\\.7")]
est_plot_list <- plot_list_area_dom[names(plot_list_area_dom) %in% est_plots]

for (i in seq_along(est_plot_list)) {
  cat("# ",est_plots[i],"\n")
  print(est_plot_list[[i]])
  cat('\n\n')
}
```
:::




## NorEsm estSOLthickness + epsilon

All results so far:


:::{.panel-tabset}
```{r}
#| warning: false
#| results: asis
#| fig-height: 7

est_ep_plots <- headings[grep(x = headings, pattern = "NorEsm")]
est_ep_plot_list <- plot_list_area_dom[names(plot_list_area_dom) %in% est_ep_plots]

for (i in seq_along(est_ep_plot_list)) {
  cat("# ",est_ep_plots[i],"\n")
  print(est_ep_plot_list[[i]])
  cat('\n\n')
}
```
:::



## NorEsm vs TaiEsm estSOLthickness

Epsilon fixed at default 2.7

```{r}
#| warning: false
#| echo: false
#| fig-height: 12
#| fig-width: 8

nor_tai_plots <- headings[grep(x = headings, pattern = "^(NorEsm|TaiESM).*est[0-9.]+.*ep2\\.7")]

area_dom_nor_tai <- area_dom_landscape_year |>
  filter(treatment %in% nor_tai_plots) |>
  mutate(gcm = substr(treatment, 1, 6),
         est = str_extract(treatment, "est[0-9.]+"))

ggplot(area_dom_nor_tai, aes(x = year, y = area, group = rep_sp, colour = sp.dom)) +
  geom_line() +
  scale_colour_manual(values = colours) +
  facet_grid(est ~ gcm) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## NorEsm vs TaiEsm epsilon {.unlisted .hidden}

estSOLthick fixed at default 0.5

```{r}
#| warning: false
#| echo: false
#| fig-height: 12
#| fig-width: 8

nor_tai_plots_ep <- headings[grep(x = headings, pattern = "^(NorEsm|TaiESM).*est0\\.5")]

area_dom_nor_tai_ep <- area_dom_landscape_year |>
  filter(treatment %in% nor_tai_plots_ep) |>
  mutate(gcm = substr(treatment, 1, 6),
         ep = str_extract(treatment, "ep[0-9.]+"))

ggplot(area_dom_nor_tai_ep, aes(x = year, y = area, group = rep_sp, colour = sp.dom)) +
  geom_line() +
  scale_colour_manual(values = colours) +
  facet_grid(ep ~ gcm) +
  theme_minimal() +
  theme(legend.position = "bottom")
```
