---
title: "Debug summaries"
author:
  name: "Quinn Asena"
date: today
format:
  html:
    code-fold: false
    toc: true
    link-external-newwindow: true
    embed-resources: true
execute: 
  cache: true
theme:
  light: flatly
  dark: darkly
---


## Packages {.unlisted .hidden}

```{r}
#| echo: false

if (!requireNamespace("pak", quietly = TRUE)) {
  install.packages("pak", repos = "https://cran.r-project.org")
}

packages <- c(
  "dplyr", "tidyr", "ggplot2", "RSQLite", "DBI", "readr",
  "skimr", "DT", "dbplyr", "tibble", "forcats", "purrr"
)
pak::pak(packages)
lapply(packages, library, character.only = TRUE)

```

## Checks

```{r}
#| code-fold: true
#| output: false

in_dir  <- "//10.60.2.10/FF_Lab/project_data/na_boreal/sensitivity_analysis/data/output/debug_data/NorV2debug"
nor_debug <- readr::read_delim(file.path(in_dir, "temp/NorV2debug_establishment.csv"), delim = ";")

dbconn <- DBI::dbConnect(
  RSQLite::SQLite(),
  dbname = file.path(in_dir, "output/NorV2HistBaseFecundDebug.sqlite"))

# Water db
water = tbl(dbconn, "water") |> 
  mutate(
    SOLLayer = ifelse(is.na(SOLLayer), 0, SOLLayer),
    SOL.cm = SOLLayer * 100 + mossLayer * 100) |>
  collect()

dbDisconnect(dbconn)

```

Join up with water output for SOL

```{r}
#| output: false

dim(nor_debug)
# Joining with water for sol
nor_debug <- left_join(nor_debug, water) |>
  mutate(species = forcats::fct(species, levels = c("Pima", "Potr", "Pigl", "Bene")))
dim(nor_debug)

```

Gut check on ru and rid. No ru or rid == -1 (not the cause of `NaN` in TACAAbioticEnv)

```{r}
# range(nor_debug$ru)
# length(unique(nor_debug$ru))

# number of unique rid matches unique ru. good.
# range(nor_debug$rid)
# length(unique(nor_debug$rid))


nor_debug |>
  summarise(across(
    all_of(c("ru", "rid")),
    list(unique = ~ n_distinct(.x, na.rm = TRUE),
         max = ~ max(.x, na.rm = TRUE),
         min = ~ min(.x, na.rm = TRUE)),
         .names = "{.col}_{.fn}"))
```

## What is causing Nans in TACA abioticenv?

Copied from the wiki:

TACAAbioticEnv:

> TACAAbioticEnv result after abiotic environment (after multiplication with fEnvYr) (result of tolerance pow(sqrt(frostdays after budbirst)) * fEnvYr)

fEnvYr:

> still missing: f_ref: parameter that scales response values to the range 0..1 (1 for best growth conditions)

TACAFrostAfterBud:

> frostdays after budbirst are counted to the end of vegetation period

Tree and climate assessment model; [TACA model](https://doi.org/10.1016/j.ecolmodel.2007.07.026)


Possible that the `NaN`s result from underlying calculation `pow(sqrt(frostdays after budbirst)) * fEnvYr`?

Briefly looked into [cpp code](https://github.com/edfm-tum/iland-model/blob/058749ffb2d174d171ca1fc8541baa0f127babbf/src/core/establishment.cpp#L263C13-L263C86), but tests would be time consuming!

How many rids have `NaN`?

```{r}
#| include: false
 
# length(unique(nor_debug$rid)) - length(unique(nor_nan$rid))
nor_nan <- nor_debug[which(is.na(nor_debug$TACAAbioticEnv)), ]
```

Only `r length(unique(nor_nan$rid))` rids, or `r nrow(nor_nan)` rows of data.

The following table is interactive with sort and filter fields, but is excessvely large and I don't see any obvious cause for the `NaN` values.

```{r}
options(htmlwidgets.TOJSON_ARGS = list(na = 'string'))
nor_debug |>
  select(id, year, species, ru, rid, TACAAbioticEnv, SOL.cm, stockable_area, fEnvYr, GDD) |>
  filter(is.na(TACAAbioticEnv)) |>
  datatable(filter = "top", options = list(searching = FALSE))

```



## Try to figure out if any cells have avprobdensity >0  for any of the 4 species

Nope. No cells with avgProbDensity >0 (p100 == max in table below).

```{r}
#| code-fold: true

# variables <- c("TACAAbioticEnv", "N_Established", "avgProbDensity")
# nor_debug |>
#   group_by(species) |>
#   summarise(across(
#     all_of("TACAAbioticEnv"),
#     list(median = ~ median(.x, na.rm = TRUE),
#          q25 = ~ quantile(.x, 0.25, na.rm = TRUE),
#          q75 = ~ quantile(.x, 0.75, na.rm = TRUE),
#          max = ~ max(.x, na.rm = TRUE),
#          min = ~ min(.x, na.rm = TRUE)),
#          .names = "{.col}_{.fn}")) |>
#   datatable() |>
#   formatRound(columns = 2:6, digits = 4)


nor_debug_skim <- nor_debug |>
  select(avgProbDensity, waterLimitation, GDD, TACAAbioticEnv, SOL.cm, fEnvYr)

skim(nor_debug_skim)
```



## 3.Summary plots:

a. N established as a function of TACabioticenv by speices

The following will remove rows where `TACAAbioticEnv` == `NaN`. Not a big loss, only `r length(unique(nor_nan$rid))` rids (`r length(nor_nan$rid)` rows).

```{r}

spp_colours <- c("Pima" = "#8c564b", "Pigl" = "#1f77b4", 
                 "Potr" = "#2ca02c", "Bene" = "green")

nor_debug_spp <- nor_debug |>
  group_by(species)

ggplot(nor_debug_spp, aes (x = N_Established, y = TACAAbioticEnv, colour = species)) +
  geom_point() +
  scale_colour_manual(values = spp_colours) +
  theme_minimal()

ggplot(nor_debug_spp, aes (x = N_Established, y = TACAAbioticEnv, colour = species)) +
  geom_point() +
  scale_colour_manual(values = spp_colours) +
  facet_wrap(~ species) +
  theme_minimal()
```

Same plots but xlim set to 300

```{r}

ggplot(nor_debug_spp, aes (x = N_Established, y = TACAAbioticEnv, colour = species)) +
  geom_point() +
  scale_colour_manual(values = spp_colours) +
  xlim(0, 300) +
  theme_minimal()

ggplot(nor_debug_spp, aes (x = N_Established, y = TACAAbioticEnv, colour = species)) +
  geom_point() +
  scale_colour_manual(values = spp_colours) +
  xlim(0, 300) +
  facet_wrap(~ species) +
  theme_minimal()
```



b. link RID with other output and plot TACabioticenv vs SOL depths

```{r}

ggplot(nor_debug, aes (x = SOL.cm, y = TACAAbioticEnv, colour = species)) +
  geom_point() +
  scale_colour_manual(values = spp_colours) +
  theme_minimal()

ggplot(nor_debug_spp, aes (x = SOL.cm, y = TACAAbioticEnv, colour = species)) +
  geom_point() +
  scale_colour_manual(values = spp_colours) +
  facet_wrap(~ species) +
  theme_minimal()
```



```{r}
nor_debug |>
  filter(species %in% c("Pigl", "Bene"),
         TACAAbioticEnv == 0,
         SOL.cm < 6) |>
  select(species, rid, TACAgdd, TACAFrostAfterBud, GDD, TACAAbioticEnv, N_Established, stockable_area, SOL.cm) |>
  datatable(filter = "top", options = list(searching = FALSE)) |>
  formatRound(columns = "SOL.cm", digits = 3)


```




c. summarizing full ranges of avprobdensity, tacaabiotic env, fenv, and Nestablished by species.

```{r}
#| code-fold: true

variables <- c("TACAAbioticEnv", "N_Established", "avgProbDensity")
nor_debug_tables <- map(variables, \(var) {
  tab <- nor_debug |>
    group_by(species) |>
    summarise(across(
      all_of(var),
      list(median = ~ median(.x, na.rm = TRUE),
           q25 = ~ quantile(.x, 0.25, na.rm = TRUE),
           q75 = ~ quantile(.x, 0.75, na.rm = TRUE),
           max = ~ max(.x, na.rm = TRUE),
           min = ~ min(.x, na.rm = TRUE)),
           .names = "{.col}_{.fn}"))
}) |>
set_names(variables)

tab_list <- map(nor_debug_tables, \(tab) {
  var_tab <- datatable(tab, options = list(searching = FALSE))
  formatRound(var_tab, columns = 2:ncol(tab), digits = 4)
  var_tab <- htmltools::div(var_tab)
})

# tab_list
```


:::{.panel-tabset}
```{r, results='asis'}
#| warning: false

for (i in seq_along(tab_list)) {
  cat("# ", variables[i],"\n")
  tab_list[[i]]
  cat('\n\n')
}
```
:::


::: {.panel-tabset}
## TACAAbioticEnv

```{r}
#| echo: false
tab_list[[1]]
```

## N_Established

```{r}
#| echo: false
tab_list[[2]]
```

## avgProbDensity

```{r}
#| echo: false
tab_list[[3]]
```

:::


```{r}
ggplot(nor_debug, aes (x = fEnvYr, fill = species)) +
  geom_histogram() +
  scale_fill_manual(values = spp_colours) +
  theme_minimal()

ggplot(nor_debug, aes (x = SOL.cm)) +
  geom_histogram() +
  theme_minimal()  
```

## ~~If they have avprobdensity >0 try to figure out where the cells are relative to nearby dominant forest types~~

probDensity all 0
