---
title: "Seedling density"
author: "FFL"
date: today
---

# Data {.unnumbered .unlisted .hidden}


```{r}
#| label: data processing
#| cache: false
source("./scripts/install.R")

out_dir  <- here("data", "estSOLthickness", "seeddens_processed")
out_file <- here("data", "estSOLthickness", "seeddens_processed", "seedling_density_data.rds")
in_dir   <- "//10.60.2.10/FF_Lab/project_data/na_boreal/sensitivity_analysis/data/output/estSOLthickness/seeddens"
if(!file.exists(in_dir)) {
  stop("Are you connected to Janet?")
}

dir.create(out_dir, recursive = TRUE)

if (!file.exists(out_file)) {
  seeddens_files <- list.files(in_dir, full.names = TRUE)

  # plan(sequential)
  plan(multisession, workers = 5)
  seedling_density_data <- furrr::future_map_dfr(seeddens_files, \(x) {
    read_delim(x, show_col_types = FALSE) |>
      select(year, rid, ru, fire.year, fireId, reburned_rid, rel.year,
             SOL.cm, starts_with(c("sapling")), treatment, replicate)
  })


  seedling_density_data_fire_filt <- seedling_density_data |> 
    group_by(treatment, replicate, reburned_rid) |> 
    drop_na(reburned_rid) |> # Drop anything burned only once
    distinct(fireId, fire.year) |> # include fireId for case when more than one fire occurred in the same year
    mutate(tsf_group_diff = coalesce(lead(fire.year) - fire.year, fire.year)) |>  # Use coalesce to replace NA from lead with fire.  year
    right_join(seedling_density_data) |>
    mutate(tsf = year - fire.year)
  
  # Here we are filtering for rid with 4 years since fire
  # including rids that burned twice if more that 4 years between fires
  fire_filt_pos <- seedling_density_data_fire_filt |> 
    filter((tsf_group_diff >= 4 | is.na(tsf_group_diff)), # Keep NAs in tsf group because NA are rids that did not reburn
            tsf > 0, tsf <= 4) |> # filtering out 0 year
    arrange(reburned_rid) # |>
  # mutate(treatment = str_remove_all(treatment, "CPCRW_|_1_seedling_dens"))
  # file names were getting too long, so I clipped them in the mutate ^
  # Can remove when no longer necessary
  
  # the following is all Lora's code.
  # ONLY SMALL SAPLINGS < 1.3m
  small_seedl_dens_m2 <- fire_filt_pos |>
    rename(
      Bene.nseedl = sapling_count_ha_small.Bene,
      Potr.nseedl = sapling_count_ha_small.Potr,
      Pima.nseedl = sapling_count_ha_small.Pima,
      Pigl.nseedl = sapling_count_ha_small.Pigl) |>
    dplyr::select(
      treatment, replicate, reburned_rid,
      year, ru, fireId, tsf, Bene.nseedl,
      Potr.nseedl, Pima.nseedl, Pigl.nseedl) |>
    mutate(
      Pigl.seedl.dens = ifelse(is.na(Pigl.nseedl), 0, Pigl.nseedl/10000),
      Pima.seedl.dens = ifelse(is.na(Pima.nseedl), 0, Pima.nseedl/10000),
      Potr.seedl.dens = ifelse(is.na(Potr.nseedl), 0, Potr.nseedl/10000),
      Bene.seedl.dens = ifelse(is.na(Bene.nseedl), 0, Bene.nseedl/10000))
  
  # Create the fire severity code based on SOL thickness in the first year after
  levels <- c("low", "mod-low", "moderate", "mod-high", "high")
  fire_severity <- fire_filt_pos |>
    filter(tsf == 1) |>
    mutate(soil.sev = levels[cut(SOL.cm, breaks = c(0, 0.5, 1.5, 3, 10, 100))]) |>
    dplyr::select(treatment, replicate, reburned_rid, ru, fireId, soil.sev)
  
    # Add the fire severity to the seedling data and subset to the same years
  # Originally filtered by observational data: filter(rel.year %in% c(2,4,7,13)) 
  fire_severity <- left_join(small_seedl_dens_m2, fire_severity) |>
    pivot_longer(cols = ends_with(".dens"), names_to = "species", values_to = "seedl.dens") |>
    mutate(species = fct(str_remove(species, ".seedl.dens"),
                         levels = c("Bene", "Pima", "Potr", "Pigl")),
           soil.sev = fct(soil.sev, levels = c("low", "mod-low", "moderate", "mod-high", "high")))
  
  fire_severity_summary <- fire_severity |>
    group_by(treatment, replicate, soil.sev, species, tsf) |> 
    summarise(avg.seedl.den = mean(seedl.dens, na.rm = TRUE))
  
  
  small_seedl_dens_m2_long <- small_seedl_dens_m2 |>
    pivot_longer(cols = ends_with(".dens"), names_to = "species", values_to = "seedl.dens") |>
    mutate(species = fct(str_remove(species, ".seedl.dens"),
                         levels = c("Bene", "Pima", "Potr", "Pigl"))) |>
    group_by(treatment, replicate, species, tsf) |>
    summarise(mean.seedl = mean(seedl.dens))
  
  # combining sapling variables
  all_seedl_dens_m2 <- fire_filt_pos |>
    filter(tsf == 2) |>
    ungroup() |> # Ungroup necessary for cut, I think ok to ungroup
    mutate(sap1.Bene = ifelse(is.na(sapling_count_ha_small.Bene), 0, sapling_count_ha_small.Bene),
           sap2.Bene = ifelse(is.na(sapling_count_ha.Bene), 0, sapling_count_ha.Bene),
           sap1.Potr = ifelse(is.na(sapling_count_ha_small.Potr), 0, sapling_count_ha_small.Potr),
           sap2.Potr = ifelse(is.na(sapling_count_ha.Potr), 0, sapling_count_ha.Potr),
           sap1.Pima = ifelse(is.na(sapling_count_ha_small.Pima), 0, sapling_count_ha_small.Pima),
           sap2.Pima = ifelse(is.na(sapling_count_ha.Pima), 0, sapling_count_ha.Pima),
           sap1.Pigl = ifelse(is.na(sapling_count_ha_small.Pigl), 0, sapling_count_ha_small.Pigl),
           sap2.Pigl = ifelse(is.na(sapling_count_ha.Pigl), 0, sapling_count_ha.Pigl),
           Bene.nseedl = (sap1.Bene + sap2.Bene),
           Potr.nseedl = (sap1.Potr + sap2.Potr),
           Pima.nseedl = (sap1.Pima + sap2.Pima),
           Pigl.nseedl = (sap1.Pigl + sap2.Pigl),
           Pigl.seedl.dens = ifelse(is.na(Pigl.nseedl), 0, Pigl.nseedl/10000),
           Pima.seedl.dens = ifelse(is.na(Pima.nseedl), 0, Pima.nseedl/10000),
           Potr.seedl.dens = ifelse(is.na(Potr.nseedl), 0, Potr.nseedl/10000),
           Bene.seedl.dens = ifelse(is.na(Bene.nseedl), 0, Bene.nseedl/10000),
           soil.thick = cut(SOL.cm, breaks = seq(0, ceiling(max(SOL.cm) / 5) * 5, by = 5))) |>
  # filter(rel.year %in% c(2,4,7,13)) |>
    pivot_longer(cols = ends_with(".dens"), names_to = "species", values_to = "seedl.dens") |>
    mutate(species = fct(str_remove(species, ".seedl.dens"),
                         levels = c("Pima", "Pigl", "Potr", "Bene"))) |>
    select(replicate, treatment, rid, tsf, species, soil.thick, seedl.dens)
  
  
  levels(all_seedl_dens_m2$soil.thick) <- sapply(levels(all_seedl_dens_m2$soil.thick), 
                      function(x){gsub("]","",strsplit(x, ",")[[1]][2], fixed=T)})
  
  # We need a few objects from this data wrangling, we can save an RDS file instead of individual csv's
  # RDS is a native R format so, no issues there. It just saves one list object with all the data
  saveRDS(list(
    fire_severity_summary = fire_severity_summary,
    small_seedl_dens_m2_long = small_seedl_dens_m2_long,
    all_seedl_dens_m2 = all_seedl_dens_m2), out_file)

} else {
  seedling_density_data <- readRDS(out_file)
}

```

# Figures


```{r}
#| echo: false
#| warning: false
list2env(seedling_density_data, envir = .GlobalEnv)

fill_colours <- c(Pima = "#8c564b", Pigl = "#1f77b4", 
                  Potr = "#2ca02c", Bene = "green")

colours <- c(
  "Pima" = "#8c564b", "Pigl" = "#1f77b4",
  "Potr" = "#2ca02c", "Bene" = "green",
  "Mixed.spruce" = "#9467bd", "Mixed.deciduous" = "#bcbd22")

all_seedl_dens_m2 <- all_seedl_dens_m2 |> 
  mutate(
    fecund    = as.numeric(str_extract(treatment, "(?<=basefecund_)[0-9.]+")),
    ep_part   = replace_na(str_extract(treatment, "ep[0-9.]+"), "ep2.7"),
    treatment = paste0(
      substr(treatment, 1, 6),
      "_base_est", fecund,
      "_", ep_part,
      "_dbh0.1"
    ),
    treatment = factor(treatment, levels = unique(treatment[order(fecund)]))
  )

```

```{r}
#| label: post fire seedling fixed
#| echo: false
#| fig-height: 10

# Since the plots of each treatment are independent plots with facet_wrap, setting the ylim does not work because it will fix ylim across all fecets, and using scales = will not be the same across treatments.

# We can trick ggplot by making a blank plot with fixed dimensions for each facet and apply it across treatments. I had done this manually, but it broke when the data changed and had a different range.

# Now I have the dummy plot generating programatically from the range of the data, rounding up to the nearest decimal to ensure the range stays fixed.

facet_dummy <- all_seedl_dens_m2 |>
  group_by(species) |>
  slice_max(seedl.dens) |>
  mutate(ylim = ceiling(seedl.dens * 10) / 10) |>
  select(species, soil.thick, ylim)


all_seedl_dens_m2_plot_list2 <- all_seedl_dens_m2 |>
  split(~ treatment) |>
  lapply(\(x){
    ggplot(x, aes(x = soil.thick, y = seedl.dens, fill = species)) +
      geom_blank(data = facet_dummy, aes(x = soil.thick, y = ylim)) +
      geom_boxplot() +
      scale_fill_manual(values = fill_colours, guide = "none") +
      theme_minimal() + 
    #  facet_grid(rows = vars(species), cols = vars(tsf), scales = "free") +
      facet_wrap(~ species, scales = "free") +
      theme(text = element_text(color = "black", family = "sans", size = 10),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            axis.text.x = element_text(colour = "black", size = 10),
            axis.text.y = element_text(colour = "black", size = 10),
            legend.text = element_text(size = 8),
            legend.position= "bottom",
            legend.background = element_rect(colour = "transparent", fill = 'transparent'),
            legend.key = element_rect(colour = "transparent", fill = 'transparent'),
            legend.key.size = unit(0.5, "cm")) +
            labs(x = "Surface organic layer (cm)", y = bquote(Seedlings ~ m^-2))
})

headings <- names(all_seedl_dens_m2_plot_list2)


```



## NorEsm epsilon

Here are figures for NorEsm looking at default estSOLthickness = 0.5 and varying epsilon (default2.7)

:::{.panel-tabset}
```{r}
#| warning: false
#| results: asis
#| fig-height: 7

epsilon_plots <- headings[grep(x = headings, pattern = "^NorEsm.*est0\\.5")]
epsilon_plot_list <- all_seedl_dens_m2_plot_list2[names(all_seedl_dens_m2_plot_list2) %in% epsilon_plots]

for (i in seq_along(epsilon_plot_list)) {
  cat("# ",epsilon_plots[i],"\n")
  print(epsilon_plot_list[[i]])
  cat('\n\n')
}
```
:::


## NorEsm estSOLthickness

These figures vary estSOLthickness and keep epsilon at default 2.7:

:::{.panel-tabset}
```{r}
#| warning: false
#| results: asis
#| fig-height: 7

est_plots <- headings[grep(x = headings, pattern = "^NorEsm.*est[0-9.]+.*ep2\\.7")]
est_plot_list <- all_seedl_dens_m2_plot_list2[names(all_seedl_dens_m2_plot_list2) %in% est_plots]

for (i in seq_along(est_plot_list)) {
  cat("# ",est_plots[i],"\n")
  print(est_plot_list[[i]])
  cat('\n\n')
}
```
:::


## NorEsm estSOLthickness + epsilon

Same figures as above, but all of them.

:::{.panel-tabset}
```{r}
#| warning: false
#| results: asis
#| fig-height: 7

est_ep_plots <- headings[grep(x = headings, pattern = "NorEsm")]
est_ep_plot_list <- all_seedl_dens_m2_plot_list2[names(all_seedl_dens_m2_plot_list2) %in% est_ep_plots]

for (i in seq_along(est_ep_plot_list)) {
  cat("# ",est_ep_plots[i],"\n")
  print(est_ep_plot_list[[i]])
  cat('\n\n')
}
```
:::




## NorEsm vs TaiEsm estSOLthickness

Epsilon fixed at default 2.7

```{r}
#| warning: false
#| echo: false
#| fig-height: 12
#| fig-width: 8

nor_tai_plots <- headings[grep(x = headings, pattern = "^(NorEsm|TaiESM).*est[0-9.]+.*ep2\\.7")]

seedl_dens_nor_tai <- all_seedl_dens_m2 |>
  filter(treatment %in% nor_tai_plots) |>
  mutate(gcm = substr(treatment, 1, 6),
         est = str_extract(treatment, "est[0-9.]+"))

ggplot(seedl_dens_nor_tai, aes(x = soil.thick, y = seedl.dens, fill = species)) +
  geom_blank(data = facet_dummy, aes(x = soil.thick, y = ylim)) +
  geom_boxplot() +
  scale_fill_manual(values = fill_colours, guide = "none") +
  theme_minimal() + 
#  facet_grid(rows = vars(species), cols = vars(tsf), scales = "free") +
  facet_grid(est ~ gcm) +
  theme(text = element_text(color = "black", family = "sans", size = 10),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(colour = "black", size = 10),
        axis.text.y = element_text(colour = "black", size = 10),
        legend.text = element_text(size = 8),
        legend.position= "bottom",
        legend.background = element_rect(colour = "transparent", fill = 'transparent'),
        legend.key = element_rect(colour = "transparent", fill = 'transparent'),
        legend.key.size = unit(0.5, "cm")) +
        labs(x = "Surface organic layer (cm)", y = bquote(Seedlings ~ m^-2))
```

## NorEsm vs TaiEsm epsilon

estSOLthick fixed at default 0.5

```{r}
#| warning: false
#| echo: false
#| fig-height: 12
#| fig-width: 8

nor_tai_plots_ep <- headings[grep(x = headings, pattern = "^(NorEsm|TaiESM).*est[0-9.]+.*ep2\\.7"
)]

seedl_dens_nor_tai_ep <- all_seedl_dens_m2 |>
  filter(treatment %in% nor_tai_plots_ep) |>
  mutate(gcm = substr(treatment, 1, 6),
         ep = str_extract(treatment, "ep[0-9.]+"))

ggplot(seedl_dens_nor_tai_ep, aes(x = soil.thick, y = seedl.dens, fill = species)) +
  geom_blank(data = facet_dummy, aes(x = soil.thick, y = ylim)) +
  geom_boxplot() +
  scale_fill_manual(values = fill_colours, guide = "none") +
  theme_minimal() + 
#  facet_grid(rows = vars(species), cols = vars(tsf), scales = "free") +
  facet_grid(ep ~ gcm) +
  theme(text = element_text(color = "black", family = "sans", size = 10),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(colour = "black", size = 10),
        axis.text.y = element_text(colour = "black", size = 10),
        legend.text = element_text(size = 8),
        legend.position= "bottom",
        legend.background = element_rect(colour = "transparent", fill = 'transparent'),
        legend.key = element_rect(colour = "transparent", fill = 'transparent'),
        legend.key.size = unit(0.5, "cm")) +
        labs(x = "Surface organic layer (cm)", y = bquote(Seedlings ~ m^-2))
```


## Hide {.unnumbered .unlisted .hidden}

### Post-fire regeneration with soil organic layer thickness post-fire

This plot of regeneration post-fire with the fire severity index is not in the main POLE-FM paper. It is in Lora's code, so I recreated it while going through. Each column is the number of years post-fire from 1-4, each row is a treatment.

-   Only sapling_count_small

```{r}
#| label: fire severity
#| fig-height: 10

ggplot(data = fire_severity_summary, aes(x = soil.sev, y = avg.seedl.den)) +
   geom_col(aes(fill=species)) +
   theme_bw() +
   scale_fill_manual(values = fill_colours) +
   labs(title = paste("iLand seedling regeneration post-fire"),
        x = "Fire severity on soil", y = "Mean Seedlings / m2") +
   facet_grid(rows = vars(treatment), cols = vars(tsf))

```

### Average seedling density post-fire

This plot is also not in the main POLE-FM paper. Shows mean sapling density for four years after fire.

-   Only sapling_count_small

```{r}
#| label: average seedling density

ggplot(small_seedl_dens_m2_long, aes(x = tsf, y = mean.seedl, fill = species)) +
  geom_col() +
  facet_wrap(~ treatment) +
  scale_fill_manual(values = fill_colours) +
  theme_minimal()

```

