---
title: "Seedling density spinup"
author: "FFL"
date: today
---

# Data {.unnumbered .unlisted .hidden}


```{r}
#| label: data_processing
#| warning: false

source("./scripts/install.R")

out_dir  <- here("data", "spinup", "seeddens_processed")

seeddens_dom_dir <- list.dirs("//10.60.2.10/FF_Lab/project_data/na_boreal/sensitivity_analysis/data/output/spinup_summaries/seeddens")
rep_dirs <- grep(seeddens_dom_dir, pattern = "rep_")
dirs <- seeddens_dom_dir[rep_dirs]
dirs <- dirs[!grepl(dirs, pattern = "x")]

if(any(!file.exists(dirs))) {
  stop("Are you connected to Janet?")
}

dir.create(out_dir, recursive = TRUE)

if (!file.exists(file.path(out_dir, "spinup_seedling_data.rds"))) {
  seeddens_files <- list.files(dirs, full.names = TRUE)
  seeddens_files <- seeddens_files[as.integer(sub(".*chunk_(\\d+)_.*\\.parquet$", "\\1", seeddens_files)) >= 250]

  seedling_density_data <- arrow::open_dataset(seeddens_files, unify_schemas = TRUE) |>
    select(year, rid, ru, fire.year, fireId, reburned_rid, rel.year, SOL.cm, 
           starts_with("sapling"), treatment, replicate) |>
    filter(year >= 260) |>
    collect()

  format(object.size(seedling_density_data), units = "Gb")

  seedling_density_data_fire_filt <- seedling_density_data |> 
    group_by(treatment, replicate, reburned_rid) |> 
    drop_na(reburned_rid) |> # Drop anything burned only once
    distinct(fireId, fire.year) |> # include fireId for case when more than one fire occurred in the same year
    mutate(rel.year_group_diff = coalesce(lead(fire.year) - fire.year, fire.year)) |>  # Use coalesce to replace NA from lead with fire.  year
    right_join(seedling_density_data)
  
  rm(seedling_density_data)

  # Here we are filtering for rid with 4 years since fire
  # including rids that burned twice if more that 4 years between fires
  fire_filt_pos <- seedling_density_data_fire_filt |> 
    filter((rel.year_group_diff >= 4 | is.na(rel.year_group_diff)), # Keep NAs in rel.year group because NA are rids that did not reburn
            rel.year > 0, rel.year <= 4) |> # filtering out 0 year
    arrange(reburned_rid) # |>
  # mutate(treatment = str_remove_all(treatment, "CPCRW_|_1_seedling_dens"))
  # file names were getting too long, so I clipped them in the mutate ^
  # Can remove when no longer necessary
  
  # the following is all Lora's code.
  # ONLY SMALL SAPLINGS < 1.3m
  small_seedl_dens_m2 <- fire_filt_pos |>
    rename(
      Bene.nseedl = sapling_count_ha_small.Bene,
      Potr.nseedl = sapling_count_ha_small.Potr,
      Pima.nseedl = sapling_count_ha_small.Pima,
      Pigl.nseedl = sapling_count_ha_small.Pigl) |>
    dplyr::select(
      treatment, replicate, reburned_rid,
      year, ru, fireId, rel.year, Bene.nseedl,
      Potr.nseedl, Pima.nseedl, Pigl.nseedl) |>
    mutate(
      Pigl.seedl.dens = ifelse(is.na(Pigl.nseedl), 0, Pigl.nseedl/10000),
      Pima.seedl.dens = ifelse(is.na(Pima.nseedl), 0, Pima.nseedl/10000),
      Potr.seedl.dens = ifelse(is.na(Potr.nseedl), 0, Potr.nseedl/10000),
      Bene.seedl.dens = ifelse(is.na(Bene.nseedl), 0, Bene.nseedl/10000))
  
  # Create the fire severity code based on SOL thickness in the first year after
  levels <- c("low", "mod-low", "moderate", "mod-high", "high")
  fire_severity <- fire_filt_pos |>
    filter(rel.year == 1) |>
    mutate(soil.sev = levels[cut(SOL.cm, breaks = c(0, 0.5, 1.5, 3, 10, 100))]) |>
    dplyr::select(treatment, replicate, reburned_rid, ru, fireId, soil.sev)
  
    # Add the fire severity to the seedling data and subset to the same years
  # Originally filtered by observational data: filter(rel.year %in% c(2,4,7,13)) 
  fire_severity <- left_join(small_seedl_dens_m2, fire_severity) |>
    pivot_longer(cols = ends_with(".dens"), names_to = "species", values_to = "seedl.dens") |>
    mutate(species = fct(str_remove(species, ".seedl.dens"),
                         levels = c("Bene", "Pima", "Potr", "Pigl")),
           soil.sev = fct(soil.sev, levels = c("low", "mod-low", "moderate", "mod-high", "high")))
  
  fire_severity_summary <- fire_severity |>
    group_by(treatment, replicate, soil.sev, species, rel.year) |> 
    summarise(avg.seedl.den = mean(seedl.dens, na.rm = TRUE))
  
  
  small_seedl_dens_m2_long <- small_seedl_dens_m2 |>
    pivot_longer(cols = ends_with(".dens"), names_to = "species", values_to = "seedl.dens") |>
    mutate(species = fct(str_remove(species, ".seedl.dens"),
                         levels = c("Bene", "Pima", "Potr", "Pigl"))) |>
    group_by(treatment, replicate, species, rel.year) |>
    summarise(mean.seedl = mean(seedl.dens))
  
  # combining sapling variables
  all_seedl_dens_m2 <- fire_filt_pos |>
    filter(rel.year == 2) |>
    ungroup() |> # Ungroup necessary for cut, I think ok to ungroup
    mutate(sap1.Bene = ifelse(is.na(sapling_count_ha_small.Bene), 0, sapling_count_ha_small.Bene),
           sap2.Bene = ifelse(is.na(sapling_count_ha.Bene), 0, sapling_count_ha.Bene),
           sap1.Potr = ifelse(is.na(sapling_count_ha_small.Potr), 0, sapling_count_ha_small.Potr),
           sap2.Potr = ifelse(is.na(sapling_count_ha.Potr), 0, sapling_count_ha.Potr),
           sap1.Pima = ifelse(is.na(sapling_count_ha_small.Pima), 0, sapling_count_ha_small.Pima),
           sap2.Pima = ifelse(is.na(sapling_count_ha.Pima), 0, sapling_count_ha.Pima),
           sap1.Pigl = ifelse(is.na(sapling_count_ha_small.Pigl), 0, sapling_count_ha_small.Pigl),
           sap2.Pigl = ifelse(is.na(sapling_count_ha.Pigl), 0, sapling_count_ha.Pigl),
           Bene.nseedl = (sap1.Bene + sap2.Bene),
           Potr.nseedl = (sap1.Potr + sap2.Potr),
           Pima.nseedl = (sap1.Pima + sap2.Pima),
           Pigl.nseedl = (sap1.Pigl + sap2.Pigl),
           Pigl.seedl.dens = ifelse(is.na(Pigl.nseedl), 0, Pigl.nseedl/10000),
           Pima.seedl.dens = ifelse(is.na(Pima.nseedl), 0, Pima.nseedl/10000),
           Potr.seedl.dens = ifelse(is.na(Potr.nseedl), 0, Potr.nseedl/10000),
           Bene.seedl.dens = ifelse(is.na(Bene.nseedl), 0, Bene.nseedl/10000),
           soil.thick = cut(SOL.cm, breaks = seq(0, ceiling(max(SOL.cm) / 5) * 5, by = 5))) |>
  # filter(rel.year %in% c(2,4,7,13)) |>
    pivot_longer(cols = ends_with(".dens"), names_to = "species", values_to = "seedl.dens") |>
    mutate(species = fct(str_remove(species, ".seedl.dens"),
                         levels = c("Pima", "Pigl", "Potr", "Bene"))) |>
    select(replicate, treatment, rid, rel.year, species, soil.thick, seedl.dens)
  
  
  levels(all_seedl_dens_m2$soil.thick) <- sapply(levels(all_seedl_dens_m2$soil.thick), 
                      function(x){gsub("]","",strsplit(x, ",")[[1]][2], fixed=T)})
  
  # We need a few objects from this data wrangling, we can save an RDS file instead of individual csv's
  # RDS is a native R format so, no issues there. It just saves one list object with all the data
  saveRDS(list(
    fire_severity_summary = fire_severity_summary,
    small_seedl_dens_m2_long = small_seedl_dens_m2_long,
    all_seedl_dens_m2 = all_seedl_dens_m2), file.path(out_dir, "spinup_seedling_data.rds"))

  } else {
    spinup_seedling_data <- readRDS(file.path(out_dir, "spinup_seedling_data.rds"))
}
```




# Figures


```{r}
#| echo: false
#| warning: false
#| output: false

list2env(spinup_seedling_data, envir = .GlobalEnv)


fill_colours <- c(Pima = "#8c564b", Pigl = "#1f77b4", 
                  Potr = "#2ca02c", Bene = "green")


all_seedl_dens_m2 <- all_seedl_dens_m2 |>
  mutate(
    rep_sp  = factor(paste(species, replicate, sep = "_"))
    )

u <- unique(all_seedl_dens_m2[, c("species", "rep_sp")])
fill_colours_full <- setNames(fill_colours[u$species], u$rep_sp)
```

```{r}
#| label: post fire seedling fixed
#| echo: false
#| fig-height: 10


facet_dummy <- all_seedl_dens_m2 |>
  group_by(rep_sp) |>
  slice_max(seedl.dens) |>
  mutate(ylim = ceiling(seedl.dens * 10) / 10) |>
  select(species, soil.thick, ylim)


all_seedl_dens_m2_plot_list <- all_seedl_dens_m2 |>
  filter(grepl("NorEsm", treatment),
         grepl("2.7", treatment)) |>
  split(~ list(treatment)) |>
  lapply(\(x){
    ggplot(x, aes(x = soil.thick, y = seedl.dens, fill = rep_sp)) +
      geom_blank(data = facet_dummy, aes(x = soil.thick, y = ylim)) +
      geom_boxplot() +
      scale_fill_manual(values = fill_colours_full, guide = "none") +
      theme_minimal() + 
    #  facet_grid(rows = vars(species), cols = vars(rel.year), scales = "free") +
      facet_wrap(~ species, scales = "free") +
      theme(text = element_text(color = "black", family = "sans", size = 12),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            axis.text.x = element_text(colour = "black", size = 12),
            axis.text.y = element_text(colour = "black", size = 12),
            legend.text = element_text(size = 8),
            legend.position= "bottom",
            legend.background = element_rect(colour = "transparent", fill = 'transparent'),
            legend.key = element_rect(colour = "transparent", fill = 'transparent'),
            legend.key.size = unit(0.5, "cm"),
            strip.text.x = element_text(size = 10)) +
            labs(x = "Surface organic layer (cm)", y = bquote(Seedlings ~ m^-2))
})

headings <- names(all_seedl_dens_m2_plot_list)


```



## Postfire seedling densities 2 years after fire IN THE LAST 40 years of simulation

:::{.panel-tabset}
```{r}
#| warning: false
#| results: asis
#| fig-height: 7

for (i in seq_along(all_seedl_dens_m2_plot_list)) {
  cat("# ",headings[i],"\n")
  print(all_seedl_dens_m2_plot_list[[i]])
  cat('\n\n')
}
```
:::



```{r}
#| label: BySpecies
#| echo: false
#| fig-height: 10

all_seedl_dens_m2_plot_list2 <- all_seedl_dens_m2 |>
  filter(grepl("NorEsm", treatment),
         grepl("2.7", treatment)) |>
  split(~ list(treatment, species)) |>
  lapply(\(x){
    
    facet_dummy <- x |>
      slice_max(seedl.dens) |>
      mutate(ylim = ceiling(seedl.dens * 10) / 10) |>
      select(species, soil.thick, ylim, rep_sp)

    ggplot(x, aes(x = soil.thick, y = seedl.dens, fill = rep_sp)) +
      geom_blank(data = facet_dummy, aes(x = soil.thick, y = ylim)) +
      geom_boxplot() +
      scale_fill_manual(values = fill_colours_full, guide = "none") +
      theme_minimal() + 
      facet_wrap(~ treatment, scales = "free") +
      theme(text = element_text(color = "black", family = "sans", size = 10),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            axis.text.x = element_text(colour = "black", size = 10),
            axis.text.y = element_text(colour = "black", size = 10),
            legend.text = element_text(size = 8),
            legend.position= "bottom",
            legend.background = element_rect(colour = "transparent", fill = 'transparent'),
            legend.key = element_rect(colour = "transparent", fill = 'transparent'),
            legend.key.size = unit(0.5, "cm"),
            strip.text.x = element_text(size = 8)) +
            labs(x = "Surface organic layer (cm)", y = bquote(Seedlings ~ m^-2))
})

headings2 <- names(all_seedl_dens_m2_plot_list2)


```



:::{.panel-tabset}
```{r}
#| warning: false
#| results: asis
#| fig-height: 7

for (i in seq_along(all_seedl_dens_m2_plot_list2)) {
  cat("# ",headings2[i],"\n")
  print(all_seedl_dens_m2_plot_list2[[i]])
  cat('\n\n')
}
```
:::


## Hide {.unnumbered .unlisted .hidden}

### Post-fire regeneration with soil organic layer thickness post-fire

This plot of regeneration post-fire with the fire severity index is not in the main POLE-FM paper. It is in Lora's code, so I recreated it while going through. Each column is the number of years post-fire from 1-4, each row is a treatment.

-   Only sapling_count_small

```{r}
#| label: fire severity
#| fig-height: 10

ggplot(data = fire_severity_summary, aes(x = soil.sev, y = avg.seedl.den)) +
   geom_col(aes(fill=species)) +
   theme_bw() +
   scale_fill_manual(values = fill_colours) +
   labs(title = paste("iLand seedling regeneration post-fire"),
        x = "Fire severity on soil", y = "Mean Seedlings / m2") +
   facet_grid(rows = vars(treatment), cols = vars(rel.year))

```

### Average seedling density post-fire

This plot is also not in the main POLE-FM paper. Shows mean sapling density for four years after fire.

-   Only sapling_count_small

```{r}
#| label: average seedling density

ggplot(small_seedl_dens_m2_long, aes(x = rel.year, y = mean.seedl, fill = species)) +
  geom_col() +
  facet_wrap(~ treatment) +
  scale_fill_manual(values = fill_colours) +
  theme_minimal()

```



```{r}
#| eval: false

images_out <- "./images/seed_dens"
dir.create(images_out)

imap(all_seedl_dens_m2_plot_list, \(plt, nm) {
  print(nm)
  num <- str_extract(nm, "[0-9]+\\.[0-9]+")
  new_nm <- paste0("Epsilon ", num)
  plt <- plt + labs(title = new_nm)
  ggsave(
    file.path(images_out, paste0("seeddens_", nm, ".png")),
    plt,
    device = "png",
    dpi = 300,
    width = 12,
    height = 6,
    units = "in"
  )
})


```


```{r}
#| eval: false


all_seedl_dens_m2_4 <- all_seedl_dens_m2 |>
  filter(treatment == "ep2.7",
         replicate == 4)

facet_names <- c(
  Bene             = "Birch",
  Pigl             = "White spruce",
  Pima             = "Black spruce",
  Potr             = "Aspen"
)

seedling_plt <- ggplot(all_seedl_dens_m2_4, aes(x = soil.thick, y = seedl.dens, fill = species)) +
  geom_boxplot() +
  scale_fill_manual(values = fill_colours, guide = "none") +
  facet_wrap(~ species, scales = "free", labeller = labeller(species = facet_names)) +
  labs(x = "Surface organic layer (cm)", y = bquote(Seedlings ~ m^-2)) +
  theme_minimal() + 
  theme(
    text = element_text(color = "black", family = "sans", size = 10),
    legend.position= "bottom",
    legend.background = element_rect(colour = "transparent", fill = 'transparent'),
    legend.key = element_rect(colour = "transparent", fill = 'transparent'),
    legend.key.size = unit(0.5, "cm"),
    legend.text = element_text(size = 26, colour = "black"),
    legend.title = element_text(size = 26, colour = "black"),
    legend.key.width = unit(4, "cm"),
    axis.text = element_text(size = 26, colour = "black"),
    axis.title = element_text(size = 28, colour = "black"),
    strip.text = element_text(size = 26, colour = "black")
    )


height_cm <- 32
width_cm  <- height_cm * 1.45

ggsave(
  file.path("C:/Users/asenaq/Documents/workshops_conferences/AGU-2025/poster/seedling_plt.svg"),
  seedling_plt,
  device = "svg",
  width = width_cm,
  height = height_cm,
  units = "cm"
)  


```