---
title: "Percent post-fire dominance"
author: "FFL"
date: today
---

# Data {.unnumbered .unlisted .hidden}

```{r}
#| echo: false
#| warning: false
source("./scripts/install.R")

basal_out_dir  <- here("data", "spinup", "basal_processed")
dir.create(basal_out_dir, recursive = TRUE)

basal_area_dir <- list.dirs("//10.60.2.10/FF_Lab/project_data/na_boreal/sensitivity_analysis/data/output/spinup_summaries/basal_area")
rep_dirs <- grep(basal_area_dir, pattern = "rep_", value = TRUE)

if(any(!file.exists(rep_dirs))) {
  stop("Are you connected to Janet?")
}

if (!file.exists(file.path(basal_out_dir, "basal_area_spinup.rds"))) {
  basal_area_files <- list.files(rep_dirs, full.names = TRUE)

  basal_area_burned_area_year300 <- arrow::open_dataset(basal_area_files) |>
    collect()
  
  saveRDS(basal_area_burned_area_year300, file.path(basal_out_dir, "basal_area_spinup.rds"))

} else {
  basal_area_burned_area_year300 <- readRDS(file.path(basal_out_dir, "basal_area_spinup.rds"))
}

```



```{r}
#| echo: false
#| warning: false

map <- terra::rast("Z:/project_data/na_boreal/sensitivity_analysis/data/output/spinup_summaries/fire_regime/NorEsm2-MMssp126V2_species_param_basefecund_0.5_ep2.7_dbh2.5/rep_1/spinup_300.asc")
map
terra::app(map,max)
map_df <- as.data.frame(map, xy=T)
map_df <- map_df |>
  rename(ru = spinup_300)

coord <- map_df |> 
  dplyr::select(x,y)

init_species <- terra::rast("Z:/project_data/na_boreal/sensitivity_analysis/data/output/spinup_summaries/forest_species_init.tif")
#plot(init_species)
data <- data.frame(cbind(map_df, extract(init_species, coord)))
final_init <- data |>
  rename("init_species" = forest_species_init) |>
  mutate(init_species = replace_na(init_species, 0))

basal_area_burned_area_year300 |>
  group_by(treatment, replicate) |>
  summarise(max_ru = max(ru),
            min_ru = min(ru))

```


```{r}
#| echo: false
#| warning: false

basal_area_burned_area_year300 |>
  group_by(treatment, replicate) |>
  summarize(n = n())

# the original script formats year 0 and 300 as columns
# columns does not work very well with multiple replicates
# but to maintain the identical workflow, we need to filter RUs by total.density.ad.sap
# the following code temporarily joins year 0 data to filter RUs by total.density.ad.sap from year 300
# Year 300 is then dopped and calculated separately below
year_0 <- left_join(basal_area_burned_area_year300, final_init, by = "ru") |>
  filter(total.density.ad.sap > 50) |>
  tibble() |>
  mutate(
    dom.Pima_0 = as.factor(ifelse(init_species==1, "Dominant","Absent")),
    dom.Pigl_0 = as.factor(ifelse(init_species==2, "Dominant","Absent")),
    dom.Potr_0 = as.factor(ifelse(init_species==3, "Dominant","Absent")),
    dom.Bene_0 = as.factor(ifelse(init_species==4, "Dominant","Absent")),
    dom.Mixed_0 = as.factor(ifelse(init_species==5, "Dominant","Absent")),
    forest_present_0 = as.factor(ifelse(init_species==1|init_species==2|init_species==3|init_species==4|init_species == 5, "Forested", "Absent")))

year_0 |>
  group_by(treatment, replicate) |>
  summarize(n = n())

year_0 |>
  group_by(treatment, replicate) |>
  select(treatment, replicate, forest_present_0) |>
  filter(forest_present_0 == "Forested") |>
  summarize(n = n()) |>
  print(n = 21)

year_0 |>
  group_by(treatment, replicate) |>
  filter(forest_present_0 == "Absent") |>
  summarize(n = n()) |>
  print(n = 21)

year_0_perc <- year_0 |>
  pivot_longer(
    cols = c(dom.Pima_0, dom.Pigl_0, dom.Potr_0,
             dom.Bene_0, dom.Mixed_0, forest_present_0),
    names_to  = "variable", values_to = "state" ) |>
  count(treatment, replicate, variable, state) |>
  group_by(treatment, replicate, variable) |>
  mutate(perc = n / max(basal_area_burned_area_year300$ru) * 100) |>
  ungroup()


year_0_perc |>
  filter(state == "Dominant")


```




```{r}
#| echo: false
#| warning: false

year_300 <- left_join(basal_area_burned_area_year300, final_init, by="ru") |>
  filter(total.density.ad.sap > 50) |>
  (\(x) replace(x, is.na(x), 0))() |>
  mutate(
   dom_Pima_300 = as_factor(ifelse(IV.ad.sap_Pima > 1, "Dominant","Absent")),
   dom_Pigl_300 = as_factor(ifelse(IV.ad.sap_Pigl > 1, "Dominant","Absent")),
   dom_Potr_300 = as_factor(ifelse(IV.ad.sap_Potr > 1, "Dominant","Absent")),
   dom_Bene_300 = as_factor(ifelse(IV.ad.sap_Bene > 1, "Dominant","Absent")),
   dom_mixed_spruce_300 = as_factor(ifelse(dom_Pima_300 =="Absent" & dom_Pigl_300 == "Absent" & dom_Potr_300 =="Absent" & dom_Bene_300 == "Absent" & IV.ad.sap_Pima+IV.ad.sap_Pigl > 1,"Dominant","Absent")),
   dom_mixed_deciduous_300 = as_factor(ifelse(dom_Pima_300 =="Absent" & dom_Pigl_300 == "Absent" & dom_Potr_300 =="Absent" & dom_Bene_300 == "Absent" & IV.ad.sap_Bene + IV.ad.sap_Potr > 1,"Dominant","Absent")),
   forest_present_300 = as_factor(ifelse(count.total.ad.sap_Pima > 50 |count.total.ad.sap_Pigl > 50 |count.total.ad.sap_Potr > 50 | count.total.ad.sap_Bene > 50, "Forested", "Non forested")))

year_300 |>
  group_by(treatment, replicate) |>
  summarize(n = n())

year_300_perc <- year_300 |>
 group_by(treatment, replicate, year) |>
 select(year, replicate, treatment,
        dom_Pima_300, dom_Pigl_300, dom_Potr_300, dom_Bene_300,
        dom_mixed_spruce_300, dom_mixed_deciduous_300, forest_present_300) |>
 pivot_longer(
    cols = where(is.factor),
    names_to  = "variable",
    values_to = "state"
  ) |>
  count(treatment, replicate, variable, state) |>
  group_by(treatment, replicate, variable) |>
  mutate(perc = n / max(basal_area_burned_area_year300$ru) * 100) |>
  arrange(treatment, replicate, variable) |>
  ungroup()

treatment_labs <- c(
  # "NorEsm2-MMssp126V2_species_param_basefecund_0.5_ep1.8_dbh2.5" =
  #   "NorEsm\nε = 1.8",

  "NorEsm2-MMssp126V2_species_param_basefecund_0.5_ep2.7_dbh2.5" =
    "NorEsm\nε = 2.7",

  # "TaiESM1ssp126V2_species_param_basefecund_0.5_ep2.7_dbh2.5" =
  #   "TaiESM\nε = 2.7",

  "NorEsm2-MMssp126V2_species_param_basefecund_pigl0.35_ep2.7_dbh2.5" =
    "NorEsm\npigl = 0.35\nε = 2.7",

  "NorEsm2-MMssp126V2_species_param_pa01pl035be05pr07_ep2.7_dbh2.5" =
    "NorEsm\npa01 | pl035 | be05 | pr07\nε = 2.7",

  "NorEsm2-MMssp126V2_species_param_pa01pl035be05pr07_ep2.7_dbh2.5_shift" =
    "NorEsm\npa01 | pl035 | be05 | pr07\nε = 2.7 | shift",

  "NorEsm2-MMssp126V3_species_param_pa01pl035be05pr07_ep2.7_dbh2.5_filt" =
    "NorEsm\npa01 | pl035 | be05 | pr07\nε = 2.7 | filt",

  "NorEsm2-MMssp126V3_species_param_pa01pl035be05pr07_ep2.7_dbh2.5_filt_1980" =
    "NorEsm\npa01 | pl035 | be05 | pr07\nε = 2.7 | filt 1980"
)
```


# Percentage forest cover for years 0 and 300

The code has been converted from Winslow's original script to handle the multiple replicates. I have varified that the results are equivalant to the original code.

- The script filters for burned area (i.e., not landscape wide percentage dominance).
- Original script calculates percentage from the total land area not the burned area (`t=length(which(var=="Dominant"))/60826*100`). I have followed the same code (max ru differs by 1 for me)


### Year 0

Results are comparable to POLE-FM. Slight differences occur from different fire regime per replicate but they seem negligible.

```{r}
#| echo: false

library(DT)

year_0_perc_tbl <- year_0_perc |>
  ungroup() |>
  select(treatment, replicate, state, variable, perc) |>
  filter(treatment %in% names(treatment_labs)) |>
  pivot_wider(values_from = perc, names_from = treatment)

year_0_perc_tbl |>
  filter(state == "Dominant") |>
  select(replicate, state, variable, all_of(names(treatment_labs))) |>
  DT::datatable(
    rownames = FALSE, colnames = c("replicate", "state", "variable", unname(treatment_labs)
    )) |>
  DT::formatRound(c(4:9), 3)

```


<br>


```{r}
#| echo: false
#| eval: false

year_0_perc_tbl |>
  select(replicate, state, variable, all_of(names(treatment_labs))) |>
  DT::datatable(
    rownames = FALSE, colnames = c(
      "replicate", "state", "variable", unname(treatment_labs))) |>
  DT::formatRound(c(4:9), 3)
```

<br>

The starting forested area also is comparable to POLE-FM.

<br>

```{r}
#| echo: false

year_0 |>
  group_by(treatment, replicate) |>
  count(treatment, replicate, forest_present_0) |>
  filter(forest_present_0 == "Forested") |>
  pivot_wider(values_from = n, names_from = treatment) |>
  select(replicate, forest_present_0, all_of(names(treatment_labs))) |>
  DT::datatable(
    rownames = FALSE, colnames = c("replicate", "forest_present_0", unname(treatment_labs)))

```

### Year 300

This table is already filtered to dominant species. You can use the search box to filter to species. Black spruce seems pretty stocastic! Reps 1 and 4 were the likely best matches in the fire regime.

```{r}
#| echo: false

year_300_perc_tbl <- year_300_perc |>
  ungroup() |>
  select(treatment, replicate, state, variable, perc) |>
  pivot_wider(values_from = perc, names_from = treatment)


year_300_perc_tbl |>
  filter(state == "Dominant") |>
  select(replicate, state, variable, all_of(names(treatment_labs))) |>
  DT::datatable(
    rownames = FALSE, colnames = c("replicate", "state", "variable", unname(treatment_labs))) |>
  DT::formatRound(c(4:9), digits = 2)


```

<br>

<br>

```{r}
#| echo: false
#| eval: false

year_300_perc_tbl |>
  select(replicate, state, variable, all_of(names(treatment_labs))) |>
  DT::datatable(
    rownames = FALSE, colnames = c("replicate", "state", "variable", unname(treatment_labs))) |>
  DT::formatRound(c(4:9), digits = 2)
```




```{r}
#| echo: false
# treatment_labs <- c(
#   "NorEsm2-MMssp126V2_species_param_basefecund_0.5_ep1.8_dbh2.5" =
#     "NorEsm\nε = 1.8",

#   "NorEsm2-MMssp126V2_species_param_basefecund_0.5_ep2.7_dbh2.5" =
#     "NorEsm\nε = 2.7",

#   "TaiESM1ssp126V2_species_param_basefecund_0.5_ep2.7_dbh2.5" =
#     "TaiESM\nε = 2.7",

#   "NorEsm2-MMssp126V2_species_param_basefecund_pigl0.35_ep2.7_dbh2.5" =
#     "NorEsm\npigl = 0.35\nε = 2.7",

#   "NorEsm2-MMssp126V2_species_param_pa01pl035be05pr07_ep2.7_dbh2.5" =
#     "NorEsm\npa01 pl035 be05 pr07\nε = 2.7",

#   "NorEsm2-MMssp126V2_species_param_pa01pl035be05pr07_ep2.7_dbh2.5_shift" =
#     "NorEsm\npa01 pl035 be05 pr07\nε = 2.7 shift"
# )

year_300_0 <- year_0_perc |>
  mutate(year = 0) |>
  bind_rows(year_300_perc) |>
  mutate(
    variable = str_remove(variable, "_[0-9]+$"),
    variable = str_replace(variable, "\\.", "_"),
    year = factor(year, levels = c(0, 300)),
    treatment = factor(treatment, levels = names(treatment_labs))
  ) |>
  filter(state == "Dominant",
         !grepl("Mixed|mixed", variable),
         treatment %in% names(treatment_labs))

colours <- c(
  "dom_Pima" = "#8c564b", "dom_Pigl" = "#1f77b4",
  "dom_Potr" = "#2ca02c", "dom_Bene" = "green")


```




```{r}
#| echo: false
ggplot(year_300_0,
  aes(x = year, y = perc, colour = variable, group = replicate)) +
  geom_point() +
  geom_line() +
  facet_grid(variable ~ treatment,
             labeller = labeller(treatment = as_labeller(treatment_labs))) +
  scale_colour_manual(values = colours) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    legend.position = "bottom"
  )

# year_300_0 |>
#   filter(treatment == "NorEsm pa01 pl035 be05 pr07 Epsilon 2.7 shift",
#          variable == "dom_Potr") |>
#   group_by(treatment, replicate, variable, state, year) |>
#   summarise(mean_n = mean(n),
#             mean_perc = mean(perc))
```


This one plots the treatments on top of each other. Doesn't look pretty, but shows the overall grouping for different treatments overlaid.

```{r}
#| echo: false

ggplot(year_300_0,
  aes(x = year, y = perc, colour = treatment, group = interaction(replicate, treatment))) +
  geom_point() +
  geom_line() +
  facet_wrap(~variable) +
  scale_colour_discrete(
    breaks = names(treatment_labs),
    labels = unname(treatment_labs)
  ) +  
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    legend.position = "bottom"
  )

```



## FISL meeting 2026/01/28 {.unlisted .hidden .unnumbered}


```{r}
#| echo: false
#| eval: false
year_300_0_fisl <- year_0_perc |>
  mutate(year = 0) |>
  bind_rows(year_300_perc) |>
  filter(grepl("NorEsm", treatment),
         grepl("2.7", treatment),
         state == "Dominant",
         !grepl("Mixed|mixed", variable)) |>
  mutate(
    treatment = case_when(
         treatment == "NorEsm2-MMssp126V2_species_param_basefecund_0.5_ep2.7_dbh2.5" ~ "Pima 0.15 | Pigl 0.25 | Bene 0.5 | Potr 0.5",
         TRUE ~ treatment),
    treatment = case_when(
         treatment == "NorEsm2-MMssp126V2_species_param_basefecund_pigl0.35_ep2.7_dbh2.5" ~ "Pima 0.15 | Pigl 0.35 | Bene 0.5 | Potr 0.5",
         TRUE ~ treatment),
    treatment = case_when(
         treatment == "NorEsm2-MMssp126V2_species_param_pa01pl035be05pr07_ep2.7_dbh2.5" ~ "Pima 0.1 | Pigl 0.35 | Bene 0.5 | Potr 0.7",
         TRUE ~ treatment),
    variable = str_remove(variable, "_[0-9]+$"),
    variable = str_replace(variable, "\\.", "_"),
    year = factor(year, levels = c(0, 300)),
    treatment = factor(treatment, levels = c(
      "Pima 0.15 | Pigl 0.25 | Bene 0.5 | Potr 0.5",
      "Pima 0.15 | Pigl 0.35 | Bene 0.5 | Potr 0.5",
      "Pima 0.1 | Pigl 0.35 | Bene 0.5 | Potr 0.7"
    ))
  ) |>
  filter(state == "Dominant",
         !grepl("Mixed|mixed", variable))


colours <- c(
  "dom_Pima" = "#8c564b", "dom_Pigl" = "#1f77b4",
  "dom_Potr" = "#2ca02c", "dom_Bene" = "green")


```

```{r}
#| echo: false
#| eval: false

labels <- c(
  dom_Pima = "Black spruce",
  dom_Pigl = "White spruce",
  dom_Bene = "Birch",
  dom_Potr = "Aspen"
)

pima_plot <- ggplot(
  year_300_0_fisl |>
    filter(variable == "dom_Pima"),
  aes(x = year, y = perc, colour = variable, group = replicate)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 40, colour = "red") +
  facet_grid(variable ~ treatment,
             labeller = labeller(variable = labels)) +
  scale_colour_manual(values = colours) +
  labs(y = "Percent dominated (post-fire area)",
       x = "Year", colour = "Species") +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    legend.position = "none",
    strip.text = element_text(size = 12)
  )



images_out <- "./images/prop_dom"
dir.create(images_out)

ggsave(
  file.path(images_out, "pima_plot.png"),
  pima_plot,
  device = "png",
  dpi = 300,
  width = 12,
  height = 4,
  units = "in"
)  
```


```{r}
#| echo: false
#| eval: false
birch_aspen_plot <- ggplot(
  year_300_0_fisl |>
    filter(variable == "dom_Bene" | variable == "dom_Potr"),
  aes(x = year, y = perc, colour = variable, group = replicate)) +
  geom_point() +
  geom_line() +
  facet_grid(variable ~ treatment,
             labeller = labeller(variable = labels)) +
  scale_colour_manual(values = colours) +
  labs(y = "Percent dominated (post-fire area)",
       x = "Year", colour = "Species") +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    legend.position = "none",
    strip.text = element_text(size = 12)
  )

ggsave(
  file.path(images_out, "birch_aspen_plot.png"),
  birch_aspen_plot,
  device = "png",
  dpi = 300,
  width = 12,
  height = 8,
  units = "in"
)   
```



```{r}
#| echo: false
#| eval: false
wspruce_aspen_plot <- ggplot(
  year_300_0_fisl |>
    filter(variable == "dom_Pigl" | variable == "dom_Potr"),
  aes(x = year, y = perc, colour = variable, group = replicate)) +
  geom_point() +
  geom_line() +
  facet_grid(variable ~ treatment,
             labeller = labeller(variable = labels)) +
  scale_colour_manual(values = colours) +
  labs(y = "Percent dominated (post-fire area)",
       x = "Year", colour = "Species") +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    legend.position = "bottom",
    strip.text = element_text(size = 12)
  )

ggsave(
  file.path(images_out, "wspruce_aspen_plot.png"),
  wspruce_aspen_plot,
  device = "png",
  dpi = 300,
  width = 12,
  height = 8,
  units = "in"
)    
```









## Winslows original code {.unlisted .hidden .unnumbered}

This is winslows original code wrapped in a split and lapply to work over replicates. I have varified that values are the same as my adjusted code above.

```{r}
#| eval: false

map <- terra::rast("Z:/project_data/na_boreal/sensitivity_analysis/data/output/spinup_summaries/fire_regime/NorEsm2-MMssp126V2_species_param_basefecund_0.5_ep2.7_dbh2.5/rep_1/spinup_300.asc")
map
terra::app(map,max)
map_df <- as.data.frame(map, xy=T)
map_df <- map_df |>
  rename(ru = spinup_300)

coord <- map_df |> 
  dplyr::select(x,y)

init_species <- terra::rast("Z:/project_data/na_boreal/sensitivity_analysis/data/output/spinup_summaries/forest_species_init.tif")
#plot(init_species)
data <- data.frame(cbind(map_df, extract(init_species, coord)))
final_init <- data |>
  rename("init_species" = forest_species_init)


stand.t.wide <- left_join(basal_area_burned_area_year300, final_init, by="ru")

mapping.data <- split(stand.t.wide, list(stand.t.wide$treatment, stand.t.wide$replicate)) |>
 lapply(\(x) {
  stand.t.wider = x %>% pivot_wider(names_from = year, values_from = c(count.total.ad.sap_Pima:stand.age))
  stand.t.wider =stand.t.wider %>% replace(is.na(.), 0)
  stand.t.wider = stand.t.wider %>% mutate(init_species.map = ifelse(init_species==1, "Pima",
                                                                     ifelse(init_species==2, "Pigl",
                                                                            ifelse(init_species==3, "Potr",
                                                                                   ifelse(init_species==4,"Bene",
                                                                                          ifelse(init_species==5, "Mixed","Not forested"))))))
  stand.t.wider= stand.t.wider %>%filter(total.density.ad.sap_300>50)
  stand.t.wider= stand.t.wider %>% mutate(dom.Pima_0 = as.factor(ifelse(init_species==1, "Dominant","Absent")),
                                          dom.Pigl_0 = as.factor(ifelse(init_species==2, "Dominant","Absent")),
                                          dom.Potr_0 = as.factor(ifelse(init_species==3, "Dominant","Absent")),
                                          dom.Bene_0 =as.factor(ifelse(init_species==4, "Dominant","Absent")),
                                          dom.Mixed_0 =as.factor(ifelse(init_species==5, "Dominant","Absent")),
                                          forest_present_0 = as.factor(ifelse(init_species==1|init_species==2|init_species==3|init_species==4|init_species==5, "Forested", "Absent")),
                                          dom_Pima_300 = as_factor(ifelse(IV.ad.sap_Pima_300 > 1, "Dominant","Absent")),
                                          dom_Pigl_300 = as_factor(ifelse(IV.ad.sap_Pigl_300 > 1, "Dominant","Absent")),
                                          dom_Potr_300 = as_factor(ifelse(IV.ad.sap_Potr_300 > 1, "Dominant","Absent")),
                                          dom_Bene_300 = as_factor(ifelse(IV.ad.sap_Bene_300 > 1, "Dominant","Absent")),
                                          dom_mixed_spruce_300 = as_factor(ifelse(dom_Pima_300 =="Absent" &dom_Pigl_300=="Absent" & dom_Potr_300=="Absent" &dom_Bene_300=="Absent" & IV.ad.sap_Pima_300+IV.ad.sap_Pigl_300 >1,"Dominant","Absent")),
                                          dom_mixed_deciduous_300 = as_factor(ifelse(dom_Pima_300=="Absent" &dom_Pigl_300=="Absent" & dom_Potr_300=="Absent" &dom_Bene_300=="Absent" &IV.ad.sap_Bene_300 + IV.ad.sap_Potr_300 >1,"Dominant","Absent")),
                                          forest_present_300 = as_factor(ifelse(count.total.ad.sap_Pima_300 >50 |count.total.ad.sap_Pigl_300 >50 |count.total.ad.sap_Potr_300 >50 | count.total.ad.sap_Bene_300 >50, "Forested", "Non forested")))
  
  stand.t.wider= stand.t.wider %>% mutate(Pima.transition = ifelse(dom.Pima_0=="Dominant"& dom_Pima_300 =="Dominant", "Remained dominant",
                                                                    ifelse(dom.Pima_0=="Dominant"&forest_present_300=="Non forested", "Now non forested",
                                                                   ifelse(dom.Pima_0=="Dominant"& dom_Potr_300=="Dominant", "Now Potr",
                                                                          ifelse(dom.Pima_0=="Dominant"& dom_Pigl_300=="Dominant", "Now Pigl",
                                                                                 ifelse(dom.Pima_0=="Dominant"& dom_Bene_300=="Dominant", "Now Bene",
                                                                                        ifelse(dom.Pima_0=="Dominant" & dom_mixed_spruce_300=="Dominant", "Now mixed spruce",
                                                                                               ifelse(dom.Pima_0=="Dominant" & dom_mixed_deciduous_300=="Dominant", "Now mixed deciduous",
                                                                                               ifelse(dom.Pima_0=="Absent"& dom_Pima_300 =="Dominant", "Now dominant","Never there")))))))))
  
  stand.t.wider= stand.t.wider %>% mutate(Pigl.transition = ifelse(dom.Pigl_0=="Dominant"& dom_Pigl_300 =="Dominant", "Remained dominant",
                                                                    ifelse(dom.Pigl_0=="Dominant"&forest_present_300=="Non forested", "Now non forested",
                                                                   ifelse(dom.Pigl_0=="Dominant"& dom_Potr_300=="Dominant", "Now Potr",
                                                                          ifelse(dom.Pigl_0=="Dominant"& dom_Pima_300=="Dominant", "Now Pima",
                                                                                 ifelse(dom.Pigl_0=="Dominant"& dom_Bene_300=="Dominant", "Now Bene",
                                                                                        ifelse(dom.Pigl_0=="Dominant" & dom_mixed_spruce_300=="Dominant", "Now mixed spruce",
                                                                                               ifelse(dom.Pigl_0=="Dominant" & dom_mixed_deciduous_300=="Dominant", "Now mixed deciduous",
                                                                                               ifelse(dom.Pigl_0=="Absent"& dom_Pigl_300 =="Dominant", "Now dominant","Never there")))))))))
  
  
  stand.t.wider= stand.t.wider %>% mutate(Potr.transition = ifelse(dom.Potr_0=="Dominant"& dom_Potr_300 =="Dominant", "Remained dominant",
                                                                    ifelse(dom.Potr_0=="Dominant"&forest_present_300=="Non forested", "Now non forested",
                                                                   ifelse(dom.Potr_0=="Dominant"& dom_Pigl_300=="Dominant", "Now Pigl",
                                                                          ifelse(dom.Potr_0=="Dominant"& dom_Pima_300=="Dominant", "Now Pima",
                                                                                 ifelse(dom.Potr_0=="Dominant"& dom_Bene_300=="Dominant", "Now Bene",
                                                                                        ifelse(dom.Potr_0=="Dominant" & dom_mixed_spruce_300=="Dominant", "Now mixed spruce",
                                                                                               ifelse(dom.Potr_0=="Dominant" & dom_mixed_deciduous_300=="Dominant", "Now mixed deciduous",
                                                                                               ifelse(dom.Potr_0=="Absent"& dom_Potr_300 =="Dominant", "Now dominant","Never there")))))))))
  stand.t.wider= stand.t.wider %>% mutate(Bene.transition = ifelse(dom.Bene_0=="Dominant"& dom_Bene_300 =="Dominant", "Remained dominant",
                                                                    ifelse(dom.Bene_0=="Dominant"&forest_present_300=="Non forested", "Now non forested",
                                                                   ifelse(dom.Bene_0=="Dominant"& dom_Pigl_300=="Dominant", "Now Pigl",
                                                                          ifelse(dom.Bene_0=="Dominant"& dom_Pima_300=="Dominant", "Now Pima",
                                                                                 ifelse(dom.Bene_0=="Dominant"& dom_Potr_300=="Dominant", "Now Potr",
                                                                                        ifelse(dom.Bene_0=="Dominant" & dom_mixed_spruce_300=="Dominant", "Now mixed spruce",
                                                                                               ifelse(dom.Bene_0=="Dominant" & dom_mixed_deciduous_300=="Dominant", "Now mixed deciduous",
                                                                                               ifelse(dom.Bene_0=="Absent"& dom_Bene_300 =="Dominant", "Now dominant","Never there")))))))))

  stand.t.wider$Pima.transition <- factor(stand.t.wider$Pima.transition, levels = c("Remained dominant", "Now dominant" , "Now Pima","Now Pigl","Now Potr", "Now Bene", "Now mixed spruce", "Now mixed deciduous", "Never there"))
  return(stand.t.wider)
})


length(which(mapping.data[[1]]$forest_present_0 == "Forested"))
length(which(mapping.data[[2]]$forest_present_0 == "Forested"))
length(which(mapping.data[[1]]$forest_present_0 == "Absent"))
length(which(mapping.data[[4]]$forest_present_0 == "Absent"))

```



```{r}
#| eval: false

d <- lapply(mapping.data, \(x) {
  x %>% dplyr::select(dom.Pima_0,dom_Pima_300,dom.Pigl_0,dom_Pigl_300,dom.Potr_0,dom_Potr_300,dom.Bene_0,dom_Bene_300,dom.Mixed_0,dom_mixed_spruce_300,dom_mixed_deciduous_300)
})

d <- d[[1]]

calc.percentage =function(var,species,label){
  t=length(which(var=="Dominant"))/60826*100
  print(paste0("Species ",species," year ",label, " percentage = ",round(t,digits=1)))
}
length(which(d$dom.Pigl_0=="Dominant"|d$dom.Pima_0=="Dominant"))
length(which(d$dom_Pigl_300=="Dominant"|d$dom_Pima_300=="Dominant"|d$dom_mixed_spruce_300=="Dominant"))

length(which(d$dom.Potr_0=="Dominant"|d$dom.Bene_0=="Dominant"|d$dom.Mixed_0=="Dominant"))
length(which(d$dom_Potr_300=="Dominant"|d$dom_Bene_300=="Dominant"|d$dom_mixed_deciduous_300=="Dominant"))

length(which(d$dom.Pima_0=="Dominant"|d$dom.Pigl_0=="Dominant"|d$dom.Bene_0=="Dominant"|d$dom.Potr_0=="Dominant"|d$dom.Mixed_0=="Dominant"))
length(which(d$dom_Pima_300=="Dominant"|d$dom_Pigl_300=="Dominant"|d$dom_Bene_300=="Dominant"|d$dom_Potr_300=="Dominant"|d$dom_mixed_spruce_300=="Dominant"|d$dom_mixed_deciduous_300=="Dominant"))
length(which(mapping.data$total.density.ad.sap_300>50))

calc.percentage(d$dom.Pima_0,"Pima","0")
calc.percentage(d$dom.Pigl_0,"Pigl","0")
calc.percentage(d$dom.Bene_0,"Bene","0")
calc.percentage(d$dom.Potr_0,"Potr","0")
calc.percentage(d$dom.Mixed_0,"Mixed","0")

calc.percentage(d$dom_Pima_300,"Pima","300")
calc.percentage(d$dom_Pigl_300,"Pigl","300")
calc.percentage(d$dom_Bene_300,"Bene","300")
calc.percentage(d$dom_Potr_300,"Potr","300")
calc.percentage(d$dom_mixed_spruce_300,"Mixed_spruce","300")
calc.percentage(d$dom_mixed_deciduous_300,"Mixed_deciduous","300")
```


