---
title: "Water and daily response debug"
author: "FFL"
date: today
---

# Intro

## Data {.unnumbered .unlisted .hidden}

```{r}
#| label: data processing
#| echo: false
#| warning: false
source("./scripts/install.R")
pak::pak("data.table")
library(data.table)

out_dir  <- here("data", "debug")
out_file_water <- here(out_dir, "debug_water.parquet")
out_file_daily_resp <- here(out_dir, "debug_daily_resp.parquet")
in_dir   <- "//10.60.2.10/FF_Lab/project_data/na_boreal/sensitivity_analysis/data/output/debug_data/daily_response/temp"
if(!dir.exists(in_dir)) {
  stop("Are you connected to Janet?")
}

dir.create(out_dir, recursive = TRUE)

if (!file.exists(out_file_water)) {
  water_debug <- read_delim(file.path(in_dir, "debug_water_cycle.csv"), delim = ";")

  water_debug <- water_debug |>
    select(id, rid, year, content, psi_kpa, combined_response)

  arrow::write_parquet(water_debug, out_file_water)

  } else {
    # read_parquet is not lazy and will return a tibble
    # open_dataset is lazy and will need collect()
    # Use open_dataset if RAM is an issue, but the files are not that big
    water_debug <- arrow::read_parquet(out_file_water)
}


if (!file.exists(out_file_daily_resp)) {
  daily_resp <- read_delim(file.path(in_dir, "debug_daily_responses.csv"), delim = ";")

  daily_resp <- daily_resp |>
    select(date, year, species, rid, waterResponse, tempResponse, VpdResponse, util.Radiation)

  arrow::write_parquet(daily_resp, out_file_daily_resp)

  # } else {
  # daily_resp <- arrow::read_parquet(out_file_daily_resp)
}

```

## Water cycle


```{r}
#| label: dplyr wrangling
#| eval: true
#| echo: false
#| output: false

format(object.size(water_debug), units = "Mb")
#"5078.7 Mb"

water_debug_long <- water_debug |>
  pivot_longer(-c(id, rid, year))

format(object.size(water_debug_long), units = "Mb")
# "12696.8 Mb"

rm(water_debug)

water_debug_ridyear <- water_debug_long |>
  group_by(year, rid, name) |>
  summarise(mean_val = mean(value),
  sd_val = sd(value))

format(object.size(water_debug_ridyear), units = "Mb")
# "60.3 Mb"
```



```{r}
#| label: data.table wrangling
#| eval: false
#| echo: false
#| output: false
water_debug <- as.data.table(water_debug)

water_debug_long <- melt(water_debug,
               id.vars = c("id", "rid", "year"),
               variable.name = "name",
               value.name = "value")

format(object.size(water_debug_long), units = "Mb")


water_debug_ridyear <- water_debug_long[, .(
  mean_val = mean(value, na.rm = TRUE),
  sd_val   = sd(value, na.rm = TRUE)
), by = .(year, rid, name)]
format(object.size(water_debug_ridyear), units = "Mb")

```

Distributions from daily data. From iLand wiki:

- `psi_kpa`: Soil water potential (kPa) at the beginning of the day, before evapotranspiration.
- `content`: Current soil water content at the end of the day, after evapotranspiration (mm).
- `combined_response`: Combined daily response modifier (0-1) of all species on the RU, representing the minimum of the water and VPD responses.

This is plotting all rids for all 5 years. I can do one histogram per year.


```{r}
#| label: plot daily (takes a while)
#| eval: true
#| warning: false
#| fig-height: 10
head(water_debug_long)

ggplot(water_debug_long, aes(x = value)) +
  geom_histogram() +
  facet_wrap(~name, scales = "free", ncol = 1) +
  theme_minimal()


```


Distributions from yearly average:

```{r}
#| label: Plot yearly
#| warning: false
#| fig-height: 10
head(water_debug_ridyear)

ggplot(water_debug_ridyear, aes(x = mean_val)) +
  geom_histogram() +
  facet_wrap(~name, scales = "free", ncol = 1) +
  theme_minimal()

```


## Daily response

```{r}
#| label: starting daily resp
#| echo: false
#| output: false
rm(water_debug_long, water_debug_ridyear)

daily_resp <- arrow::open_dataset(out_file_daily_resp)

format(object.size(daily_resp), units = "Mb")

```

```{r}
#| label: Making plots
#| echo: false
#| output: false
#| warning: false
vars <- c("waterResponse", "tempResponse", "VpdResponse", "util.Radiation")

daily_resp_plot_list <- lapply(vars, function(var) {
  df <- daily_resp |>
    select(all_of(var), species) |>
#    slice_sample(n = 1e+05) |>
    collect()

  plt <- ggplot(df, aes(x = .data[[var]])) +
    geom_histogram() +
    facet_wrap(~species) +
    theme_minimal()

  rm(df)
  return(plt)
})


```

Histograms of daily waterResponse, tempResponse, VpdResponse, util.Radiation

From iLand wiki:

- `resp_soil`: Monthly average of the daily soil water response modifier (0-1). A value of 1 indicates no water limitation.
- `resp_temp`: Monthly average of the daily temperature response modifier (0-1). A value of 1 indicates optimal temperature for growth.
- `resp_vpd`: Monthly average of the daily Vapour Pressure Deficit (VPD) response modifier (0-1). A value of 1 indicates no limitation from atmospheric drought.
- `utilizable_rad`: The utilizable radiation for the month (MJ/m2). This is the APAR value after being reduced by all preceding response modifiers (meteo, pheno, co2, daylength).

These are distributions for all rids and all 5 years. I can plot histograms per year, or do yearly average like above.

:::{.panel-tabset}

```{r}
#| label: LONG time to process
#| results: asis
#| warning: false

for (i in seq_along(daily_resp_plot_list)) {
  cat("# ",vars[i],"\n")
  print(daily_resp_plot_list[[i]])
  cat('\n\n')
}

```

:::


Daily response with 0 filtered out


```{r}
#|label: Making plots without 0
#| echo: false
#| output: false
vars <- c("waterResponse", "tempResponse", "VpdResponse", "util.Radiation")

daily_resp_plot_list <- lapply(vars, function(var) {
  df <- daily_resp |>
    select(all_of(var), species) |>
    filter(.data[[var]] != 0) |>
#    slice_sample(n = 1e+05) |>
    collect()

  plt <- ggplot(df, aes(x = .data[[var]])) +
    geom_histogram() +
    facet_wrap(~species) +
    theme_minimal()

  rm(df)
  return(plt)
})


```


:::{.panel-tabset}

```{r}
#| label: time to process...
#| results: asis
#| warning: false

for (i in seq_along(daily_resp_plot_list)) {
  cat("# ",vars[i],"\n")
  print(daily_resp_plot_list[[i]])
  cat('\n\n')
}

```

:::
