---
title: "Carbon parameterization"
author: "Winslow Hansen"
date: "11/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
library(terra)
v=21
```
Map raster to pull out the initial species distribution
```{r}
map = rast("D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/cpcrw/model/temp/spinup_300.asc")
map
app(map,max)
map.df=as.data.frame(map, xy=T)
map.df = map.df %>% rename(ru = spinup_300)

init.species = rast("D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/cpcrw/materials/init/forest_species_init.tif")
#plot(init.species)
coord=map.df%>%dplyr::select(x,y)
data=data.frame(cbind(map.df, extract(init.species, coord)))
final.init=data%>%rename("init.species"=forest_species_init)
```

Read in the produce a stand composition and regen dataset to check stand densities at the beginning and end of the simulation and then pair pare with the simulated carbon dataset
```{r}
db.conn <- DBI::dbConnect(RSQLite::SQLite(), dbname="D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/cpcrw/model/output/paper_run/CPCRW_sm_6.sqlite")
stand = tbl(db.conn, "stand")  %>% filter(year ==300) %>% collect()
reg.det = tbl(db.conn, "saplingdetail") %>% filter(year ==300) %>% collect()
fire =  tbl(db.conn, "fire") %>% collect()
  DBI::dbDisconnect(db.conn)
```





The first task is to compare stand density and basal area of different stands dominated by different species to RSN and CAFI benchmarks
First we need to prep the stand data to be comparable



Processing regeneration to bind with stand
```{r}
reg.det=reg.det%>%mutate(ba=pi*((dbh/100)/2)^2)
reg.det=reg.det%>%mutate(ba_all=ba*n_represented)
reg.det.sum=reg.det%>%group_by(year,ru, species)%>%summarize(count_ha_sap=sum(n_represented),
                                                             dbh_mean_sapling=weighted.mean(dbh,n_represented),
                                                             ba_sum_sapling=sum(ba_all))

summary(reg.det.sum)
```

Merging stand and regenration together to calculate total density.

```{r}
stand = stand %>% dplyr::select(-cohort_count_ha,-cohort_basal_area)
stand.t=full_join(stand,reg.det.sum,by = c("ru","year","species"))
stand.t2 =stand.t %>% replace(is.na(.), 0)
stand.t2=stand.t2%>%mutate(count.total.ad.sap = count_ha + count_ha_sap,
                         ba.total.ad.sap = basal_area_m2+ba_sum_sapling)
```
Now pivoting wider to generate IV values
```{r}
stand.t2.select = stand.t2 %>% dplyr::select(ru,rid, year,species, count.total.ad.sap:ba.total.ad.sap)
stand.t.wide = stand.t2.select %>% pivot_wider( names_from = species, values_from = c(count.total.ad.sap:ba.total.ad.sap), values_fill = 0)
stand.t.wide=stand.t.wide %>% mutate(total.density.ad.sap = count.total.ad.sap_Potr + count.total.ad.sap_Pima + count.total.ad.sap_Bene + count.total.ad.sap_Pigl,
                                     total.ba.ad.sap = ba.total.ad.sap_Potr + ba.total.ad.sap_Pima + ba.total.ad.sap_Bene + ba.total.ad.sap_Pigl)
stand.t.wide=stand.t.wide%>%mutate(IV.ad.sap_Pima = ifelse(total.density.ad.sap !=0 & total.ba.ad.sap!=0, (count.total.ad.sap_Pima/total.density.ad.sap)+(ba.total.ad.sap_Pima/total.ba.ad.sap),0),
                                   IV.ad.sap_Pigl = ifelse(total.density.ad.sap !=0 & total.ba.ad.sap!=0, (count.total.ad.sap_Pigl/total.density.ad.sap)+(ba.total.ad.sap_Pigl/total.ba.ad.sap),0),
                                   IV.ad.sap_Potr = ifelse(total.density.ad.sap !=0 & total.ba.ad.sap!=0, (count.total.ad.sap_Potr/total.density.ad.sap)+(ba.total.ad.sap_Potr/total.ba.ad.sap), 0),
                                   IV.ad.sap_Bene = ifelse(total.density.ad.sap !=0 & total.ba.ad.sap!=0,(count.total.ad.sap_Bene/total.density.ad.sap)+(ba.total.ad.sap_Bene/total.ba.ad.sap), 0))
summary(stand.t.wide)
```

Approach to identify time since fire for each grid cell in stand.t.wide
```{r}
# Let's get all the spread files (fire extents) as a raster stack
path = "D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/cpcrw/model/output/paper_run/spread_6"
fire_maps = rast(list.files(path, full.names = T))

# Use the environment grid to get ids for each grid
path = "D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/cpcrw/model/gis"
env_grid <- rast(paste0(path, "/env.grid.txt"))
crs(fire_maps)=crs(env_grid)

#env_grid.2=crop(env_grid,fire_maps)
env_grid.2=env_grid
fire <- fire %>% mutate(prop.dens.killed = n_trees_died   / n_trees,
                        prop.ba.killed   = basalArea_died / basalArea_total)
fire$prop.ba.killed  [fire$basalArea_total == 0] <- 0
fire$prop.dens.killed[fire$n_trees         == 0] <- 0

fire_year_dat <- data.frame()

for (i in 1:nrow(fire)) {
  
  # Skip fires that burned nothing
  if (fire$area_m2[i] > 0) {
    
    # Take the fire map for this fire, and make everything but the fire cells NA
    mask <- fire_maps[[paste0("spread_", fire$fireId[i])]]
    mask = ifel(mask==0, NA,mask)
    # Use the mask to extract the rids for burned grids
    burned_rids <- as.data.frame(mask(env_grid.2, mask))
burned_rids= burned_rids%>%filter(!is.na(env.grid))
burned_rids$year=fire$year[i]
  }
  fire_year_dat=rbind(fire_year_dat,burned_rids)
}
fire_year_dat = fire_year_dat %>% rename("rid" = env.grid, "fire.year"=year)
fire.year_dat.f = fire_year_dat %>% group_by(rid) %>% summarize(last.fire.year = max(fire.year))
```
Join together fire year and stand wide
```{r}
summary(stand.t.wide)
stand.t.wide = left_join(stand.t.wide, fire.year_dat.f, by="rid")  
stand.t.wide =stand.t.wide %>% replace(is.na(.), 0)
stand.t.wide = stand.t.wide %>% mutate(stand.age = 300 - last.fire.year)
```








Preparing the modeled and benchmarking data for comparison with plots
```{r}
stand.t.wide.subset=stand.t.wide %>% filter(year==300 &stand.age > 34) %>% dplyr::select(count.total.ad.sap_Pima:IV.ad.sap_Bene, stand.age)  # Here we filter to stands 35 years or older to match the observations and stands 
summary(stand.t.wide.subset)
stand.t.wide.subset$stand.age=NULL
benchmark_rsn=read.table("D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/cpcrw/model/output/landscape_stand-structure-RSN-validation.txt")
summary(benchmark_rsn)
benchmark_rsn_subset = benchmark_rsn %>% dplyr::select(count.total.ad.sap_Pima:count.total.ad.sap_Potr,ba.total.ad.sap_Pima:ba.total.ad.sap_Potr, total.density.ad.sap, total.ba.ad.sap, IV.ad.sap_Pima:IV.ad.sap_Bene)
benchmark_rsn_subset = benchmark_rsn_subset %>% mutate(form = "Observed-RSN")

benchmark_cafi = read.table("D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/cpcrw/model/output/landscape_structure-CAFI-validation.txt")
summary(benchmark_cafi)
benchmark_cafi_subset = benchmark_cafi %>% dplyr::select(count.total.ad.sap_Bene, count.total.ad.sap_Pigl, count.total.ad.sap_Potr, count.total.ad.sap_Pima, ba.total.ad.sap_Bene, ba.total.ad.sap_Pigl, ba.total.ad.sap_Potr, ba.total.ad.sap_Pima, total.density.ad.sap:IV.ad.sap_Bene)
benchmark_cafi_subset = benchmark_cafi_subset %>% mutate(form = "Observed-CAFI")

benchmark = rbind(benchmark_rsn_subset, benchmark_cafi_subset)
stand.t.wide.subset=stand.t.wide.subset %>% mutate(form = "Modeled")
data=rbind(stand.t.wide.subset, benchmark)
summary(data)
data= data %>% mutate(sp.dom = as.factor(ifelse(IV.ad.sap_Pima > 1,"Pima",
                                      ifelse(IV.ad.sap_Potr > 1, "Potr",
                                             ifelse(IV.ad.sap_Pigl > 1, "Pigl",
                                                    ifelse(IV.ad.sap_Bene > 1, "Bene","no.species"))))))

data.long = data %>% pivot_longer(cols = c(count.total.ad.sap_Pima:ba.total.ad.sap_Bene,IV.ad.sap_Pima:IV.ad.sap_Bene), names_to = c(".value","species"),names_sep = "_")
summary(data.long)
library(scales)




data.long %>% filter(sp.dom!= "no.species") %>%
ggplot(aes(x=species, y=count.total.ad.sap+1, fill=form))+
  geom_boxplot()+
  scale_y_log10()+
  facet_wrap(.~sp.dom)
summary(data.long)
data.plot = data.long %>% filter(sp.dom=="Pima" & species=="Pima"|sp.dom=="Pigl"&species=="Pigl"|sp.dom=="Potr"&species=="Potr"|sp.dom=="Bene"&species=="Bene")
summary(data.plot)
d= ggplot(data.plot, aes(x=species, y=count.total.ad.sap, fill=form))+
  geom_boxplot()+
  scale_y_log10(labels=comma)+
  theme_bw()+
  scale_fill_manual(values=c("#1b7837","#762a83","#9970ab"))+
  theme(    text  = element_text(color="black",  family="sans", size=10),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
             axis.text.x=element_text(colour="black",size=10),
             axis.text.y=element_text(colour="black",size=10),
            legend.text = element_text(size=8),
             legend.position= "",
            legend.background = element_rect(colour = "transparent", fill = 'transparent'),
            legend.key = element_rect(colour = "transparent", fill = 'transparent'),
            legend.key.size = unit(0.5, "cm"))+
  labs(x="", y=bquote(Density~(Trees~ha^-1)),fill="")


ba=ggplot(data.plot, aes(x=species, y=ba.total.ad.sap, fill=form))+
  geom_boxplot()+
  scale_y_log10(labels=comma)+
  theme_bw()+
  scale_fill_manual(values=c("#1b7837","#762a83","#9970ab"))+
  theme(    text  = element_text(color="black",  family="sans", size=10),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
             axis.text.x=element_text(colour="black",size=10),
             axis.text.y=element_text(colour="black",size=10),
            legend.text = element_text(size=8),
             legend.position= c(0.75,0.1),
            legend.background = element_rect(colour = "transparent", fill = 'transparent'),
            legend.key = element_rect(colour = "transparent", fill = 'transparent'),
            legend.key.size = unit(0.5, "cm"))+
  labs(x="", y=bquote(Basal~area~(m^2~ha^-1)),fill="")
cowplot::plot_grid(d,ba,nrow=1,labels="AUTO",label_size = 10,label_fontface = "plain", label_fontfamily = "sans")
ggsave("D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/figures/stand.density.ba.jpg", height=4.5,width=6.5, units = "in", dpi=700)


data.long %>% filter(sp.dom=="Pima" & species=="Pima"|sp.dom=="Pigl" & species=="Pigl"|sp.dom=="Potr" & species=="Potr"|sp.dom=="Bene" & species=="Bene") %>% group_by(form, sp.dom) %>% summarize(quantile.density = quantile(count.total.ad.sap, c(0.25,0.5,0.75)),q=c(0.25,0.5,0.75),
                                                                                                                                                                                                   quantile.ba=quantile(ba.total.ad.sap, c(0.25,0.5,0.75)),q=c(0.25,0.5,0.75))

data.plot%>% group_by(form, sp.dom) %>% summarize(quantile.density = quantile(count.total.ad.sap, c(0.25,0.5,0.75)),q=c(0.25,0.5,0.75),
                                                                                                                                                                                                   quantile.ba=quantile(ba.total.ad.sap, c(0.25,0.5,0.75)),q=c(0.25,0.5,0.75))
                                                  
```


Now we need to plot where we have lost forest and how forest dominance varies as a function of aspect and SOL

Mapping variables
Final maps
First generate a hillshade effect
```{r}
library(terra)
env.grid = rast("D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/cpcrw/materials/env/env.grid.tif")
dem=rast("D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/raw data/DEM/dem_repr.tif")
dem=crop(dem,env.grid)
slope=terrain(dem,v="slope",unit="radians")
aspect=terrain(dem, v="aspect",unit="radians")
hill=shade(slope,aspect,40,270,normalize=T)
hill.df =as.data.frame(hill,xy=T)
colnames(hill.df) <- c( "x", "y", "value")

```
Read in water
```{r}
library(sf)
m.river=st_read("D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/raw data/hydrology/export_1594471551598/hydro_63360_py_repr_cpcrw.shp")
c.river=st_read("D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/raw data/hydrology/attachments/cpcrw_streams.shp")
m.river2=st_transform(m.river,st_crs(c.river))
```

```{r}
stand.t.wide=left_join(stand.t.wide, final.init, by="ru")
summary(stand.t.wide)


cols.1 <- c("Dominant" = "green", "Absent" = "grey")
cols.2 <- c( "Forested" = "darkgreen", "Non forested" = "grey")
cols.3 <- c("Remained dominant"="#4daf4a","Now dominant"="#a65628","Now Pima"="#ff7f00", "Now Pigl"="#984ea3","Now Potr"="#377eb8","Now Bene"="#ffff33","Now mixed spruce" ="#01665e","Now mixed deciduous" ="#542788", "Now non forested" = "#e41a1c","Never there"="grey")
cols.4 <- c("Potr"="#377eb8","Pigl"="#984ea3","Pima"="#ff7f00", "Bene"="#ffff33", "Mixed" ="#80cdc1","mixed deciduous" ="#afcd80","Not forested"="grey")

map = function(dataset, variable, year.label, col){
plot=ggplot(dataset)+
  geom_tile(data=hill.df, aes(x=x, y=y, fill=value),show.legend = F)+
  scale_fill_gradient(low = "grey0", high = "grey100") +
  ggnewscale::new_scale_fill() +
  geom_tile(aes_string(x="x", y="y", fill=variable), alpha=0.6)+
  scale_fill_manual(values = col)+
ggthemes::theme_map() +
guides(fill=guide_legend(nrow=4)) +
  geom_sf(data=m.river2,color="cadetblue2",fill="cadetblue2")+
  geom_sf(data=c.river, color="cadetblue2")+
  theme(        legend.position = "bottom",
                text  = element_text(color="black",  family="sans", size=10),
         panel.background = element_blank(),
        plot.margin = unit(c(0, 0, 0, 0), "cm"))+
  labs(fill="")
return(plot)
}
summary(stand.t.wide)

stand.t.wider = stand.t.wide %>% pivot_wider(names_from = year, values_from = c(count.total.ad.sap_Pima:stand.age))
stand.t.wider =stand.t.wider %>% replace(is.na(.), 0)
stand.t.wider = stand.t.wider %>% mutate(init.species.map = ifelse(init.species==1, "Pima",
                                                                   ifelse(init.species==2, "Pigl",
                                                                          ifelse(init.species==3, "Potr",
                                                                                 ifelse(init.species==4,"Bene",
                                                                                        ifelse(init.species==5, "Mixed","Not forested"))))))
stand.t.wider= stand.t.wider %>%filter(total.density.ad.sap_300>50)
stand.t.wider= stand.t.wider %>% mutate(dom.Pima_0 = as.factor(ifelse(init.species==1, "Dominant","Absent")),
                                        dom.Pigl_0 = as.factor(ifelse(init.species==2, "Dominant","Absent")),
                                        dom.Potr_0 = as.factor(ifelse(init.species==3, "Dominant","Absent")),
                                        dom.Bene_0 =as.factor(ifelse(init.species==4, "Dominant","Absent")),
                                        dom.Mixed_0 =as.factor(ifelse(init.species==5, "Dominant","Absent")),
                                        forest_present_0 = as.factor(ifelse(init.species==1|init.species==2|init.species==3|init.species==4|init.species==5, "Forested", "Absent")),
                                        dom_Pima_300 = as_factor(ifelse(IV.ad.sap_Pima_300 > 1, "Dominant","Absent")),
                                        dom_Pigl_300 = as_factor(ifelse(IV.ad.sap_Pigl_300 > 1, "Dominant","Absent")),
                                        dom_Potr_300 = as_factor(ifelse(IV.ad.sap_Potr_300 > 1, "Dominant","Absent")),
                                        dom_Bene_300 = as_factor(ifelse(IV.ad.sap_Bene_300 > 1, "Dominant","Absent")),
                                        dom_mixed_spruce_300 = as_factor(ifelse(dom_Pima_300 =="Absent" &dom_Pigl_300=="Absent" & dom_Potr_300=="Absent" &dom_Bene_300=="Absent" & IV.ad.sap_Pima_300+IV.ad.sap_Pigl_300 >1,"Dominant","Absent")),
                                        dom_mixed_deciduous_300 = as_factor(ifelse(dom_Pima_300=="Absent" &dom_Pigl_300=="Absent" & dom_Potr_300=="Absent" &dom_Bene_300=="Absent" &IV.ad.sap_Bene_300 + IV.ad.sap_Potr_300 >1,"Dominant","Absent")),
                                        forest_present_300 = as_factor(ifelse(count.total.ad.sap_Pima_300 >50 |count.total.ad.sap_Pigl_300 >50 |count.total.ad.sap_Potr_300 >50 | count.total.ad.sap_Bene_300 >50, "Forested", "Non forested")))

stand.t.wider= stand.t.wider %>% mutate(Pima.transition = ifelse(dom.Pima_0=="Dominant"& dom_Pima_300 =="Dominant", "Remained dominant",
                                                                  ifelse(dom.Pima_0=="Dominant"&forest_present_300=="Non forested", "Now non forested",
                                                                 ifelse(dom.Pima_0=="Dominant"& dom_Potr_300=="Dominant", "Now Potr",
                                                                        ifelse(dom.Pima_0=="Dominant"& dom_Pigl_300=="Dominant", "Now Pigl",
                                                                               ifelse(dom.Pima_0=="Dominant"& dom_Bene_300=="Dominant", "Now Bene",
                                                                                      ifelse(dom.Pima_0=="Dominant" & dom_mixed_spruce_300=="Dominant", "Now mixed spruce",
                                                                                             ifelse(dom.Pima_0=="Dominant" & dom_mixed_deciduous_300=="Dominant", "Now mixed deciduous",
                                                                                             ifelse(dom.Pima_0=="Absent"& dom_Pima_300 =="Dominant", "Now dominant","Never there")))))))))

stand.t.wider= stand.t.wider %>% mutate(Pigl.transition = ifelse(dom.Pigl_0=="Dominant"& dom_Pigl_300 =="Dominant", "Remained dominant",
                                                                  ifelse(dom.Pigl_0=="Dominant"&forest_present_300=="Non forested", "Now non forested",
                                                                 ifelse(dom.Pigl_0=="Dominant"& dom_Potr_300=="Dominant", "Now Potr",
                                                                        ifelse(dom.Pigl_0=="Dominant"& dom_Pima_300=="Dominant", "Now Pima",
                                                                               ifelse(dom.Pigl_0=="Dominant"& dom_Bene_300=="Dominant", "Now Bene",
                                                                                      ifelse(dom.Pigl_0=="Dominant" & dom_mixed_spruce_300=="Dominant", "Now mixed spruce",
                                                                                             ifelse(dom.Pigl_0=="Dominant" & dom_mixed_deciduous_300=="Dominant", "Now mixed deciduous",
                                                                                             ifelse(dom.Pigl_0=="Absent"& dom_Pigl_300 =="Dominant", "Now dominant","Never there")))))))))


stand.t.wider= stand.t.wider %>% mutate(Potr.transition = ifelse(dom.Potr_0=="Dominant"& dom_Potr_300 =="Dominant", "Remained dominant",
                                                                  ifelse(dom.Potr_0=="Dominant"&forest_present_300=="Non forested", "Now non forested",
                                                                 ifelse(dom.Potr_0=="Dominant"& dom_Pigl_300=="Dominant", "Now Pigl",
                                                                        ifelse(dom.Potr_0=="Dominant"& dom_Pima_300=="Dominant", "Now Pima",
                                                                               ifelse(dom.Potr_0=="Dominant"& dom_Bene_300=="Dominant", "Now Bene",
                                                                                      ifelse(dom.Potr_0=="Dominant" & dom_mixed_spruce_300=="Dominant", "Now mixed spruce",
                                                                                             ifelse(dom.Potr_0=="Dominant" & dom_mixed_deciduous_300=="Dominant", "Now mixed deciduous",
                                                                                             ifelse(dom.Potr_0=="Absent"& dom_Potr_300 =="Dominant", "Now dominant","Never there")))))))))
stand.t.wider= stand.t.wider %>% mutate(Bene.transition = ifelse(dom.Bene_0=="Dominant"& dom_Bene_300 =="Dominant", "Remained dominant",
                                                                  ifelse(dom.Bene_0=="Dominant"&forest_present_300=="Non forested", "Now non forested",
                                                                 ifelse(dom.Bene_0=="Dominant"& dom_Pigl_300=="Dominant", "Now Pigl",
                                                                        ifelse(dom.Bene_0=="Dominant"& dom_Pima_300=="Dominant", "Now Pima",
                                                                               ifelse(dom.Bene_0=="Dominant"& dom_Potr_300=="Dominant", "Now Potr",
                                                                                      ifelse(dom.Bene_0=="Dominant" & dom_mixed_spruce_300=="Dominant", "Now mixed spruce",
                                                                                             ifelse(dom.Bene_0=="Dominant" & dom_mixed_deciduous_300=="Dominant", "Now mixed deciduous",
                                                                                             ifelse(dom.Bene_0=="Absent"& dom_Bene_300 =="Dominant", "Now dominant","Never there")))))))))

                                                                 
   


mapping.data=stand.t.wider
mapping.data$Pima.transition <- factor(mapping.data$Pima.transition, levels = c("Remained dominant", "Now dominant" , "Now Pima","Now Pigl","Now Potr", "Now Bene", "Now mixed spruce", "Now mixed deciduous", "Never there"))
#   summary(as.factor(mapping.data$Pigl.transition))
# Pima.dom.yr0 = map(mapping.data, "dom.Pima_0", "0",cols.1)
# Pigl.dom.yr0 = map(mapping.data, "dom.Pigl_0", "0",cols.1)
# Potr.dom.yr0 = map(mapping.data, "dom.Potr_0", "0",cols.1)
# Bene.dom.yr0 = map(mapping.data,"dom.Bene_0", "0",cols.1)
# forested.yr0 = map(mapping.data,"forest_present_0", "0",cols.2)
# 
# Pima.dom.yr150 = map(mapping.data, "dom_Pima_150", "150",cols.1)
# Pigl.dom.yr150 = map(mapping.data, "dom_Pigl_150", "150",cols.1)
# Potr.dom.yr150 = map(mapping.data, "dom_Potr_150", "150",cols.1)
# Bene.dom.yr150 = map(mapping.data, "dom_Bene_150", "150",cols.1)
# forested.yr150 = map(mapping.data, "forest_present_150", "150",cols.2)
# 
# Pima.dom.yr300 = map(mapping.data, "dom_Pima_300", "300",cols.1)
# Pigl.dom.yr300 = map(mapping.data, "dom_Pigl_300", "300",cols.1)
# Potr.dom.yr300 = map(mapping.data, "dom_Potr_300", "300",cols.1)
# Bene.dom.yr300 = map(mapping.data, "dom_Bene_300", "300",cols.1)
# mixed.dom.yr300 =map(mapping.data, "dom_mixed_300", "300",cols.1)
# forested.yr300 = map(mapping.data, "forest_present_300", "300",cols.2)
pima.lost = map(mapping.data, "Pima.transition",300,cols.3)
ggsave("D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/figures/pima.species.map.jpg",height=4.5,width=6.5,units="in")
pigl.lost = map(mapping.data, "Pigl.transition",300,cols.3)
ggsave("D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/figures/pigl.species.map.jpg",height=4.5,width=4.5,units="in")
potr.lost = map(mapping.data, "Potr.transition",300,cols.3)
ggsave("D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/figures/potr.species.map.jpg",height=4.5,width=4.5,units="in")
bene.lost = map(mapping.data, "Bene.transition",300,cols.3)
ggsave("D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/figures/bene.species.map.jpg",height=4.5,width=4.5,units="in")
# 
# 
# pdf(paste("D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/cpcrw/model/output/species_presence_dom_",v,".pdf",sep=""),height=8,width=10)
# cowplot::plot_grid(forested.yr0, forested.yr300)
# cowplot::plot_grid(Pigl.dom.yr0, Pigl.dom.yr300,pigl.lost)
# cowplot::plot_grid(Potr.dom.yr0,Potr.dom.yr300,potr.lost)
# cowplot::plot_grid(Bene.dom.yr0,Bene.dom.yr300,bene.lost)
# dev.off()



init.map=ggplot(stand.t.wider)+
  geom_tile(data=hill.df, aes(x=x, y=y, fill=value),show.legend = F)+
  scale_fill_gradient(low = "grey0", high = "grey100") +
  ggnewscale::new_scale_fill() +
  geom_raster(aes(x=x, y=y, fill=as.factor(init.species.map)),alpha=0.6)+
  scale_fill_manual(values = cols.4)+
  ggthemes::theme_map() +
  geom_sf(data=m.river2,color="cadetblue2",fill="cadetblue2")+
  geom_sf(data=c.river, color="cadetblue2")+
  theme(        legend.position = "right",
         panel.background = element_blank(),
        plot.margin = unit(c(0, 0, 0, 0), "cm"))+
  labs(fill="")
ggsave("D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/figures/init.species.map.jpg",height=4.5,width=6.5,units="in")

level.2 = cowplot::plot_grid(pima.lost,pigl.lost,potr.lost,bene.lost,nrow=2)
level.1=cowplot::plot_grid(init.map,ncol=2)
level.3 = cowplot::plot_grid(level.1,level.2,nrow=3)
```
Calculating change in species composition
```{r}
d = mapping.data %>% dplyr::select(dom.Pima_0,dom_Pima_300,dom.Pigl_0,dom_Pigl_300,dom.Potr_0,dom_Potr_300,dom.Bene_0,dom_Bene_300,dom.Mixed_0,dom_mixed_spruce_300,dom_mixed_deciduous_300)


calc.percentage =function(var,species,label){
  t=length(which(var=="Dominant"))/60826*100
  print(paste0("Species ",species," year ",label, " percentage = ",round(t,digits=1)))
}
length(which(d$dom.Pigl_0=="Dominant"|d$dom.Pima_0=="Dominant"))
length(which(d$dom_Pigl_300=="Dominant"|d$dom_Pima_300=="Dominant"|d$dom_mixed_spruce_300=="Dominant"))

length(which(d$dom.Potr_0=="Dominant"|d$dom.Bene_0=="Dominant"|d$dom.Mixed_0=="Dominant"))
length(which(d$dom_Potr_300=="Dominant"|d$dom_Bene_300=="Dominant"|d$dom_mixed_deciduous_300=="Dominant"))

length(which(d$dom.Pima_0=="Dominant"|d$dom.Pigl_0=="Dominant"|d$dom.Bene_0=="Dominant"|d$dom.Potr_0=="Dominant"|d$dom.Mixed_0=="Dominant"))
length(which(d$dom_Pima_300=="Dominant"|d$dom_Pigl_300=="Dominant"|d$dom_Bene_300=="Dominant"|d$dom_Potr_300=="Dominant"|d$dom_mixed_spruce_300=="Dominant"|d$dom_mixed_deciduous_300=="Dominant"))
length(which(mapping.data$total.density.ad.sap_300>50))

calc.percentage(d$dom.Pima_0,"Pima","0")
calc.percentage(d$dom.Pigl_0,"Pigl","0")
calc.percentage(d$dom.Bene_0,"Bene","0")
calc.percentage(d$dom.Potr_0,"Potr","0")
calc.percentage(d$dom.Mixed_0,"Mixed","0")

calc.percentage(d$dom_Pima_300,"Pima","300")
calc.percentage(d$dom_Pigl_300,"Pigl","300")
calc.percentage(d$dom_Bene_300,"Bene","300")
calc.percentage(d$dom_Potr_300,"Potr","300")
calc.percentage(d$dom_mixed_spruce_300,"Mixed_spruce","300")
calc.percentage(d$dom_mixed_deciduous_300,"Mixed_deciduous","300")

```
