---
title: "permafrost distribution"
author: "Winslow"
date: "11/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(terra)
library(tidyverse)
```





```{r}

db.conn <- DBI::dbConnect(RSQLite::SQLite(), dbname="D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/cpcrw/model/output/paper_run/CPCRW_sm_6.sqlite")
water = tbl(db.conn, "water")  %>% filter(year>270) %>% collect()
stand = tbl(db.conn, "stand")  %>% filter(year==271) %>% collect()
reg.det = tbl(db.conn, "saplingdetail") %>% filter(year==271) %>% collect()
  DBI::dbDisconnect(db.conn)
```
Processing regeneration to bind with stand
```{r}
reg.det=reg.det%>%mutate(ba=pi*((dbh/100)/2)^2)
reg.det=reg.det%>%mutate(ba_all=ba*n_represented)
reg.det.sum=reg.det%>%group_by(year,ru, species)%>%summarize(count_ha_sap=sum(n_represented),
                                                             dbh_mean_sapling=weighted.mean(dbh,n_represented),
                                                             ba_sum_sapling=sum(ba_all))


```

Merging stand and regenration together to calculate total density.

```{r}
stand = stand %>% dplyr::select(-cohort_count_ha,-cohort_basal_area)
stand.t=full_join(stand,reg.det.sum,by = c("ru","year","species"))

stand.t =stand.t %>% replace(is.na(.), 0)
stand.t=stand.t%>%mutate(count.total.ad.sap = count_ha + count_ha_sap,
                         ba.total.ad.sap = basal_area_m2+ba_sum_sapling)
```
Now pivoting wider to generate IV values
```{r}
stand.t2.select = stand.t %>% dplyr::select(ru,rid, year,species, count.total.ad.sap:ba.total.ad.sap)
stand.t.wide = stand.t2.select %>% pivot_wider( names_from = species, values_from = c(count.total.ad.sap:ba.total.ad.sap), values_fill = 0)
stand.t.wide=stand.t.wide %>% mutate(total.density.ad.sap = count.total.ad.sap_Potr + count.total.ad.sap_Pima + count.total.ad.sap_Bene + count.total.ad.sap_Pigl,
                                     total.ba.ad.sap = ba.total.ad.sap_Potr + ba.total.ad.sap_Pima + ba.total.ad.sap_Bene + ba.total.ad.sap_Pigl)
stand.t.wide=stand.t.wide%>%mutate(IV.ad.sap_Pima = ifelse(total.density.ad.sap !=0 & total.ba.ad.sap!=0, (count.total.ad.sap_Pima/total.density.ad.sap)+(ba.total.ad.sap_Pima/total.ba.ad.sap),0),
                                   IV.ad.sap_Pigl = ifelse(total.density.ad.sap !=0 & total.ba.ad.sap!=0, (count.total.ad.sap_Pigl/total.density.ad.sap)+(ba.total.ad.sap_Pigl/total.ba.ad.sap),0),
                                   IV.ad.sap_Potr = ifelse(total.density.ad.sap !=0 & total.ba.ad.sap!=0, (count.total.ad.sap_Potr/total.density.ad.sap)+(ba.total.ad.sap_Potr/total.ba.ad.sap), 0),
                                   IV.ad.sap_Bene = ifelse(total.density.ad.sap !=0 & total.ba.ad.sap!=0,(count.total.ad.sap_Bene/total.density.ad.sap)+(ba.total.ad.sap_Bene/total.ba.ad.sap), 0))

```

read in the map to make the data spatial
```{r}
env.grid = rast("D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/cpcrw/materials/env/env.grid.tif")
env.grid.df=as.data.frame(env.grid, xy=T)
env.grid.df = env.grid.df %>% rename(rid = env.grid)
```
```{r}
water= left_join(water, env.grid.df, by="rid")
summary(water)
```

generate a raster from model results to compare to the Pastick product

```{r}
water.2=water%>%filter(year>1)
water.2=water.2 %>% mutate(permafrost_Modeled.NS = ifelse(maxDepthThawed<=1, 1,0),
                       permafrost_Modeled.Deep = ifelse(maxDepthThawed>1 & maxDepthThawed < 2, 1,0))
water.subset = water.2 %>%group_by(x,y,rid)%>% summarize(sum.permafrost_Modeled.NS=sum(permafrost_Modeled.NS),
                                           sum.permafrost_Modeled.Deep = sum(permafrost_Modeled.Deep))

water.subset=water.subset %>%mutate(pf.ns =ifelse(sum.permafrost_Modeled.NS >=15,1,0),
                                    pf.deep=ifelse(sum.permafrost_Modeled.Deep>=15,1,0))
summary(water.subset)
plot.df=water.subset%>%dplyr::select(x,y,pf.ns)
plot.df=plot.df %>%filter(!is.na(x))
pf.m.ras = rast(plot.df,type="xyz") 
crs(pf.m.ras)=crs(env.grid)
plot(pf.m.ras)
water.subset2 =water.subset %>% dplyr::select(rid, "permafrost_Modeled.NS"=pf.ns, "permafrost_Modeled.Deep"=pf.deep)
```


Pastick et al. reprojected to 1ha
```{r}
pf.b=rast("D:/workspace/Winslow/permafrost/permafrost_param/benchmarking data/AK_PFProbability_1m/cpcrw-1ha-repr.tif")
pf.b
classify.f=function(data.set, val.1,val.2,val.3){
m <- c(val.1, val.2, val.3)
rclmat <- matrix(m, ncol=3, byrow=TRUE)
rc1 <- classify(data.set, rclmat, include.lowest=TRUE)
return(rc1)
}
step.1 = classify.f(pf.b, 0,50,0)
hist(step.1)
step.2=classify.f(step.1, 50,101,1)
plot(step.1)
```


Read in aspect to link to the others
```{r}
aspect=rast("D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/raw data/DEM/aspect_repr.tif")
aspect.repr= terra::project(aspect, env.grid, method="bilinear")

```

Now set the rid for the benchmarks and extract the datasets
```{r}
  points=as.points(env.grid)  #turn response raster into a points dataset in the middle of each raster cell.
data=data.frame(terra::extract(aspect.repr, points,xy=T,method="simple"), terra::extract(step.2, points,method="simple"))
data$ID.1 = NULL
data = data %>% rename("rid" = ID,
                       "aspect" = aspect_repr,
                       "permafrost_Benchmark.NS"= Layer_1)
```


Link the two and pivot longer
```{r}

water.subset2 = left_join(water.subset2, data, by ="rid")
summary(water.subset2)
data.long = water.subset2 %>% pivot_longer(cols = c(permafrost_Modeled.NS:permafrost_Modeled.Deep,permafrost_Benchmark.NS),  names_to = "form", values_to = "permafrost", names_prefix = "permafrost_")

```

summarizing percent difference in permafrost distribution

```{r}
data.long.sum = data.long %>% group_by(form) %>% summarize(perc.pf = sum(permafrost,na.rm=T)/length(permafrost)*100)

data.long.sum.wider = data.long.sum %>%pivot_wider(names_from = form, values_from = c(2))



data.long.sum.wider = data.long.sum.wider %>% mutate(pf.ns.diff = abs(Benchmark.NS-Modeled.NS)/((Benchmark.NS+Modeled.NS)/2)*100)
data.long.sum.wider

```
generate plots
First generate a hillshade effect
```{r}
dem=rast("D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/raw data/DEM/dem_repr.tif")
dem=crop(dem,env.grid)
slope=terrain(dem,v="slope",unit="radians")
aspect=terrain(dem, v="aspect",unit="radians")
hill=shade(slope,aspect,40,270,normalize=T)
hill.df =as.data.frame(hill,xy=T)
colnames(hill.df) <- c( "x", "y", "value")

```
Read in water
```{r}
library(sf)
m.river=st_read("D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/raw data/hydrology/export_1594471551598/hydro_63360_py_repr_cpcrw.shp")
c.river=st_read("D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/raw data/hydrology/attachments/cpcrw_streams.shp")
m.river2=st_transform(m.river,st_crs(c.river))
```

Generating dataset of modeled fires
```{r}
db.conn <- DBI::dbConnect(RSQLite::SQLite(), dbname="D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/cpcrw/model/output/paper_run/CPCRW_sm_6.sqlite")
fire.model = tbl(db.conn, "fire")  %>% filter(year>=2) %>% collect()
  DBI::dbDisconnect(db.conn)
  summary(fire.model)
min(fire.model$fireId)
max(fire.model$fireId)  # these need to be updated each time  
path = "D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/cpcrw/model/output/paper_run/spread_6"
  files <- list.files(path=path, pattern = "[2][4-9]|[3][0-9]\\.txt$", full.names=T)
  fire.maps=rast(files)
  crs(fire.maps)=crs(env.grid)
  fire.maps=terra::project(fire.maps,env.grid)
fire.maps.final=sum(fire.maps)
fire.maps.final.df = as.data.frame(fire.maps.final,xy=T)
colnames(fire.maps.final.df) <- c( "x", "y", "value")
fire.maps.final.df = fire.maps.final.df %>%mutate(value=ifelse(value>0,1,value))
fire.maps.final.df = fire.maps.final.df %>%mutate(value=ifelse(value==0,NA,value))
fire.maps.final.df = fire.maps.final.df %>% filter(!is.na(value))
```


```{r}
plotting.fun=function(dataset,var,dataset.2, color){
 plot= ggplot()+
  geom_tile(data=hill.df, aes(x=x, y=y, fill=value),show.legend = F)+
  scale_fill_gradient(low = "grey0", high = "grey100") +
  ggnewscale::new_scale_fill() +
  geom_tile(data =dataset, aes_string(x = "x.x", y = "y.x", fill =var ), alpha=0.6) +
scale_fill_manual(values=color)+ 
       labs(fill="")+
      ggnewscale::new_scale_fill() +
   ggthemes::theme_map() +
  geom_sf(data=m.river2,color="cadetblue2",fill="cadetblue2")+
  geom_sf(data=c.river, color="cadetblue2")+
  theme( legend.position = "",
         panel.background = element_blank(),
        plot.margin = unit(c(0, 0, 0, 0), "cm"))+
 labs(x="",y="")
  print(plot)
}


plot.modeled=water.subset2 %>%mutate(pf=as.factor(ifelse(permafrost_Modeled.NS==1,"Present","Absent")))%>%filter(pf=="Present")
plot.benchmark=water.subset2 %>%mutate(pf=as.factor(ifelse(permafrost_Benchmark.NS==1,"Present","Absent")))%>%filter(pf=="Present")

pf.m=plotting.fun(plot.modeled, "pf",fire.maps.final.df, "#b35806")
pf.b=plotting.fun(plot.benchmark, "pf",fire.df,"#542788")
maps=cowplot::plot_grid(pf.b,pf.m,nrow=1)

```
join water and stand
```{r}
data.long=left_join(data.long, stand.t.wide, by="rid")

```
```{r}
data.long= data.long %>% mutate(sp.dom = as.factor(ifelse(IV.ad.sap_Pima > 1,"Pima",
                                      ifelse(IV.ad.sap_Potr > 1, "Potr",
                                             ifelse(IV.ad.sap_Pigl > 1, "Pigl",
                                                    ifelse(IV.ad.sap_Bene > 1, "Bene",
                                                           ifelse(IV.ad.sap_Pima+IV.ad.sap_Pigl > 1, "Mixed spruce",
                                                                  ifelse(IV.ad.sap_Potr+IV.ad.sap_Bene > 1, "Mixed deciduous", "Non forest"))))))))
data.long=data.long %>%filter(!is.na(sp.dom))
data.long%>%filter(sp.dom=="NA")
summ.plot1 = data.long %>% group_by(form, sp.dom) %>% summarize(percent.pf = sum(permafrost,na.rm=T)/length(permafrost)*100)
summ.plot1.wider = summ.plot1 %>%pivot_wider(names_from = form, values_from = c(3))



summ.plot1.wider = summ.plot1.wider %>% mutate(pf.ns.diff = abs(Benchmark.NS-Modeled.NS)/((Benchmark.NS+Modeled.NS)/2)*100)


summ.plot1 = summ.plot1 %>% mutate(form.final = ifelse(form == "Benchmark.NS", "Benchmark <= 1m", 
                                                       ifelse(form == "Modeled.NS", "Modeled <=1m", "Modeled 1-2m")))
summ.plot1$form= factor(summ.plot1$form, levels= c("Benchmark.NS", "Modeled.NS", "Modeled.Deep"))
summ.plot1=summ.plot1 %>% filter(form=="Modeled.NS"|form=="Benchmark.NS")
summ.plot1$sp.dom= factor(summ.plot1$sp.dom, levels= c("Pima", "Pigl", "Mixed spruce", "Bene", "Potr", "Mixed deciduous", "Non forest"))
summ.plot1
a=ggplot(summ.plot1, aes(x=as.factor(sp.dom), y=percent.pf, fill=form.final))+
  geom_bar(stat="identity",position="dodge")+
  scale_fill_manual(values=c("#542788","#b35806"))+
  theme_bw()+
  theme(    text  = element_text(color="black",  family="sans", size=10),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
             axis.text.x=element_text(colour="black",size=10),
             axis.text.y=element_text(colour="black",size=10),
            legend.text = element_text(size=8),
             legend.position= "bottom",
            legend.background = element_rect(colour = "transparent", fill = 'transparent'),
            legend.key = element_rect(colour = "transparent", fill = 'transparent'),
            legend.key.size = unit(0.5, "cm"))+
  labs(x="Dominant species", y="Area with permafrost (%)",fill="")

```
plot the percent of landscape by aspect for modeled and benchmark
```{r}


data.long=data.long %>% filter(!is.na(aspect))

data.long=data.long%>%mutate(aspect.dir=as.factor(ifelse(aspect == -1, "Flat",
                                       ifelse(aspect>-1&aspect<=45,"North",
                                              ifelse(aspect>45&aspect<=135,"East",
                                                     ifelse(aspect>135&aspect<=225,"South",
                                                           ifelse(aspect>225&aspect<=315,"West","North")))))))

summ.plot = data.long %>% group_by(form, aspect.dir) %>% summarize(percent.pf = sum(permafrost,na.rm=T)/length(permafrost)*100)

sum.plot.wide = summ.plot %>%pivot_wider(names_from = form, values_from = c(3))

sum.plot.wide = sum.plot.wide %>% mutate(pf.ns.diff = abs(Benchmark.NS-Modeled.NS)/((Benchmark.NS+Modeled.NS)/2)*100)

summ.plot = summ.plot %>% mutate(form.final = ifelse(form == "Benchmark.NS", "Benchmark <= 1m", 
                                                       ifelse(form == "Modeled.NS", "Modeled <=1m", "Modeled 1-2m")))
summ.plot$form= factor(summ.plot$form, levels= c("Benchmark.NS", "Modeled.NS", "Modeled.Deep"))
summ.plot$aspect.dir= factor(summ.plot$aspect.dir, levels= c("North", "East", "South", "West", "Flat"))
summ.plot=summ.plot %>% filter(form=="Modeled.NS"|form=="Benchmark.NS")

b=ggplot(summ.plot, aes(x=as.factor(aspect.dir), y=percent.pf, fill=form.final))+
  geom_bar(stat="identity",position="dodge")+
  scale_fill_manual(values=c("#542788","#b35806"))+
  theme_bw()+
  theme(    text  = element_text(color="black",  family="sans", size=10),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
             axis.text.x=element_text(colour="black",size=10),
             axis.text.y=element_text(colour="black",size=10),
            legend.text = element_text(size=8),
             legend.position= "bottom",
            legend.background = element_rect(colour = "transparent", fill = 'transparent'),
            legend.key = element_rect(colour = "transparent", fill = 'transparent'),
            legend.key.size = unit(0.5, "cm"))+
  labs(x="Aspect", y="Permafrost area (%)",fill="")
cowplot::plot_grid(maps,b,labels="AUTO",nrow=2,label_size=10, label_fontface = "plain", label_fontfamily = "sans")
ggsave("D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/figures/pf.benchmark.jpg", height=6,width=6.5, units = "in", dpi=700)

```





Plotting maps of SOL
```{r}
plotting.fun=function(dataset,var,label){
 plot= ggplot()+
  geom_tile(data=hill.df, aes(x=x, y=y, fill=value),show.legend = F)+
  scale_fill_gradient(low = "grey0", high = "grey100") +
  ggnewscale::new_scale_fill() +
  geom_tile(data =dataset, aes_string(x = "x", y = "y", fill =var ), alpha=0.6) +
scale_fill_gradientn(colours = rev(terrain.colors(10))) +  
   ggthemes::theme_map() +
  geom_sf(data=m.river2,color="cadetblue2",fill="cadetblue2")+
  geom_sf(data=c.river, color="cadetblue2")+
  theme( legend.position = "bottom",
         panel.background = element_blank(),
        plot.margin = unit(c(0, 0, 0, 0), "cm"))+
    labs(fill=label)
  print(plot)
}

 
  plot.df=water %>% filter(year==300) %>%dplyr::select(x,y,SOLLayer,mossLayer)
plot.df=plot.df %>%filter(!is.na(x))
plot.df=plot.df %>%mutate(SOLLayer=SOLLayer*100,
                         mossLayer=mossLayer*100)
summary(plot.df)

plot.df %>% summarize(quantile.moss = quantile(mossLayer, c(0.25,0.5,0.75),q=c(0.25,0.5,0.75)),
                      quantile.sol = quantile(SOLLayer, c(0.25,0.5,0.75),q=c(0.25,0.5,0.75)))
s=plotting.fun(plot.df,"SOLLayer","Dead moss and \n tree litter depth (cm)")
m=plotting.fun(plot.df,"mossLayer","Live moss \n depth (cm)")
sol=cowplot::plot_grid(s,m,nrow=1)
ggsave("D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/figures/sol.benchmark.jpg", height=3,width=6.5, units = "in", dpi=700)
```

Determining SOL depth as a function of dominant forest type

Map raster
```{r}
map = rast("D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/cpcrw/model/temp/spinup_300.asc")
map
map.df=as.data.frame(map, xy=T)
map.df = map.df %>% rename(ru = spinup_300)
map.df %>% filter(ru!=-1) %>%summary(map.df)
```


Read in the produce a stand composition and regen dataset to check stand densities at the beginning and end of the simulation and then pair pare with the simulated carbon dataset
```{r}

db.conn <- DBI::dbConnect(RSQLite::SQLite(), dbname="D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/cpcrw/model/output/paper_run/CPCRW_sm_6.sqlite")
stand = tbl(db.conn, "stand")  %>% filter(year==300) %>% collect()
reg.det = tbl(db.conn, "saplingdetail") %>% filter(year==300)%>% collect ()
water=tbl(db.conn, "water") %>% filter(rid>-1) %>% collect
fire =  tbl(db.conn, "fire") %>% collect()
  DBI::dbDisconnect(db.conn)
```
Processing regeneration to bind with stand
```{r}
reg.det=reg.det%>%mutate(ba=pi*((dbh/100)/2)^2)
reg.det=reg.det%>%mutate(ba_all=ba*n_represented)
reg.det.sum=reg.det%>%group_by(year,ru, species)%>%summarize(count_ha_sap=sum(n_represented),
                                                             dbh_mean_sapling=weighted.mean(dbh,n_represented),
                                                             ba_sum_sapling=sum(ba_all))


```

Merging stand and regenration together to calculate total density.

```{r}
stand = stand %>% dplyr::select(-cohort_count_ha,-cohort_basal_area)
stand.t=full_join(stand,reg.det.sum,by = c("ru","year","species"))

stand.t =stand.t %>% replace(is.na(.), 0)
stand.t2=stand.t%>%mutate(count.total.ad.sap = count_ha + count_ha_sap,
                         ba.total.ad.sap = basal_area_m2+ba_sum_sapling)
summary(stand.t2)
```
Now pivoting wider to generate IV values
```{r}
stand.t2.select = stand.t2 %>% dplyr::select(ru,rid, year,species, count.total.ad.sap:ba.total.ad.sap, LAI)
stand.t.wide = stand.t2.select %>% pivot_wider( names_from = species, values_from = c(count.total.ad.sap:LAI), values_fill = 0)
stand.t.wide=stand.t.wide %>% mutate(total.density.ad.sap = count.total.ad.sap_Potr + count.total.ad.sap_Pima + count.total.ad.sap_Bene + count.total.ad.sap_Pigl,
                                     total.ba.ad.sap = ba.total.ad.sap_Potr + ba.total.ad.sap_Pima + ba.total.ad.sap_Bene + ba.total.ad.sap_Pigl)
stand.t.wide=stand.t.wide%>%mutate(IV.ad.sap_Pima = ifelse(total.density.ad.sap !=0 & total.ba.ad.sap!=0, (count.total.ad.sap_Pima/total.density.ad.sap)+(ba.total.ad.sap_Pima/total.ba.ad.sap),0),
                                   IV.ad.sap_Pigl = ifelse(total.density.ad.sap !=0 & total.ba.ad.sap!=0, (count.total.ad.sap_Pigl/total.density.ad.sap)+(ba.total.ad.sap_Pigl/total.ba.ad.sap),0),
                                   IV.ad.sap_Potr = ifelse(total.density.ad.sap !=0 & total.ba.ad.sap!=0, (count.total.ad.sap_Potr/total.density.ad.sap)+(ba.total.ad.sap_Potr/total.ba.ad.sap), 0),
                                   IV.ad.sap_Bene = ifelse(total.density.ad.sap !=0 & total.ba.ad.sap!=0,(count.total.ad.sap_Bene/total.density.ad.sap)+(ba.total.ad.sap_Bene/total.ba.ad.sap), 0))
summary(stand.t.wide)
```

Approach to identify time since fire for each grid cell in stand.t.wide
```{r}
# Let's get all the spread files (fire extents) as a raster stack
path = "D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/cpcrw/model/output/paper_run/spread_6"
fire_maps = rast(list.files(path, full.names = T))

# Use the environment grid to get ids for each grid
path = "D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/cpcrw/model/gis"
env_grid <- rast(paste0(path, "/env.grid.txt"))
crs(fire_maps)=crs(env_grid)

#env_grid.2=crop(env_grid,fire_maps)
env_grid.2=env_grid
fire <- fire %>% mutate(prop.dens.killed = n_trees_died   / n_trees,
                        prop.ba.killed   = basalArea_died / basalArea_total)
fire$prop.ba.killed  [fire$basalArea_total == 0] <- 0
fire$prop.dens.killed[fire$n_trees         == 0] <- 0

fire_year_dat <- data.frame()

for (i in 1:nrow(fire)) {
  
  # Skip fires that burned nothing
  if (fire$area_m2[i] > 0) {
    
    # Take the fire map for this fire, and make everything but the fire cells NA
    mask <- fire_maps[[paste0("spread_", fire$fireId[i])]]
    mask[mask==0] <- NA
    
    # Use the mask to extract the rids for burned grids
    burned_rids <- as.data.frame(raster::mask(env_grid.2, mask))
burned_rids= burned_rids%>%filter(!is.na(env.grid))
burned_rids$year=fire$year[i]
  }
  fire_year_dat=rbind(fire_year_dat,burned_rids)
}
fire_year_dat = fire_year_dat %>% rename("rid" = env.grid) %>% mutate(fire=1)
summary(fire_year_dat)
```

Join together fire year and stand wide
```{r}
summary(stand.t.wide)
stand.t.wide = left_join(stand.t.wide, fire_year_dat, by=c("rid","year"))  

stand.t.wide =stand.t.wide %>% replace(is.na(.), 150)


stand.t.wide=stand.t.wide%>%
  group_by(rid) %>%
  mutate(idx=cumsum(fire==1))%>%
  group_by(rid,idx) %>%
  mutate(time_since_fire=row_number())%>%
  ungroup() %>% 
    dplyr::select(-idx)

test=stand.t.wide %>% filter(rid ==  50746)

summary(test)
```
```{r}

stand.t.final=left_join(stand.t.wide,water,by = c("rid","year"))
summary(stand.t.final)
stand.t.final %>% filter(IV.ad.sap_Pima>1)%>% summary()
```
```{r}
library(scales)
stand.t.final= stand.t.final %>% mutate(sp.dom = as.factor(ifelse(IV.ad.sap_Pima > 1,"Pima",
                                      ifelse(IV.ad.sap_Potr > 1, "Potr",
                                             ifelse(IV.ad.sap_Pigl > 1, "Pigl",
                                                    ifelse(IV.ad.sap_Bene > 1, "Bene",
                                                           ifelse(IV.ad.sap_Pima+IV.ad.sap_Pigl>1, "Mixed spruce",
                                                                  ifelse(IV.ad.sap_Bene+IV.ad.sap_Potr, "Mixed deciduous","Non forest"))))))))
stand.t.final$sp.dom <- factor(stand.t.final$sp.dom, levels = c("Pima", "Pigl",  "Potr", "Bene","Mixed spruce", "Mixed deciduous", "Non forest"))
stand.t.final=stand.t.final %>% mutate(total=mossLayer*100+SOLLayer*100)

stand.t.final %>%filter(year==300) %>% group_by(sp.dom)%>%filter(total>0) %>%summarise(quantile.total  = quantile(total, c(0.25,0.5,0.75)),q=c(0.25,0.5,0.75))
stand.t.final = stand.t.final %>% filter(sp.dom!="Non forest")
comparison.plot2=ggplot(stand.t.final, aes(x=total,y=sp.dom, height = ..ndensity..))+
  ggridges::geom_density_ridges(alpha=0.5)+
  theme_bw()+
  scale_x_continuous(labels = comma)+
  theme(    text  = element_text(color="black",  family="sans", size=10),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
             axis.text.x=element_text(colour="black",size=10),
             axis.text.y=element_text(colour="black",size=10),
            legend.position = "bottom",
            legend.background = element_rect(colour = "transparent", fill = 'transparent'),
            legend.key = element_rect(colour = "transparent", fill = 'transparent'))+
  labs(x=bquote(Surface~organic~layer~(cm)),y="Dominant species", fill="")
ggsave("D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/figures/SOL.moss.depth.stand.jpg", height=4.25,width=8.5, units = "in", dpi=700)

```

```{r}
# Whether we've saved the water and soil table or need to create it again
data_saved <- F
key <- "spec"
tag <- "PIMA 0.1 decid 0.35"
```

```{r}
output_file <- "D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/cpcrw/model/output/paper_run/CPCRW_sm_6.sqlite"
#output_file <- "C:/Users/lora/Documents/Winslow/for_Lora/output.sqlite"
#-----------------------------------------------------------------------------#
# Load iLand fire table
#-----------------------------------------------------------------------------#
db.conn <- DBI::dbConnect(RSQLite::SQLite(), dbname=output_file)
fire <- dbReadTable(db.conn, "fire")
dbDisconnect(db.conn)
fire$area_ha <- fire$area_m2 / 10000
```

# Assembling a dataset of burned stuff, one year pre-fire to 30 years post-fire:

```{r}
if (!data_saved) {
  #---------------------------------------------------------------------------#
  # Read in the water table, which has info on soil organic layer; get this 
  # info for specific burned areas
  #---------------------------------------------------------------------------#
  db.conn <- DBI::dbConnect(RSQLite::SQLite(), dbname=output_file)
  water <- tbl(db.conn, "water") %>% collect()
  dbDisconnect(db.conn)
  
  # Let's get all the spread files (fire extents) as a raster stack
  path = "D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/cpcrw/model/output/paper_run/spread_6"
  #path = "E:/storage/Lora/from_winslow/output/spread"
  fire_maps = stack(list.files(path, full.names = T))
  # Use the environment grid to get ids for each grid
  path = "D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/cpcrw/materials/env"
  env_grid <- raster(paste0(path, "/env.grid.tif"))
  crs(fire_maps)=crs(env_grid)
  new.extent = extent(fire_maps)
  
#env_grid.2=crop(env_grid,new.extent)
  env_grid.2=env_grid                  ###This is where you adjust for a subset of the landscape or whole landscape

  fire <- fire %>% mutate(prop.dens.killed = n_trees_died   / n_trees,
                          prop.ba.killed   = basalArea_died / basalArea_total)
  fire$prop.ba.killed  [fire$basalArea_total == 0] <- 0
  fire$prop.dens.killed[fire$n_trees         == 0] <- 0
  
  #-----------------------------------------------------------------------------#
  # Build a dataset of SOL the year before a burn through 30 years after it. 
  #-----------------------------------------------------------------------------#
  fire_sol_dat <- NULL
  for (i in 1:nrow(fire)) {
    
    # Skip fires that burned nothing
    if (fire$area_m2[i] > 0) {
      
      # Take the fire map for this fire, and make everything but the fire cells NA
      mask <- fire_maps[[paste0("spread_", fire$fireId[i])]]
      mask[mask==0] <- NA
      
      # Use the mask to extract the rids for burned grids
      burned_rids <- getValues(raster::mask(env_grid.2, mask))
      burned_rids <- burned_rids[!is.na(burned_rids)]
      
      if (length(burned_rids) > 0) { # should never happen but...
        
        # Years we want info: year before the fire, through 30 years post-fire
        fire_years <- (fire$year[i] - 1):(fire$year[i] + 30)
        
        firedat <- water %>% filter(year %in% fire_years  &
                                      rid %in%  burned_rids)
        firedat$prop.dens.killed <- fire$prop.dens.killed[i]
        firedat$prop.ba.killed   <- fire$prop.ba.killed  [i]
        
        # Add the year of the record relative to the fire, and soil organic layer
        # thickness (go from m to cm)
        firedat <- firedat %>% 
          mutate(rel.year  = year - fire$year[i],
                 fireId    = fire$fireId[i],
                 fire.year = fire$year[i],
                 SOLLayer  = ifelse(is.na(SOLLayer) ,0,SOLLayer),
                 mossLayer = ifelse(is.na(mossLayer),0,mossLayer),
                 SOL.cm    = SOLLayer*100+mossLayer*100)
        
        fire_sol_dat <- rbind(fire_sol_dat,
                              firedat %>% dplyr::select(year, ru, rid,
                                                        fireId, rel.year, 
                                                        fire.year, 
                                                        SOLLayer, mossLayer,
                                                        SOL.cm,
                                                        prop.dens.killed,
                                                        prop.ba.killed))
        
      } else {
        stop("Better check on this...")
      }
      
    }
  }
  
  
  #---------------------------------------------------------------------------#
  # Also assemble a dataset of unburned pixels, for comparison
  #---------------------------------------------------------------------------#
  
  # Determine where fires happened anywhere
  fires <- sum(fire_maps) 
  
  # Flip around for the unburned pixels - this is where fires never happened
  mask <- sum(fire_maps) 
  mask[mask>0] <- NA
  
  # Use the mask to extract the rids for unburned grids
  unburned_rids <- getValues(raster::mask(env_grid.2, mask))
  unburned_rids <- unburned_rids[!is.na(unburned_rids)]
  
  nofire_sol_dat <- water                                                  %>% 
    filter(rid %in% unburned_rids)                         %>%
    mutate(SOLLayer  = ifelse(is.na(SOLLayer),0,SOLLayer),
           SOL.cm    = SOLLayer*100+mossLayer*100)         %>%
    dplyr::select(year, ru, rid, SOL.cm)
  
  rm(water)
}
```


Now we calculate species importance value. This is the way to determine which species was/is dominant...
Just do this for every resource unit, every year

```{r}
if (!data_saved) {
  #---------------------------------------------------------------------------#
  # Processing of sapling and stand tables. Sapling detail in the newest 
  # version is far too large to load completely, even in the capacious memory
  # of the workstation. So it is loaded and summarized in chunks.
  #---------------------------------------------------------------------------#
  
  # Get the stand and sapling tables and chop them down some
  db.conn <- DBI::dbConnect(RSQLite::SQLite(), dbname=output_file)
  stand          <- tbl(db.conn, "stand")         %>% collect() %>% filter(ru != -1) %>%                          
    dplyr::select(year, ru, rid, species, area_ha, count_ha, basal_area_m2)
  sapling        <- tbl(db.conn, "sapling")       %>% collect() %>% filter(ru != -1) %>%
    dplyr::select(ru, year, species, "sapling_count_ha"=count_ha, 
                  "sapling_count_ha_small" = count_small_ha)
  #sapling.detail <- tbl(db.conn, "saplingdetail") %>% filter(ru != -1) %>% collect()
  sapling.detail <- tbl(db.conn, "saplingdetail") %>% filter(year %in% 1:10) %>% collect()
  
  
  DBI::dbDisconnect(db.conn)
  
  # Add basal area to sapling.detail
  sapling.detail    = sapling.detail %>% mutate(basal.area.ind   = pi*((dbh/100)/2)^2)
  sapling.detail    = sapling.detail %>% mutate(basal.area.cohort= basal.area.ind*n_represented)
  
  # Group sapling info by year, location, species
  sapling.detail.sum= sapling.detail %>% group_by(year,ru, species) %>%   
    summarize(dbh_mean=weighted.mean(dbh,n_represented),
              basal.area.total=sum(basal.area.cohort))
  
  start = 11
  end = start + 9
  while (start < 300) {
    db.conn <- DBI::dbConnect(RSQLite::SQLite(), dbname=output_file)
    sapling.detail <- tbl(db.conn, "saplingdetail") %>% filter(year %in% start:end) %>% collect()
    DBI::dbDisconnect(db.conn)
  
  # Add basal area to sapling.detail
  sapling.detail    = sapling.detail %>% mutate(basal.area.ind   = pi*((dbh/100)/2)^2)
  sapling.detail    = sapling.detail %>% mutate(basal.area.cohort= basal.area.ind*n_represented)
  
  # Group sapling info by year, location, species
  sapling.detail.sum= rbind(sapling.detail.sum, 
                            sapling.detail %>% group_by(year,ru, species) %>%   
                            summarize(dbh_mean=weighted.mean(dbh,n_represented),
                            basal.area.total=sum(basal.area.cohort)))
  start <- start + 10
  end = start + 9
  }
  
  # Join all this up
  stand.sap.1=full_join(stand       ,sapling.detail.sum, by = c("ru","year","species"))
  final.trees=full_join(stand.sap.1 ,sapling,            by = c("ru","year","species"))
  final.trees=final.trees %>% replace(is.na(.), 0)
  
 # summary(final.trees)
  
  # Total density and basal area sums; this here will be by species
  final.trees = final.trees %>%
    mutate(count_total    = count_ha + sapling_count_ha + sapling_count_ha_small,
           basal.area_sum = basal_area_m2 + basal.area.total)
  
  # Widen table so each record has all species in it
  final.trees.wide = final.trees %>% 
    pivot_wider(names_from  = species, 
                names_sep   = ".", 
                values_from = c(area_ha:basal.area_sum), 
                values_fill = 0)
  #summary(final.trees.wide)
  
  # Final sums across species
  final.trees.wide = final.trees.wide %>% 
    mutate(total_density = count_total.Potr + count_total.Pima + count_total.Bene + count_total.Pigl,
           total_ba = basal.area_sum.Potr + basal.area_sum.Pima + basal.area_sum.Bene + basal.area_sum.Pigl)
  
  # Final importance value
  final.trees.wide = final.trees.wide%>%
    mutate(IV.Pima = ifelse(total_density !=0 & total_ba!=0, 
                            (count_total.Pima/total_density)+(basal.area_sum.Pima/total_ba),0),
           IV.Pigl = ifelse(total_density !=0 & total_ba!=0,
                            (count_total.Pigl/total_density)+(basal.area_sum.Pigl/total_ba),0),
           IV.Potr = ifelse(total_density !=0 & total_ba!=0,
                            (count_total.Potr/total_density)+(basal.area_sum.Potr/total_ba), 0),
           IV.Bene = ifelse(total_density !=0 &total_ba!=0,
                            (count_total.Bene/total_density)+(basal.area_sum.Bene/total_ba), 0))
  
  # Smallen the dataset a bit
  #final.trees.wide2 = final.trees.wide %>% dplyr::select(year, ru, count_total.Pima:IV.Bene) 
  #summary(final.trees.wide2)
  
  # Assign a stand type by whatever is the species with the highest IV
  iv_cols <- grep("IV", names(final.trees.wide))
  spp <- sapply(names(final.trees.wide)[iv_cols], function(x){strsplit(x, ".", fixed=T)[[1]][2]})
  ind <- apply(final.trees.wide[,iv_cols], MARGIN = 1, FUN = which.max)
  final.trees.wide$stand.type <- as.factor(spp[ind])
  
  rm(final.trees, #final.trees.wide, 
     stand, sapling, sapling.detail, stand.sap.1)
}
```

Join it up to our fire dataset and get some info by stand type

```{r}
if (!data_saved) {
  # We want to make sure we are looking at stand type at the right time. This will
  # give us the stand info along each year.
  fire_dat <- left_join(fire_sol_dat, final.trees.wide, by=c("ru","year"))
  
  # But we also want to know the stand type at the time of the fire, so we can
  # see if it changes. Add that as a separate variable.
  fire.dat.2 <- fire_dat %>% filter(rel.year == 0) %>% 
    rename(starting.stand.type = stand.type) %>%
    dplyr::select(ru, fire.year, starting.stand.type)
  ind <- match(paste0(fire_dat  $ru, fire_dat  $fire.year, sep="."), 
               paste0(fire.dat.2$ru, fire.dat.2$fire.year, sep="."))
  fire_dat$starting.stand.type <- fire.dat.2$starting.stand.type[ind]
  fire_dat$rel.year <- as.factor(fire_dat$rel.year)
  rm(fire.dat.2)
  
  
  nofire_dat <- left_join(nofire_sol_dat, final.trees.wide, by=c("ru", "year"))
  
  # Save these datasets for later
  saveRDS(fire_dat, file="E:/storage/Winslow/for_Lora/fire_dat_spec.rds")
  saveRDS(nofire_dat, file="E:/storage/Winslow/for_Lora/nofire_dat_spec.rds")
} else {
  fire_dat <- readRDS("E:/storage/Winslow/for_Lora/fire_dat_spec.rds")
  nofire_dat <- readRDS("E:/storage/Winslow/for_Lora/nofire_dat_spec.rds")
}

# Histogram of never-burned pixels, by stand type
#ggplot(nofire_dat[!is.na(nofire_dat$stand.type),], aes(x=SOL.cm, after_stat(density))) + 
#  geom_histogram() +
#  theme_bw() + 
#  labs(title="Soil organic layer thickness for never-burned areas, all years",
#       x = "SOL (cm)") +
#  facet_wrap(~stand.type)

# Repeat for only the final 50 years of a run
ggplot(nofire_dat[(!is.na(nofire_dat$stand.type) & nofire_dat$year > 90),], aes(x=SOL.cm, after_stat(density))) + 
  geom_histogram() +
  theme_bw() + 
  labs(title=paste("Soil organic layer thickness for never-burned areas, years 260-300\n", tag),
       x = "SOL (cm)") +
  facet_wrap(~stand.type)

summary(fire_dat)
s=fire_dat %>% filter(starting.stand.type=="Pima" |starting.stand.type == "Pigl") %>% group_by(rel.year) %>% summarize(quantile.SOL =  quantile(SOL.cm, c(0.25,0.5,0.75)),q=c(0.25,0.5,0.75))
s.wide= s %>% pivot_wider(names_from = c(rel.year, q), values_from = quantile.SOL)
10.7 - 5.0
16.2 - 7.3
20.3-8.3

s=fire_dat %>% filter(starting.stand.type=="Potr" |starting.stand.type == "Bene") %>% group_by(rel.year) %>% summarize(quantile.SOL =  quantile(SOL.cm, c(0.25,0.5,0.75)),q=c(0.25,0.5,0.75))
s.wide= s %>% pivot_wider(names_from = c(rel.year, q), values_from = quantile.SOL)
3.1-2.9
3.6 - 3.3
4.1 - 3.8



```































