---
title: "Fire and soil organic layer evaluation"
author: "Lora Murphy"
date: "10/13/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# estSOL: black spruce = 0.1 and aspen/birch = 0.35

# Goal:
Plot modeled surface organic layer depth (live moss+dead moss+litter) with time since fire (30 years after fire) for stands dominated by black spruce, white spruce, and deciduous vegetation and compare to the average prefire surface organic layer depths for stands of the same forest types from inventory data of black spruce and published results for deciduous tree species. 


```{r echo=FALSE, message=FALSE}
library(RSQLite)
library(tidyverse)
library(rgdal)
library(ggplot2)
library(raster)

# Whether we've saved the water and soil table or need to create it again
data_saved <- T
key <- "spec"
tag <- "PIMA 0.1 decid 0.35"
```

```{r}
output_file <- "D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/cpcrw/model/output/paper_run/CPCRW_sm_6.sqlite"
#output_file <- "C:/Users/lora/Documents/Winslow/for_Lora/output.sqlite"
#-----------------------------------------------------------------------------#
# Load iLand fire table
#-----------------------------------------------------------------------------#
db.conn <- DBI::dbConnect(RSQLite::SQLite(), dbname=output_file)
fire <- dbReadTable(db.conn, "fire")
dbDisconnect(db.conn)
fire$area_ha <- fire$area_m2 / 10000
```

# Assembling a dataset of burned stuff, one year pre-fire to 30 years post-fire:

```{r}
if (!data_saved) {
  #---------------------------------------------------------------------------#
  # Read in the water table, which has info on soil organic layer; get this 
  # info for specific burned areas
  #---------------------------------------------------------------------------#
  db.conn <- DBI::dbConnect(RSQLite::SQLite(), dbname=output_file)
  water <- tbl(db.conn, "water") %>% collect()
  dbDisconnect(db.conn)
  
  # Let's get all the spread files (fire extents) as a raster stack
  path = "D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/cpcrw/model/output/paper_run/spread_6"
  #path = "E:/storage/Lora/from_winslow/output/spread"
  fire_maps = stack(list.files(path, full.names = T))
  # Use the environment grid to get ids for each grid
  path = "D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/cpcrw/materials/env"
  env_grid <- raster(paste0(path, "/env.grid.tif"))
  crs(fire_maps)=crs(env_grid)
  new.extent = extent(fire_maps)
  
#env_grid.2=crop(env_grid,new.extent)
  env_grid.2=env_grid                  ###This is where you adjust for a subset of the landscape or whole landscape

  fire <- fire %>% mutate(prop.dens.killed = n_trees_died   / n_trees,
                          prop.ba.killed   = basalArea_died / basalArea_total)
  fire$prop.ba.killed  [fire$basalArea_total == 0] <- 0
  fire$prop.dens.killed[fire$n_trees         == 0] <- 0
  
  #-----------------------------------------------------------------------------#
  # Build a dataset of SOL the year before a burn through 30 years after it. 
  #-----------------------------------------------------------------------------#
  fire_sol_dat <- NULL
  for (i in 1:nrow(fire)) {
    
    # Skip fires that burned nothing
    if (fire$area_m2[i] > 0) {
      
      # Take the fire map for this fire, and make everything but the fire cells NA
      mask <- fire_maps[[paste0("spread_", fire$fireId[i])]]
      mask[mask==0] <- NA
      
      # Use the mask to extract the rids for burned grids
      burned_rids <- getValues(raster::mask(env_grid.2, mask))
      burned_rids <- burned_rids[!is.na(burned_rids)]
      
      if (length(burned_rids) > 0) { # should never happen but...
        
        # Years we want info: year before the fire, through 30 years post-fire
        fire_years <- (fire$year[i] - 1):(fire$year[i] + 30)
        
        firedat <- water %>% filter(year %in% fire_years  &
                                      rid %in%  burned_rids)
        firedat$prop.dens.killed <- fire$prop.dens.killed[i]
        firedat$prop.ba.killed   <- fire$prop.ba.killed  [i]
        
        # Add the year of the record relative to the fire, and soil organic layer
        # thickness (go from m to cm)
        firedat <- firedat %>% 
          mutate(rel.year  = year - fire$year[i],
                 fireId    = fire$fireId[i],
                 fire.year = fire$year[i],
                 SOLLayer  = ifelse(is.na(SOLLayer) ,0,SOLLayer),
                 mossLayer = ifelse(is.na(mossLayer),0,mossLayer),
                 SOL.cm    = SOLLayer*100+mossLayer*100)
        
        fire_sol_dat <- rbind(fire_sol_dat,
                              firedat %>% dplyr::select(year, ru, rid,
                                                        fireId, rel.year, 
                                                        fire.year, 
                                                        SOLLayer, mossLayer,
                                                        SOL.cm,
                                                        prop.dens.killed,
                                                        prop.ba.killed))
        
      } else {
        stop("Better check on this...")
      }
      
    }
  }
  
  
  #---------------------------------------------------------------------------#
  # Also assemble a dataset of unburned pixels, for comparison
  #---------------------------------------------------------------------------#
  
  # Determine where fires happened anywhere
  fires <- sum(fire_maps) 
  
  # Flip around for the unburned pixels - this is where fires never happened
  mask <- sum(fire_maps) 
  mask[mask>0] <- NA
  
  # Use the mask to extract the rids for unburned grids
  unburned_rids <- getValues(raster::mask(env_grid.2, mask))
  unburned_rids <- unburned_rids[!is.na(unburned_rids)]
  
  nofire_sol_dat <- water                                                  %>% 
    filter(rid %in% unburned_rids)                         %>%
    mutate(SOLLayer  = ifelse(is.na(SOLLayer),0,SOLLayer),
           SOL.cm    = SOLLayer*100+mossLayer*100)         %>%
    dplyr::select(year, ru, rid, SOL.cm)
  
  rm(water)
}
```


Now we calculate species importance value. This is the way to determine which species was/is dominant...
Just do this for every resource unit, every year

```{r}
if (!data_saved) {
  #---------------------------------------------------------------------------#
  # Processing of sapling and stand tables. Sapling detail in the newest 
  # version is far too large to load completely, even in the capacious memory
  # of the workstation. So it is loaded and summarized in chunks.
  #---------------------------------------------------------------------------#
  
  # Get the stand and sapling tables and chop them down some
  db.conn <- DBI::dbConnect(RSQLite::SQLite(), dbname=output_file)
  stand          <- tbl(db.conn, "stand")         %>% collect() %>% filter(ru != -1) %>%                          
    dplyr::select(year, ru, rid, species, area_ha, count_ha, basal_area_m2)
  sapling        <- tbl(db.conn, "sapling")       %>% collect() %>% filter(ru != -1) %>%
    dplyr::select(ru, year, species, "sapling_count_ha"=count_ha, 
                  "sapling_count_ha_small" = count_small_ha)
  #sapling.detail <- tbl(db.conn, "saplingdetail") %>% filter(ru != -1) %>% collect()
  sapling.detail <- tbl(db.conn, "saplingdetail") %>% filter(year %in% 1:10) %>% collect()
  
  
  DBI::dbDisconnect(db.conn)
  
  # Add basal area to sapling.detail
  sapling.detail    = sapling.detail %>% mutate(basal.area.ind   = pi*((dbh/100)/2)^2)
  sapling.detail    = sapling.detail %>% mutate(basal.area.cohort= basal.area.ind*n_represented)
  
  # Group sapling info by year, location, species
  sapling.detail.sum= sapling.detail %>% group_by(year,ru, species) %>%   
    summarize(dbh_mean=weighted.mean(dbh,n_represented),
              basal.area.total=sum(basal.area.cohort))
  
  start = 11
  end = start + 9
  while (start < 300) {
    db.conn <- DBI::dbConnect(RSQLite::SQLite(), dbname=output_file)
    sapling.detail <- tbl(db.conn, "saplingdetail") %>% filter(year %in% start:end) %>% collect()
    DBI::dbDisconnect(db.conn)
  
  # Add basal area to sapling.detail
  sapling.detail    = sapling.detail %>% mutate(basal.area.ind   = pi*((dbh/100)/2)^2)
  sapling.detail    = sapling.detail %>% mutate(basal.area.cohort= basal.area.ind*n_represented)
  
  # Group sapling info by year, location, species
  sapling.detail.sum= rbind(sapling.detail.sum, 
                            sapling.detail %>% group_by(year,ru, species) %>%   
                            summarize(dbh_mean=weighted.mean(dbh,n_represented),
                            basal.area.total=sum(basal.area.cohort)))
  start <- start + 10
  end = start + 9
  }
  
  # Join all this up
  stand.sap.1=full_join(stand       ,sapling.detail.sum, by = c("ru","year","species"))
  final.trees=full_join(stand.sap.1 ,sapling,            by = c("ru","year","species"))
  final.trees=final.trees %>% replace(is.na(.), 0)
  
 # summary(final.trees)
  
  # Total density and basal area sums; this here will be by species
  final.trees = final.trees %>%
    mutate(count_total    = count_ha + sapling_count_ha + sapling_count_ha_small,
           basal.area_sum = basal_area_m2 + basal.area.total)
  
  # Widen table so each record has all species in it
  final.trees.wide = final.trees %>% 
    pivot_wider(names_from  = species, 
                names_sep   = ".", 
                values_from = c(area_ha:basal.area_sum), 
                values_fill = 0)
  #summary(final.trees.wide)
  
  # Final sums across species
  final.trees.wide = final.trees.wide %>% 
    mutate(total_density = count_total.Potr + count_total.Pima + count_total.Bene + count_total.Pigl,
           total_ba = basal.area_sum.Potr + basal.area_sum.Pima + basal.area_sum.Bene + basal.area_sum.Pigl)
  
  # Final importance value
  final.trees.wide = final.trees.wide%>%
    mutate(IV.Pima = ifelse(total_density !=0 & total_ba!=0, 
                            (count_total.Pima/total_density)+(basal.area_sum.Pima/total_ba),0),
           IV.Pigl = ifelse(total_density !=0 & total_ba!=0,
                            (count_total.Pigl/total_density)+(basal.area_sum.Pigl/total_ba),0),
           IV.Potr = ifelse(total_density !=0 & total_ba!=0,
                            (count_total.Potr/total_density)+(basal.area_sum.Potr/total_ba), 0),
           IV.Bene = ifelse(total_density !=0 &total_ba!=0,
                            (count_total.Bene/total_density)+(basal.area_sum.Bene/total_ba), 0))
  
  # Smallen the dataset a bit
  #final.trees.wide2 = final.trees.wide %>% dplyr::select(year, ru, count_total.Pima:IV.Bene) 
  #summary(final.trees.wide2)
  
  # Assign a stand type by whatever is the species with the highest IV
  iv_cols <- grep("IV", names(final.trees.wide))
  spp <- sapply(names(final.trees.wide)[iv_cols], function(x){strsplit(x, ".", fixed=T)[[1]][2]})
  ind <- apply(final.trees.wide[,iv_cols], MARGIN = 1, FUN = which.max)
  final.trees.wide$stand.type <- as.factor(spp[ind])
  
  rm(final.trees, #final.trees.wide, 
     stand, sapling, sapling.detail, stand.sap.1)
}
```

Join it up to our fire dataset and get some info by stand type

```{r}
if (!data_saved) {
  # We want to make sure we are looking at stand type at the right time. This will
  # give us the stand info along each year.
  fire_dat <- left_join(fire_sol_dat, final.trees.wide, by=c("ru","year"))
  
  # But we also want to know the stand type at the time of the fire, so we can
  # see if it changes. Add that as a separate variable.
  fire.dat.2 <- fire_dat %>% filter(rel.year == 0) %>% 
    rename(starting.stand.type = stand.type) %>%
    dplyr::select(ru, fire.year, starting.stand.type)
  ind <- match(paste0(fire_dat  $ru, fire_dat  $fire.year, sep="."), 
               paste0(fire.dat.2$ru, fire.dat.2$fire.year, sep="."))
  fire_dat$starting.stand.type <- fire.dat.2$starting.stand.type[ind]
  fire_dat$rel.year <- as.factor(fire_dat$rel.year)
  rm(fire.dat.2)
  
  
  nofire_dat <- left_join(nofire_sol_dat, final.trees.wide, by=c("ru", "year"))
  
  # Save these datasets for later
  saveRDS(fire_dat, file="E:/storage/Winslow/for_Lora/fire_dat_spec.rds")
  saveRDS(nofire_dat, file="E:/storage/Winslow/for_Lora/nofire_dat_spec.rds")
} else {
  fire_dat <- readRDS("E:/storage/Winslow/for_Lora/fire_dat_spec.rds")
  nofire_dat <- readRDS("E:/storage/Winslow/for_Lora/nofire_dat_spec.rds")
}

# Histogram of never-burned pixels, by stand type
#ggplot(nofire_dat[!is.na(nofire_dat$stand.type),], aes(x=SOL.cm, after_stat(density))) + 
#  geom_histogram() +
#  theme_bw() + 
#  labs(title="Soil organic layer thickness for never-burned areas, all years",
#       x = "SOL (cm)") +
#  facet_wrap(~stand.type)

# Repeat for only the final 50 years of a run
ggplot(nofire_dat[(!is.na(nofire_dat$stand.type) & nofire_dat$year > 90),], aes(x=SOL.cm, after_stat(density))) + 
  geom_histogram() +
  theme_bw() + 
  labs(title=paste("Soil organic layer thickness for never-burned areas, years 260-300\n", tag),
       x = "SOL (cm)") +
  facet_wrap(~stand.type)






```

Compare to the average prefire surface organic layer depths for stands of the same forest types from inventory data of black spruce and published results for deciduous tree species (7.6cm for paper birch Melvin et al. 2015). 

Use the prefire surface organic layer depth from the 90 stands in data_sets/Organic layer and regeneration. The environment file is called 342_JFSP_sitedata.txt, the seedling data is called 398_2004burns_seedlings.txt. They link through Site variable. Metadata available here: https://www.lter.uaf.edu/data/data-detail/id/398

Applicable site variables:

* Resid.org: Residual (postfire) organic layer depth (cm), measured at 11 random points per site by Leslie Boby in 2006. Missing data for DC33 and BF92 filled in by Jill Johnstone in 2007.

* PreFireOrg: Estimated pre-fire depth of the organic layer (cm), based on post-fire depth, corrected using adventitious root data (Leslie Boby dataset).

* %OrgLoss: Estimated fire consumption of the organic layer, based on comparisons of estimated pre-fire depth and measured residual depth (Leslie Boby dataset).


```{r fig.width=8}
sol_dat <- read.table("\\\\10.60.2.10\\FF_Lab\\project_data\\na_boreal\\data_sets\\Organic layer and tree regen\\Johnstone_LTER\\342_JFSP_sitedata.txt", sep="\t", header=T)
seed_dat <- read.csv("\\\\10.60.2.10\\FF_Lab\\project_data\\na_boreal\\data_sets\\Organic layer and tree regen\\Johnstone_LTER\\398_2004burns_seedlings.txt")
names(seed_dat)[1] <- "site"

# Weirdly, there's a negative basal area value; assume should be positive
sol_dat$PB.ba <- abs(sol_dat$PB.ba)

# Add the time since fire, since the post-fire SOL was measured in 2007
sol_dat$rel.year = as.factor(2007 - as.numeric(format(as.Date(sol_dat$Burn.date, format = "%d-%B-%y"),"%Y")))

# Add species IV, as best we can
sol_dat = sol_dat %>% 
  mutate(total_density = BS.dens + TA.dens + PB.dens,
         total_ba      = BS.ba   + TA.ba   + PB.ba) %>%
  mutate(IV.Pima = ifelse(total_density !=0 & total_ba!=0, 
                          (BS.dens/total_density)+(BS.ba/total_ba),0),
         IV.Potr = ifelse(total_density !=0 & total_ba!=0,
                          (TA.dens/total_density)+(TA.ba/total_ba), 0),
         IV.Bene = ifelse(total_density !=0 &total_ba!=0,
                          (PB.dens/total_density)+(PB.ba/total_ba), 0))

# Assign a stand type by whatever is the species with the highest IV
iv_cols <- grep("IV", names(sol_dat))
spp <- sapply(names(sol_dat)[iv_cols], function(x){strsplit(x, ".", fixed=T)[[1]][2]})
ind <- apply(sol_dat[,iv_cols], MARGIN = 1, FUN = which.max)
sol_dat$stand.type <- as.factor(spp[ind])
sol_dat.plot= sol_dat  %>% dplyr::select(rel.year, Resid.org, PreFireOrg, stand.type)
sol_dat.long = sol_dat.plot %>% pivot_longer(cols=c("PreFireOrg","Resid.org"),names_to = "layer", values_to = "depth")
summary(sol_dat.long)
sol_dat.long%>% group_by(layer,stand.type)%>%summarize(mean(depth,na.rm=T),
                                            median(depth,na.rm=T))
sol_dat.long = sol_dat.long %>% mutate(rel.year = as.factor(ifelse(layer=="Resid.org",3,-1)))

#-----------------------------------------------------------------------------#
# Make a dataset that will allow us to display iLand and observed data together
# ALL YEARS IN THE RUN
#-----------------------------------------------------------------------------#
df <- bind_rows(fire_dat %>% 
                  dplyr::select(rel.year, SOL.cm, starting.stand.type) %>% 
                  mutate(type="iLand"),
                sol_dat.long  %>% dplyr::select(rel.year, depth, stand.type) %>% 
                  rename(SOL.cm = depth, starting.stand.type = stand.type) %>% 
                  mutate(type="Observed"))
summary(df %>% filter(type=="Observed"))
#ggplot(data = df, aes(x=rel.year, y=SOL.cm, color=type)) +
#  geom_boxplot(outlier.shape = NA) +
#  scale_color_manual(values = c("black", "red")) +
#  ylim(c(0,45)) +
#  theme_bw() + 
#  labs(title = "Soil organic thickness in response to fire, all stands, all years; boxes are 25-75 percentiles",
#       x     = "Years since fire",
#       y     = "SOL (cm)")

ggplot(data = df[!is.na(df$starting.stand.type),],
       aes(x = rel.year, y=SOL.cm, color=type))  +
  scale_color_manual(values = c("black", "red")) +
  geom_boxplot(outlier.shape = NA) +
  ylim(c(0,45)) +
  theme_bw() + 
  labs(title = paste("Soil organic thickness in response to fire, by stand type, all years\n", tag),
       x     = "Years since fire",
       y     = "SOL (cm)") +
  facet_wrap(~starting.stand.type)



#-----------------------------------------------------------------------------#
# Make a dataset that will allow us to display iLand and observed data together
# LAST 50 YEARS ONLY
#-----------------------------------------------------------------------------#
df <- bind_rows(fire_dat %>% 
                  filter(year > 250) %>%
                  dplyr::select(rel.year, SOL.cm, starting.stand.type) %>% 
                  mutate(type="iLand"),
                sol_dat.long  %>% dplyr::select(rel.year, depth, stand.type) %>% 
                  rename(SOL.cm = depth, starting.stand.type = stand.type) %>% 
                  mutate(type="Observed"))
df %>% filter(type=="Observed")
ggplot(data = df, aes(x=rel.year, y=SOL.cm, color=type)) +
  #geom_boxplot(outlier.shape = NA) +
  geom_boxplot() +
  scale_color_manual(values = c("black", "red")) +
  ylim(c(0,45)) +
  theme_bw() + 
  labs(title = "Soil organic thickness in response to fire, all stands, years 251-300",
       x     = "Years since fire",
       y     = "SOL (cm)")
df$starting.stand.type <- factor(df$starting.stand.type, levels = c("Potr", "Bene", "Pima", "Pigl"))
df.summary = df %>% group_by(rel.year,starting.stand.type,type) %>% summarise(median.SOL.cm = median(SOL.cm, na.rm=T),
                                                                              sd.SOL.cm =sd(SOL.cm, na.rm=T))
df.summary=df.summary%>%filter(!is.na(starting.stand.type))
ggplot(data = df.summary,
       aes(x = rel.year, y=mean.SOL.cm, fill=type))  +
  geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=mean.SOL.cm-sd.SOL.cm, ymax=mean.SOL.cm+sd.SOL.cm), width=.2,
                 position=position_dodge(.9))+
  geom_boxplot() +
  ylim(c(0,35)) +
   
  theme_bw() + 
  scale_fill_manual(values=c("#1b7837","#762a83"),labels=c("Modeled", "Observed"))+
  theme(    text  = element_text(color="black",  family="sans", size=10),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
             axis.text.x=element_text(colour="black",size=10),
             axis.text.y=element_text(colour="black",size=10),
            legend.text = element_text(size=8),
             legend.position= "bottom",
            legend.background = element_rect(colour = "transparent", fill = 'transparent'),
            legend.key = element_rect(colour = "transparent", fill = 'transparent'),
            legend.key.size = unit(0.5, "cm"))+
  labs(x     = "Years since fire",
       y     = "Surface organic layer (cm)") +
  facet_wrap(~starting.stand.type, nrow=4)


#-----------------------------------------------------------------------------#
# Ribbon plot of SOL thickness
#-----------------------------------------------------------------------------#
df <- df %>% filter(!is.na(starting.stand.type) & !is.na(SOL.cm))
df %>% group_by(starting.stand.type, rel.year) %>% summarize(n=length(starting.stand.type))

df$rel.year <- as.numeric(df$rel.year)

# Calculate median, 25% and 75% lines, for iLand only
df_ribbon <- df %>% 
  filter(type == "iLand") %>% 
  group_by(rel.year, starting.stand.type) %>%
  summarise(lev25SOL = quantile(x=SOL.cm, probs=0.25),
            medianSOL = median(SOL.cm),
            lev75SOL = quantile(x=SOL.cm, probs=0.75))
df_ribbon$rel.year <- as.numeric(df_ribbon$rel.year)

# Get outliers. Find out what the quantiles are for each observation
#df$lev25 <- NA
#for (x in 1:nrow(df)) {
#  i <- which(df_ribbon$rel.year == df$rel.year[x] &
#             df_ribbon$starting.stand.type == df$starting.stand.type[x])
#  if (length(i)==1) df$lev25[x] = df_ribbon$lev25SOL[i]
#}
#df$lev75 <- NA
#for (x in 1:nrow(df)) {
#  i <- which(df_ribbon$rel.year == df$rel.year[x] &
#             df_ribbon$starting.stand.type == df$starting.stand.type[x])
#  if (length(i)==1) df$lev75[x] = df_ribbon$lev75SOL[i]
#}
# Use that info to mark whether an observation is an outlier. A lot of them are
#df$outlier <- ifelse(df$SOL.cm < df$lev25 | df$SOL.cm > df$lev75, T, F)
#df_outlier <- df %>% filter(outlier==T & type == "iLand")
#df_outlier$rel.year <- as.numeric(df_outlier$rel.year)

#----- Ribbon plot -----------------------------------------------------------#\
# Many datasets being used here!
ggplot(data = NULL)  +
  # Plot the ribbon
  geom_ribbon(data = df_ribbon, 
              aes(x = rel.year, ymin = lev25SOL, ymax = lev75SOL), fill = "#1b7837", alpha=0.8) +
  
  # Plot a median line on top
  geom_line(data = df_ribbon, aes(x = rel.year, y = medianSOL)) +
  
  # For the observed - a ribbon doesn't work with just a couple observations,
  # so put a boxplot on top of that
  geom_boxplot(data = (df %>% filter(type=="Observed")), 
               aes(x = rel.year, y=SOL.cm, group=rel.year), fill="#762a83") +
  
  # All the outliers; make mostly transparent so they're easier to see
  #geom_point(data = df_outlier, aes(x = rel.year, y = SOL.cm), alpha=0.1) +
  theme_bw() + 
  
  # I'm leaving this in but it doesn't do anything. We don't have a legend
  # at this point. Gotta figure that one out
  theme(    text  = element_text(color="black",  family="sans", size=10),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
             axis.text.x=element_text(colour="black",size=10),
             axis.text.y=element_text(colour="black",size=10),
            legend.text = element_text(size=8),
             legend.position= "bottom",
            legend.background = element_rect(colour = "transparent", fill = 'transparent'),
            legend.key = element_rect(colour = "transparent", fill = 'transparent'),
            legend.key.size = unit(0.5, "cm"))+
  labs(x     = "Years since fire",
       y     = "Surface organic layer (cm)") +
  
  # Scales = free should allow each plot to rescale itself, but with the 
  # outliers that doesn't really happen
  facet_wrap(~starting.stand.type, nrow=4, scales = "free")
ggsave("C:/Users/lora/Documents/Winslow/for_Lora/postfire_sol.jpg", height=6.5,width=8.5, units = "in", dpi=700)



```

# Regeneration
```{r}
#-----------------------------------------------------------------------------#
# Dataset of observed seedling data
#-----------------------------------------------------------------------------#
obs_dat <- left_join(
  seed_dat %>% rename(TA.nseedl = TA.tot, 
                      PB.nseedl = PB.tot, 
                      BS.nseedl = BS.tot,
                      TA.seedl.dens = TA.dens,
                      PB.seedl.dens = PB.dens,
                      BS.seedl.dens = BS.dens),
  sol_dat, by="site")

# Add fire severity info
df <- left_join(seed_dat %>% pivot_longer(cols = ends_with(".dens")),
                sol_dat  %>% dplyr::select(site, soil.sev, Burn.date), 
                by="site") %>% 
      filter(!is.na(soil.sev)) %>% 
      rename(species = name, seedl.dens = value)

# Make the species nice and have them match our species designations
df$species <- gsub(".dens", "", df$species, fixed = T)
df$species[df$species=="PB"] <- "Bene"
df$species[df$species=="TA"] <- "Potr"
df$species[df$species=="BS"] <- "Pima"

# Make the fire severity on soil a correctly-ordered factor, so it will graph 
# well
df$soil.sev <- factor(df$soil.sev, 
                      levels = c("low", "mod-low", "moderate", "mod-high", "high"),
                      ordered = T)
df$seedl.dens <- as.numeric(df$seedl.dens) # NAs introduced by coercion are in fact NA

# Add the time since fire, since the post-fire SOL was measured in 2007
df$rel.year = as.factor(df$Year - as.numeric(format(as.Date(df$Burn.date, format = "%d-%B-%y"),"%Y")))


# Get the mean seedling density by species, year, and fire severity
df2 <- df %>% group_by(soil.sev, species, rel.year) %>% 
       summarise(avg.seedl.den = mean(seedl.dens, na.rm=T))


# Plot actual seedling data by fire severity
ggplot(data = df2, aes(x = soil.sev, y=avg.seedl.den)) +
  geom_col(aes(fill=species)) +
  theme_bw() +
  labs(title = "Observed seedling regeneration post-fire, by years since fire",
       x = "Fire severity on soil", y = "Mean Seedlings / m2") +
  scale_fill_manual(values = c("darkred", "blue", "darkgreen")) +
  facet_wrap(~rel.year)
```

An attempt to compare our data. First things first: classify fire severity, based on its effects on soil. The observed regeneration dataset from the Bonanza Creek LTER had a rough assessmet of this. Our challenge: to apply similar levels to the iLand data in order to compare apples to apples.

soil.sev codes:

* Low: no min soil exposed, thick residual organic layers
* Mod-low: up to 5% min soil exposure and/or shallower (<10 cm) but mostly continuous organic layers
* Moderate: 10-20% min soil and/or mostly thin residual organic layers (<3 cm)
* Mod-high: 20-50% mineral soil exposed
* High: Over 50% mineral soil exposed

My attempt to quantify this by SOL thickness (SOL plus moss), the year post-fire (rel.year = 1). Note that this will not take into account amount actually lost - just trying to get a direct comparison going

* Low: >= 10 cm thick
* Mod-low: 3-10 cm thick
* Moderate: 1.5-3 cm thick
* Mod-high: 0.5 - 1.5 cm thick
* High: SOL < 0.5 cm

```{r}

#-----------------------------------------------------------------------------#
# Apples-to-apples categorized seedling density and fire severity
#-----------------------------------------------------------------------------#
# Apply fire intensity code, and get seedling density in #/m2
levels <- c("low", "mod-low", "moderate", "mod-high", "high")
df <- fire_dat %>% rename(Bene.nseedl = sapling_count_ha_small.Bene,
                          Potr.nseedl = sapling_count_ha_small.Potr,
                          Pima.nseedl = sapling_count_ha_small.Pima,
                          Pigl.nseedl = sapling_count_ha_small.Pigl) %>%
                   filter(year > 100) %>%
                   dplyr::select(year, ru, fireId, rel.year, 
                                 Bene.nseedl, Potr.nseedl, Pima.nseedl,
                                 Pigl.nseedl) %>%
                   mutate(Pigl.seedl.dens = 
                            ifelse(is.na(Pigl.nseedl), 0, Pigl.nseedl/10000),
                          Pima.seedl.dens = 
                            ifelse(is.na(Pima.nseedl), 0, Pima.nseedl/10000),
                          Potr.seedl.dens = 
                            ifelse(is.na(Potr.nseedl), 0, Potr.nseedl/10000),
                          Bene.seedl.dens = 
                            ifelse(is.na(Bene.nseedl), 0, Bene.nseedl/10000))

# Create the fire severity code based on SOL thickness in the first year after
# a fire
df2 <- fire_dat %>% 
  filter(rel.year == 1) %>% 
  mutate(soil.sev = levels[cut(SOL.cm, breaks = c(0,0.5,1.5,3,10,100))]) %>%
  dplyr::select(ru, fireId, soil.sev)

# Add the fire severity to the seedling data and subset to the same years
# as the observed data
df <- left_join(df, df2, by=c("ru", "fireId")) %>% 
  filter(rel.year %in% c(2,4,7,13)) %>% 
  pivot_longer(cols = ends_with(".dens")) %>% 
  rename(species = name, seedl.dens = value)
df$species = gsub(".seedl.dens", "", df$species, fixed=T)
df$species = factor(df$species,
                    levels = c("Bene", "Pima", "Potr", "Pigl"),
                    ordered = T)
            
df$soil.sev <- factor(df$soil.sev, 
                      levels = c("low", "mod-low", "moderate", "mod-high", "high"),
                      ordered = T)

df2 <- df %>% group_by(soil.sev, species, rel.year) %>% 
       summarise(avg.seedl.den = mean(seedl.dens, na.rm=T))

# Plot iLand seedling data by fire severity
ggplot(data = df2, aes(x = soil.sev, y=avg.seedl.den)) +
  geom_col(aes(fill=species)) +
  theme_bw() +
  scale_fill_manual(values = c("darkred", "blue", "darkgreen", "orange")) +
  labs(title = paste("iLand seedling regeneration post-fire, by years since fire, years 251-300\n", tag),
       x = "Fire severity on soil", y = "Mean Seedlings / m2") +
  facet_wrap(~rel.year)

```



Compare by residual soil organic layer depth; group into classes. Winslow notes: By 13 years postfire, a lot of the deciduous species will have grown out of the count_small_ha (which is the sapling cohort shorter than 1.3 m). So to get total saplings we need to do count_ha + count_small_ha from the sapling fire which will give us all saplings shorter than 4m.

FILTER DOWN TO LAST 100 YEARS OF SIMULATION PER MANUSCRIPT
```{r}
# iLand: seedling density in #/m2
df <- fire_dat %>% mutate(sap1.Bene = ifelse(is.na(sapling_count_ha_small.Bene), 0, 
                                                   sapling_count_ha_small.Bene),
                          sap2.Bene = ifelse(is.na(sapling_count_ha.Bene), 0, 
                                                   sapling_count_ha.Bene),
                          sap1.Potr = ifelse(is.na(sapling_count_ha_small.Potr), 0, 
                                                   sapling_count_ha_small.Potr),
                          sap2.Potr = ifelse(is.na(sapling_count_ha.Potr), 0, 
                                                   sapling_count_ha.Potr),
                          sap1.Pima = ifelse(is.na(sapling_count_ha_small.Pima), 0, 
                                                   sapling_count_ha_small.Pima),
                          sap2.Pima = ifelse(is.na(sapling_count_ha.Pima), 0, 
                                                   sapling_count_ha.Pima),
                          sap1.Pigl = ifelse(is.na(sapling_count_ha_small.Pigl), 0, 
                                                   sapling_count_ha_small.Pigl),
                          sap2.Pigl = ifelse(is.na(sapling_count_ha.Pigl), 0, 
                                                   sapling_count_ha.Pigl)) %>%
                   mutate(Bene.nseedl = (sap1.Bene + sap2.Bene),
                          Potr.nseedl = (sap1.Potr + sap2.Potr),
                          Pima.nseedl = (sap1.Pima + sap2.Pima),
                          Pigl.nseedl = (sap1.Pigl + sap2.Pigl)) %>%
                   dplyr::select(year, ru, fireId, rel.year, SOL.cm,
                                 Bene.nseedl, Potr.nseedl, Pima.nseedl,
                                 Pigl.nseedl) %>%
                   filter(rel.year %in% c(2,4,7,13) & year > 200) %>% 
                   mutate(Pigl.seedl.dens = 
                            ifelse(is.na(Pigl.nseedl), 0, Pigl.nseedl/10000),
                          Pima.seedl.dens = 
                            ifelse(is.na(Pima.nseedl), 0, Pima.nseedl/10000),
                          Potr.seedl.dens = 
                            ifelse(is.na(Potr.nseedl), 0, Potr.nseedl/10000),
                          Bene.seedl.dens = 
                            ifelse(is.na(Bene.nseedl), 0, Bene.nseedl/10000) )%>% 
                   pivot_longer(cols = ends_with(".dens")) %>% 
                   rename(species = name, seedl.dens = value)
df$species = gsub(".seedl.dens", "", df$species, fixed=T)


# Observed data
obs_dat <- left_join(seed_dat %>% rename(TA.nseedl = TA.tot, 
                                         PB.nseedl = PB.tot, 
                                         BS.nseedl = BS.tot,
                                         TA.seedl.dens = TA.dens,
                                         PB.seedl.dens = PB.dens,
                                         BS.seedl.dens = BS.dens),
                     sol_dat %>% dplyr::select(site, Resid.org, Burn.date, rel.year), 
                     by="site")
# Add the time since fire, since the post-fire SOL was measured in 2007
obs_dat$rel.year = as.factor(obs_dat$Year - as.numeric(format(as.Date(obs_dat$Burn.date, format = "%d-%B-%y"),"%Y")))

# Lengthen the table for ggplot
df2 <- obs_dat %>% #filter(rel.year %in% c(2,4,7,13)) %>% 
            pivot_longer(cols = ends_with(".dens")) %>% 
            rename(species = name, seedl.dens = value)
# Make the species nice and have them match our species designations
df2$species <- gsub(".seedl.dens", "", df2$species, fixed = T)
df2$species[df2$species=="PB"] <- "Bene"
df2$species[df2$species=="TA"] <- "Potr"
df2$species[df2$species=="BS"] <- "Pima"
df2$seedl.dens <- as.numeric(df2$seedl.dens) # NAs introduced by coercion are in fact NA

# Join up iLand and observed so we can plot them together
df <- bind_rows(df  %>% dplyr::select(year, SOL.cm, species, seedl.dens, rel.year) %>% 
                       mutate(type="iLand"),
                df2 %>% dplyr::select(Year, Resid.org, species, seedl.dens, rel.year) %>%
                       rename(SOL.cm = Resid.org, year = Year) %>%
                       mutate(type = "Observed")) %>%
      mutate(soil.thick = cut(SOL.cm, breaks = seq(0,max(SOL.cm),by=5)))
levels(df$soil.thick) <- sapply(levels(df$soil.thick), 
                    function(x){gsub("]","",strsplit(x, ",")[[1]][2], fixed=T)})


df$species <- factor(df$species, levels = c("Pima","Pigl","Potr","Bene"))

s.plot = df %>% filter(rel.year==2 & !is.na(soil.thick) & year >200) 


# g4 <- ggplot(data = s.plot, 
#        aes(x=soil.thick, y=seedl.dens, fill=species)) +
#   geom_boxplot() +
#   ylim(c(0,50)) +
#   theme_bw() + 
#  scale_fill_manual(values=c("#f0027f","#386cb0","#33a02c","#ffff33","#ff7f00"),labels = c("Pima","Pigl","Potr","Bene"))+ 
#   theme(    text  = element_text(color="black",  family="sans", size=10),
#             panel.grid.major = element_blank(),
#             panel.grid.minor = element_blank(),
#              axis.text.x=element_text(colour="black",size=10),
#              axis.text.y=element_text(colour="black",size=10),
#             legend.text = element_text(size=8),
#              legend.position= "bottom",
#             legend.background = element_rect(colour = "transparent", fill = 'transparent'),
#             legend.key = element_rect(colour = "transparent", fill = 'transparent'),
#             legend.key.size = unit(0.5, "cm"))+
#   labs(x     = "Surface organic layer (cm)",
#        y     = bquote(Seedlings~m^2)) +
#   facet_wrap(~type)

#-----------------------------------------------------------------------------#
# Winslow: It would be better to switch this: put the iland and observed 
# together and make a separate panel for each species.
#-----------------------------------------------------------------------------#

# LEM: Some of these outliers absolutely kill the Y axis. I'm filtering them out
# this way
g4 <- ggplot(data = s.plot %>% filter(seedl.dens < 50), 
       aes(x=soil.thick, y=seedl.dens, fill=type)) +
  geom_boxplot() +
  #ylim(c(0,50)) +
  theme_bw() + 
 scale_fill_manual(values=c("#1b7837","#762a83"),labels=c("Modeled", "Observed"))+ 
  theme(    text  = element_text(color="black",  family="sans", size=10),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
             axis.text.x=element_text(colour="black",size=10),
             axis.text.y=element_text(colour="black",size=10),
            legend.text = element_text(size=8),
             legend.position= "bottom",
            legend.background = element_rect(colour = "transparent", fill = 'transparent'),
            legend.key = element_rect(colour = "transparent", fill = 'transparent'),
            legend.key.size = unit(0.5, "cm"))+
  labs(x     = "Surface organic layer (cm)",
       y     = bquote(Seedlings~m^-2)) +
  facet_wrap(~species, scales="free")
g4 
ggsave("D:/workspace/Winslow/permafrost/permafrost_param/iLand_landscapes/landscape_building/figures/postfire_seedling.jpg", height=4.5,width=6.5, units = "in", dpi=700)
#list species seedling densities
s.plot = df %>% filter(rel.year==2 & !is.na(soil.thick) & year >200) %>% group_by(species, type, soil.thick) %>% summarize(density=quantile(seedl.dens, c(0.25,0.5,0.75)),q=c(0.25,0.5,0.75))
s.plot %>%filter(species=="Pigl")
s.plot %>% filter(soil.thick==5 &type=="iLand")


```
Looking at iLand seedling numbers
```{r}
# iLand: seedling density in #/m2
df <- fire_dat %>% rename(Bene.nseedl = sapling_count_ha_small.Bene,
                          Potr.nseedl = sapling_count_ha_small.Potr,
                          Pima.nseedl = sapling_count_ha_small.Pima,
                          Pigl.nseedl = sapling_count_ha_small.Pigl) %>%
                   dplyr::select(ru, fireId, rel.year, SOL.cm,
                                 Bene.nseedl, Potr.nseedl, Pima.nseedl,
                                 Pigl.nseedl) %>%
                   mutate(Pigl.seedl.dens = 
                            ifelse(is.na(Pigl.nseedl), 0, Pigl.nseedl/10000),
                          Pima.seedl.dens = 
                            ifelse(is.na(Pima.nseedl), 0, Pima.nseedl/10000),
                          Potr.seedl.dens = 
                            ifelse(is.na(Potr.nseedl), 0, Potr.nseedl/10000),
                          Bene.seedl.dens = 
                            ifelse(is.na(Bene.nseedl), 0, Bene.nseedl/10000) )%>% 
                   pivot_longer(cols = ends_with(".dens")) %>% 
                   rename(species = name, seedl.dens = value)
df$species = gsub(".seedl.dens", "", df$species, fixed=T)
df <- df %>% group_by(species, rel.year) %>% summarize(mean.seedl = mean(seedl.dens))

ggplot(data = df, aes(x = rel.year, y=mean.seedl)) +
  geom_col(aes(fill=species)) +
  theme_bw()
```
